---
title: "08_Analysis_Fish"
author: "Laura Givens"
date: "2025-05-28"
output: html_document
---
# Load  
Load libraries  
```{r,message=FALSE}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(vegan)
library(lme4)
library(patchwork)
```

```{r}
set.seed(999)
```

Load data  
```{r load}
dir_data <- '/Users/lauragivens/Desktop/R/BZrookery_eDNA/Rdata'
dir_results <- "/Volumes/Fuji/Mangroves/2025_0319_Givens_Canty_Rookery_COI/cutadapt/results"

ps.nt <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.nt.troph.rds"))
ps.mar <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.troph.rds"))
ps.bold <- readRDS(paste0(dir_results,"/prune.decontam.ps.bold.rds"))

ps.nt <- ps.nt %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")
ps.mar <- ps.mar %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")
ps.bold <- ps.bold %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")

taxa_nt <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.nt.troph.rds"))
taxa_mar <- readRDS(paste0(dir_results,"/taxtable.v2.wtroph.rds"))
taxa_bold <- readRDS(paste0(dir_results,"/taxtable.bold.rds"))

meta <- readRDS(paste0(dir_results,"/metadata.edit.rds"))
samplelist <- meta %>% filter(Rookery!="Practice" & Rookery !="Blank")
```

Or alternatively,  
```{r}
load("RData/08_Analysis_Fish.RData")
```

Convert rookery to 0/1  
```{r}
samplelist<- samplelist %>% mutate(RookeryBinary = case_when(Rookery=="Rookery" ~ 1,
                                Rookery!="Rookery" ~ 0),.after=Rookery )
samplelist <- samplelist %>% mutate(NewName=word(rownames(.),sep="_",end=2))

sample_data(ps.nt) <- samplelist
sample_data(ps.mar) <- samplelist
sample_data(ps.bold) <- samplelist
```

```{r}
ggplot2::theme_set(theme_cowplot() + theme(
  legend.position = "bottom",
  panel.border = element_rect(color='black'),
  legend.box.margin = margin(5,5,5,5),
  legend.spacing = unit(0,"pt"),
  plot.title = element_text(face="bold"),
  plot.caption = element_text(face="italic"),
  text = element_text(size=(15),family="Times"),
  legend.key.height = unit(10,"pt"),
  legend.key.width = unit(10,"pt")
))
```

# Prune to fish   
Select only fish BOLD samples (since the mar dataset is already only chordates)
May need to further subsample this actually
```{r}
ps.bold.fsh <- ps.bold %>% subset_taxa(Phylum=="Chordata" & Class!="Ascidiacea" & Class != "Reptilia" & Class != "Aves") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
```

Select only fish nt samples (since the mar dataset is already only chordates)
May need to further subsample this actually
```{r}
ps.nt.fsh <- ps.nt %>% subset_taxa(Phylum=="Chordata" & Class=="Actinopteri" |  Class == "Chondrichthyes") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
```

# Conversions  
Presence/absence  
```{r}
ps.nt.pa <- ps.nt.fsh %>% transform_sample_counts(., function(abund) 1*(abund > 0))
ps.mar.pa <- ps.mar %>% transform_sample_counts(., function(abund) 1*(abund > 0))
ps.bold.pa <- ps.bold.fsh %>% transform_sample_counts(., function(abund) 1*(abund > 0))
```

Relative abundance  
```{r}
ps.nt.rel <- ps.nt.fsh %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.mar.rel <- ps.mar %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.bold.rel <- ps.bold.fsh %>% transform_sample_counts(.,function(x) x/sum(x) ) 
```

Log transform  
```{r}
ps.nt.log <- ps.nt.fsh %>% transform_sample_counts(., function(x) log(1 + x)) %>% 
      prune_samples(sample_sums(.) > 0,.)
ps.mar.log <- ps.mar %>% transform_sample_counts(., function(x) log(1 + x)) %>% 
      prune_samples(sample_sums(.) > 0,.)
ps.bold.log <- ps.bold.fsh %>% transform_sample_counts(., function(x) log(1 + x)) %>% 
      prune_samples(sample_sums(.) > 0,.)
```

```{r}
tab.bold <- otu_table(ps.bold.fsh) %>% as.matrix() %>% as.data.frame() %>% t()
tab.nt <- otu_table(ps.nt.fsh) %>% as.matrix() %>% as.data.frame() %>% t()
tab.mar <- otu_table(ps.mar) %>% as.matrix() %>% as.data.frame() %>% t()

tab.bold.pa <- otu_table(ps.bold.pa) %>% as.matrix() %>% as.data.frame() %>% t()
tab.bold.rel <- otu_table(ps.bold.rel) %>% as.matrix() %>% as.data.frame() %>% t()
tab.bold.log <- otu_table(ps.bold.log) %>% as.matrix() %>% as.data.frame() %>% t()

tab.nt.pa <- otu_table(ps.nt.pa) %>% as.matrix() %>% as.data.frame() %>% t()
tab.nt.rel <- otu_table(ps.nt.rel) %>% as.matrix() %>% as.data.frame() %>% t()
tab.nt.log <- otu_table(ps.nt.log) %>% as.matrix() %>% as.data.frame() %>% t()

tab.mar.pa <- otu_table(ps.mar.pa) %>% as.matrix() %>% as.data.frame() %>% t()
tab.mar.rel <- otu_table(ps.mar.rel) %>% as.matrix() %>% as.data.frame() %>% t()
tab.mar.log <- otu_table(ps.mar.log) %>% as.matrix() %>% as.data.frame() %>% t()

ps.bold.pa.melt <- psmelt(ps.bold.pa)
ps.nt.pa.melt <- psmelt(ps.nt.pa)
ps.mar.pa.melt <- psmelt(ps.mar.pa)

ps.bold.rel.melt <- psmelt(ps.bold.rel)
ps.nt.rel.melt <- psmelt(ps.nt.rel)
ps.mar.rel.melt <- psmelt(ps.mar.rel)

ps.nt.log.melt <- psmelt(ps.nt.log)
ps.bold.log.melt <- psmelt(ps.bold.log)
ps.mar.log.melt <- psmelt(ps.mar.log)
```

```{r}
#ps.mar.pa.melt %>% filter(Abundance > 0) %>% View()
write.csv((ps.mar.pa.melt %>% filter(Abundance > 0)),
          file="/Users/lauragivens/Downloads/bzrookery.fish.nozero.csv")
```

# Normality tests    

## Untransformed  
Shannon diversity index  
```{r}
rich.nt <- diversity(tab.nt,index="shannon")
rich.mar <- diversity(tab.mar,index="shannon")
rich.bold <- diversity(tab.bold,index="shannon")
```
The only one that passes is the nt dataset  
```{r}
shapiro.test(rich.nt)
shapiro.test(rich.mar)
shapiro.test(rich.bold)
```

```{r}
hist(rich.nt, main="Shannon diversity", xlab="", breaks=10)
hist(rich.mar, main="Shannon diversity", xlab="", breaks=10)
hist(rich.bold, main="Shannon diversity", xlab="", breaks=10)
```

Observed richness  
```{r}
obs.nt <- specnumber(tab.nt)
obs.bold <- specnumber(tab.bold)
obs.mar <- specnumber(tab.mar)
```
Again, the only one that passes is the nt dataset  
```{r}
shapiro.test(obs.nt)
shapiro.test(obs.bold)
shapiro.test(obs.mar)
```
```{r}
hist(obs.nt, main="Observed richness", xlab="", breaks=10)
hist(obs.mar, main="Observed richness", xlab="", breaks=10)
hist(obs.bold, main="Observed richness", xlab="", breaks=10)
```

```{r,eval=FALSE}
#with phyloseq
rich.nt <- estimate_richness(ps.nt.fsh,measures=c("Shannon","Observed"))
rich.bold <- estimate_richness(ps.bold.fsh,measures=c("Shannon","Observed"))
rich.mar <- estimate_richness(ps.mar,measures=c("Shannon","Observed"))

shapiro.test(rich.nt$Observed)
shapiro.test(rich.bold$Observed)
shapiro.test(rich.mar$Observed)
#Untransformed data passes the Shapiro test (p>0.05) for the nt and mar databases (barely for MAR) for Observed richness. We can also see this in the histogram:   
hist(rich.nt$Observed, main="Observed richness", xlab="", breaks=10)
hist(rich.bold$Observed, main="Observed richness", xlab="", breaks=10)
hist(rich.mar$Observed, main="Observed richness", xlab="", breaks=10)

shapiro.test(rich.nt$Shannon)
shapiro.test(rich.bold$Shannon)
shapiro.test(rich.mar$Shannon)
#For Shannon diversity, nt database is again ok but the other two are not.  
hist(rich.nt$Shannon, main="Shannon richness", xlab="", breaks=10)
hist(rich.bold$Shannon, main="Shannon richness", xlab="", breaks=10)
hist(rich.mar$Shannon, main="Shannon richness", xlab="", breaks=10)
```

## Presence/absence      
Shannon diversity index  
```{r}
rich.nt <- diversity(tab.nt.pa,index="shannon")
rich.mar <- diversity(tab.mar.pa,index="shannon")
rich.bold <- diversity(tab.bold.pa,index="shannon")
```
None of the presence/absence datasets pass the test for normality   
```{r}
shapiro.test(rich.nt)
shapiro.test(rich.mar)
shapiro.test(rich.bold)
```

```{r}
hist(rich.nt, main="Shannon diversity", xlab="", breaks=10)
hist(rich.mar, main="Shannon diversity", xlab="", breaks=10)
hist(rich.bold, main="Shannon diversity", xlab="", breaks=10)
```

Observed richness  
```{r}
obs.nt <- specnumber(tab.nt)
obs.bold <- specnumber(tab.bold)
obs.mar <- specnumber(tab.mar)
```
Core nt and MAR datasets pass, BOLD fails   
```{r}
shapiro.test(obs.nt)
shapiro.test(obs.bold)
shapiro.test(obs.mar)
```
```{r}
hist(obs.nt, main="Observed richness", xlab="", breaks=10)
hist(obs.mar, main="Observed richness", xlab="", breaks=10)
hist(obs.bold, main="Observed richness", xlab="", breaks=10)
```

## Relative abundance  
Shannon diversity index  
```{r}
rich.nt <- diversity(tab.nt.rel,index="shannon")
rich.mar <- diversity(tab.mar.rel,index="shannon")
rich.bold <- diversity(tab.bold.rel,index="shannon")
```
The only one that passes is the nt dataset  
```{r}
shapiro.test(rich.nt)
shapiro.test(rich.mar)
shapiro.test(rich.bold)
```

```{r}
hist(rich.nt, main="Shannon diversity", xlab="", breaks=10)
hist(rich.mar, main="Shannon diversity", xlab="", breaks=10)
hist(rich.bold, main="Shannon diversity", xlab="", breaks=10)
```

Observed richness  
```{r}
obs.nt <- specnumber(tab.nt)
obs.bold <- specnumber(tab.bold)
obs.mar <- specnumber(tab.mar)
```
core nt and mar datasets both pass   
```{r}
shapiro.test(obs.nt)
shapiro.test(obs.bold)
shapiro.test(obs.mar)
```
```{r}
hist(obs.nt, main="Observed richness", xlab="", breaks=10)
hist(obs.mar, main="Observed richness", xlab="", breaks=10)
hist(obs.bold, main="Observed richness", xlab="", breaks=10)
```

## Log  
Shannon diversity index  
```{r}
rich.nt <- diversity(tab.nt.log,index="shannon")
rich.mar <- diversity(tab.mar.log,index="shannon")
rich.bold <- diversity(tab.bold.log,index="shannon")
```
None pass    
```{r}
shapiro.test(rich.nt)
shapiro.test(rich.mar)
shapiro.test(rich.bold)
```

```{r}
hist(rich.nt, main="Shannon diversity", xlab="", breaks=10)
hist(rich.mar, main="Shannon diversity", xlab="", breaks=10)
hist(rich.bold, main="Shannon diversity", xlab="", breaks=10)
```

Observed richness  
```{r}
obs.nt <- specnumber(tab.nt)
obs.bold <- specnumber(tab.bold)
obs.mar <- specnumber(tab.mar)
```
core nt and mar datasets pass    
```{r}
shapiro.test(obs.nt)
shapiro.test(obs.bold)
shapiro.test(obs.mar)
```
```{r}
hist(obs.nt, main="Observed richness", xlab="", breaks=10)
hist(obs.mar, main="Observed richness", xlab="", breaks=10)
hist(obs.bold, main="Observed richness", xlab="", breaks=10)
```

# Comparisons  

## Percentage of class level assignments  
```{r}
bold.fsh.plot <- ps.bold.rel.melt %>% ungroup() %>% 
  filter(.,Abundance>0) %>%  group_by(.,Class) %>% 
  reframe(.,Class=unique(Class), 
            db="BOLDv5",
            Abund=sum(Abundance)) %>% 
  mutate(RelAbund=Abund/sum(Abund)) %>%
  ggplot(aes(x=db,y=RelAbund,fill=Class)) + 
  geom_bar(stat="identity",position="stack") + theme(axis.title.x=element_blank(),legend.text = element_text(size=8))

nt.fsh.plot <- ps.nt.rel.melt %>% ungroup() %>% 
  filter(.,Abundance>0) %>%  group_by(.,Class) %>% 
  reframe(.,Class=unique(Class), 
            db="core nt",
            Abund=sum(Abundance)) %>% 
  mutate(RelAbund=Abund/sum(Abund)) %>%
  ggplot(aes(x=db,y=RelAbund,fill=Class)) + 
  geom_bar(stat="identity",position="stack")+ theme(axis.title.x=element_blank(),axis.title.y=element_blank(),legend.position="none") 

mar.fsh.plot <- ps.mar.rel.melt %>% ungroup() %>% 
  filter(.,Abundance>0) %>%  group_by(.,Class) %>% 
  reframe(.,Class=unique(Class), 
            db="MAR fish nt alias",
            Abund=sum(Abundance)) %>% 
  mutate(RelAbund=Abund/sum(Abund)) %>%
  ggplot(aes(x=db,y=RelAbund,fill=Class)) + 
  geom_bar(stat="identity",position="stack") + theme(axis.title.x=element_blank(),axis.title.y=element_blank(),legend.text = element_text(size=8))

patchwork::wrap_plots(bold.fsh.plot,nt.fsh.plot,mar.fsh.plot,guides="collect") 
#ggsave("Images/taxassgn.comp.fsh.png",width=9,height=5)
```

## Species of importance   
```{r}
#str_detect(names(ps.mar.pa.melt),"Comm")

ps.mar.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  #select(contains("Commer") | contains("Small.scale")) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% 
  ggplot(aes(x=Month,fill=New,y=Rel)) + geom_bar(stat="identity",position="stack") + facet_grid(Site ~ Rookery) + guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) + theme_bw() + labs(y="Relative Abundance")
```

That is a lot of commercially relevant species. Does that track?  
```{r}
#ps.mar.pa.melt %>% filter(Abundance > 0) %>% select(c(Species,Reported.Commercial.FishbaseSpeciesPage)) %>% distinct() 

ps.mar.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  select(c(Species,contains("Commer") | contains("Small.scale") )) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% 
  rowwise() %>% 
  mutate(Relevant.Species= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% filter(!is.na(Species)) %>% distinct() # %>% write.csv(., file="/Users/lauragivens/Downloads/fishbase.commercialspecies.csv")
```

```{r}
ps.mar.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% 
  ggplot(aes(x=Month,fill=New,y=Rel)) + geom_bar(stat="identity",position="stack") + facet_grid(Site ~ Rookery) + guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) + theme_bw() + labs(y="Relative Abundance")
```

table of commercially important and small scale fisheries species  
```{r}
comm.smsc.fisheries.tab <- ps.mar.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  #select(contains("Commer") | contains("Small.scale")) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% filter(New=="TRUE" & Abundance > 0) %>%  select(c(Site,Rookery,Genus,Species)) %>% #select(c(Month,Site,Rookery,Side,Domain,Phylum,Class,Order,Family,Genus,Species)) %>% 
  distinct() %>% mutate(n=1) %>% pivot_wider(names_from=c(Site,Rookery),values_from=n) %>%
  select(c(Species,starts_with("Drown"),starts_with("Hicks"),starts_with("Turnef"))) %>%
  filter(!is.na(Species)) %>%
  arrange(Species) #%>%
  #select(c(Genus,Species,starts_with("Drown"),starts_with("Hicks"),starts_with("Turnef"))) %>%
  #arrange(Genus) %>% 
  #View()

write.csv(comm.smsc.fisheries.tab,file="/Users/lauragivens/Downloads/commercial.smallscale.bysite.csv")
```

## Species number by sampling effort  
```{r}
ps.mar.pa.melt %>% 
  group_by(Month,Rookery,Site,Species) %>%
  summarize(Abund=sum(Abundance),spec=unique(Species)) %>% 
  filter(Abund>0) %>% ungroup() %>% 
  group_by(Month,Rookery,Site) %>% 
  summarize(n=length(Species)) %>% pivot_wider(names_from="Month",values_from="n") %>% arrange(desc(Site)) %>% flextable::flextable()
```

## Leeward v windward  
```{r}
ps.mar.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>%
  rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% 
  ggplot(aes(x=Side,fill=New,y=Rel)) + 
  geom_bar(stat="identity",position="stack") + 
  facet_grid(Site ~ Rookery) + 
  guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) + theme_bw() + labs(y="Relative Abundance")
```

```{r}
plot_ordination(ps.mar.pa,ordinate(ps.mar.pa,method="NMDS"),color="Side") + facet_grid(Site~Rookery) + theme_bw()
#ggsave("/Users/lauragivens/Downloads/leeward.windward.nmds.png",width=7,height=5)
```

# Richness  
```{r}
richdf <- rowSums(tab.mar.pa) %>% 
  as.data.frame() %>% 
  `colnames<-` ("samplerichness") %>% 
  mutate(Sample=rownames(.)) %>% left_join(select((samplelist %>% mutate(Sample=rownames(.))),Sample,Month,Site,Rookery))
```
```{r}
wrap_plots(
ggplot(richdf) + geom_boxplot(aes(x=Month,y=samplerichness)) + labs(x="Month",y="Sample Richness"),
ggplot(richdf) + geom_boxplot(aes(x=Site,y=samplerichness)) + labs(x="Site",y="Sample Richness"),
ggplot(richdf) + geom_boxplot(aes(x=Rookery,y=samplerichness)) + labs(x="Rookery",y="Sample Richness")
)


wrap_plots(
ggplot(richdf) + geom_boxplot(aes(x=Site,y=samplerichness)) + labs(x="Site",y="Number of taxa"),
ggplot(richdf) + geom_boxplot(aes(x=Rookery,y=samplerichness)) + labs(x="Rookery",y="Number of taxa")
) & theme_bw() & theme(axis.title.x=element_blank())

#ggsave("Images/richness.boxplots.png",width=10,height=5,bg="white")
```

```{r}
ps.mar.pa.melt %>% 
  group_by(Rookery,Site,Species) %>%
  summarize(Abund=sum(Abundance),spec=unique(Species)) %>% 
  filter(Abund>0) %>% ungroup() %>% 
  group_by(Rookery,Site) %>% 
  summarize(n=length(Species)) %>% pivot_wider(names_from="Rookery",values_from="n")
```

```{r}
ps.mar.pa.melt %>% 
  group_by(Rookery,Site,Species) %>%
  summarize(Abund=sum(Abundance),spec=unique(Species)) %>% 
  filter(Abund>0) %>% ungroup() %>% 
  group_by(Rookery,Site) %>% 
  summarize(n=length(Species)) %>% pivot_wider(names_from="Rookery",values_from="n") %>% flextable::flextable()
```
```{r}
ps.mar.pa.melt %>% 
  group_by(Month,Rookery,Site,Species) %>%
  summarize(Abund=sum(Abundance),spec=unique(Species)) %>% 
  filter(Abund>0) %>% ungroup() %>% 
  group_by(Month,Rookery,Site) %>% 
  summarize(n=length(Species)) %>% pivot_wider(names_from="Month",values_from="n") %>% arrange(desc(Site)) %>% flextable::flextable()
```

```{r}
hc <- get_taxa_unique(ps.mar %>% subset_samples(Site=="Hicks Caye") %>% prune_taxa(taxa_sums(.)>0,.),'Species')
dc <- get_taxa_unique(ps.mar %>% subset_samples(Site=="Drowned Caye") %>% prune_taxa(taxa_sums(.)>0,.),'Species')
ta <- get_taxa_unique(ps.mar %>% subset_samples(Site=="Turneffe Atoll") %>% prune_taxa(taxa_sums(.)>0,.),'Species')

hc[!hc%in%c(dc,ta)]
dc[!dc%in%c(hc,ta)]
ta[!ta%in%c(hc,dc)]
```

```{r}
kruskal.test(richdf$samplerichness ~ richdf$Month)
kruskal.test(richdf$samplerichness ~ richdf$Site)
kruskal.test(richdf$samplerichness ~ richdf$Rookery)
```

# Jaccard distance  
Jaccard distance is a presence/absence based distance matrix
```{r}
jacc.mar.dist <- distance(ps.mar.pa,method='jaccard')
jacc.bold.dist <- distance(ps.bold.pa,method='jaccard')
jacc.nt.dist <- distance(ps.nt.pa,method='jaccard')
```

### distance-based Redundancy analysis   
We want to know how much variation in community composition can be explained by the environmental variables we measured.  
To do that, we are going to use a distance-based redundancy analysis (db-RDA) to identify the significant variables. (Different from a regular RDA because it can use non-Euclidean dissimilarity indices - if you use a Euclidean index, you will get the same result as rda() ).   

```{r}
#using vegan 
rda.jacc.mar <- dbrda(jacc.mar.dist ~ RookeryBinary+Site+Month+Side,samplelist)
rda.jacc.bold <- dbrda(jacc.bold.dist ~ RookeryBinary+Site+Month+Side,as.data.frame(as.matrix(sample_data(ps.bold.pa))))
rda.jacc.nt <- dbrda(jacc.nt.dist ~ RookeryBinary+Site+Month+Side,as.data.frame(as.matrix(sample_data(ps.nt.pa))))
```

Calculate significance  
```{r}
step.mar <- vegan::anova.cca(rda.jacc.mar,permutations=1000,by='terms') #anova.cca performs an ANOVA like permutation test for RDA to assess the significance of constraints  
step.bold <- vegan::anova.cca(rda.jacc.bold,permutations=1000,by='terms')
step.nt <- vegan::anova.cca(rda.jacc.nt,permutations=1000,by='terms')
```

```{r}
step.mar
```

```{r}
step.bold
```

```{r}
step.nt
```

#### Visualize  
```{r}
#bar plot of relative eigenvalues. This is the percentage variance explained by each axis
plot_scree(rda.jacc.mar)
plot_scree(rda.jacc.bold)
plot_scree(rda.jacc.nt)
```

```{r}
jacc.mar.pa.plot <- plot_ordination(ps.mar.pa,rda.jacc.mar,color="Rookery",shape="Month") + facet_wrap(~Site) + labs(title="distance-based RDA plot",subtitle="Jaccard distance",caption="tax assigned using MAR alias database")
jacc.mar.pa.plot
#ggsave("Images/dbrda.jacc.mar.fsh.png",width=7,height=5)

jacc.bold.pa.plot <- plot_ordination(ps.bold.pa,rda.jacc.bold,color="Rookery",shape="Side") + facet_wrap(~Site) + labs(title="distance-based RDA plot",subtitle="Jaccard distance",caption="tax assigned using BOLD v5 database")
jacc.bold.pa.plot
#ggsave("Images/dbrda.jacc.bold.fsh.png",width=7,height=5)

jacc.nt.pa.plot <- plot_ordination(ps.nt.pa,rda.jacc.nt,color="Rookery",shape="Side") + facet_wrap(~Site) + labs(title="distance-based RDA plot",subtitle="Jaccard distance",caption="tax assigned using core nt database")
jacc.nt.pa.plot
#ggsave("Images/dbrda.jacc.nt.fsh.png",width=7,height=5)
```
```{r}
patchwork::wrap_plots(jacc.mar.pa.plot,jacc.nt.pa.plot,jacc.bold.pa.plot,nrow=1,guides = "collect")
#ggsave("Images/dbrda.jacc.allfsh.png",width=12,height=6)
```

```{r}
arrow_map_rda = aes(xend = dbRDA1, 
                yend = dbRDA2,
                x = 0, y = 0, 
                shape = NULL, color = NULL,label=labels
                )
label_map_rda = aes(x = 1.5 * dbRDA1, 
                y = 1.5 * dbRDA2, 
                shape = NULL, color = NULL,label=labels
                )
arrowhead = arrow(length = unit(0.03, "npc"))
```
```{r}
# get scores from rda  
sigscrs_rda_mar <- scores(rda.jacc.mar,display="bp") %>% as.data.frame() %>% filter(., abs(dbRDA1)>0.05 & abs(dbRDA2)>0.05)
# add labels  
arrowdf_rda_mar <- data.frame(labels = rownames(sigscrs_rda_mar), sigscrs_rda_mar)

jacc.mar.pa.scrs.plot <- plot_ordination(ps.mar.pa,rda.jacc.mar,shape="Rookery") + 
  geom_point(aes(color=Site)) + 
  geom_segment(arrow_map_rda, data=arrowdf_rda_mar,arrow=arrowhead) + #add vector line  
  #geom_text(label_map_rda, size = 4, data = arrowdf_rda_mar, color = "black")  #add label 
    ggrepel::geom_text_repel(label_map_rda,data = arrowdf_rda_mar, 
                           cex = 4, direction = "both", 
                           segment.size = 0.25,color="black") + labs(title="distance-based RDA plot",subtitle="Jaccard distance",caption="tax assigned using MAR alias database")
#ggsave("Images/dbrda.jacc.mar.scrs.fsh.png",width=7,height=7)
```

```{r}
# get scores from rda  
sigscrs_rda_bold <- scores(rda.jacc.bold,display="bp") %>% as.data.frame() %>% filter(., abs(dbRDA1)>0.05 & abs(dbRDA2)>0.05)
# add labels  
arrowdf_rda_bold <- data.frame(labels = rownames(sigscrs_rda_bold), sigscrs_rda_bold)

jacc.bold.pa.scrs.plot <- plot_ordination(ps.bold.pa,rda.jacc.bold,shape="Rookery") + 
  geom_point(aes(color=Site)) + 
  geom_segment(arrow_map_rda, data=arrowdf_rda_bold,arrow=arrowhead) + #add vector line  
  #geom_text(label_map_rda, size = 4, data = arrowdf_rda_bold, color = "black")  #add label 
    ggrepel::geom_text_repel(label_map_rda,data = arrowdf_rda_bold, 
                           cex = 4, direction = "both", 
                           segment.size = 0.25,color="black") + labs(title="distance-based RDA plot",subtitle="Jaccard distance",caption="tax assigned using BOLD v5 database")
#ggsave("Images/dbrda.jacc.bold.scrs.fsh.png",width=7,height=7)
```

```{r}
# get scores from rda  
sigscrs_rda_nt <- scores(rda.jacc.nt,display="bp") %>% as.data.frame() %>% filter(., abs(dbRDA1)>0.05 & abs(dbRDA2)>0.05)
# add labels  
arrowdf_rda_nt <- data.frame(labels = rownames(sigscrs_rda_nt), sigscrs_rda_nt)

jacc.nt.pa.scrs.plot <- plot_ordination(ps.nt.pa,rda.jacc.nt) + 
  geom_point(aes(color=Site)) + 
  geom_segment(arrow_map_rda, data=arrowdf_rda_nt,arrow=arrowhead) + #add vector line  
  #geom_text(label_map_rda, size = 4, data = arrowdf_rda_nt, color = "black")  #add label 
    ggrepel::geom_text_repel(label_map_rda,data = arrowdf_rda_nt, 
                           cex = 4, direction = "both", 
                           segment.size = 0.25,color="black") + labs(title="distance-based RDA plot",subtitle="Jaccard distance",caption="tax assigned using core nt database")
#ggsave("Images/dbrda.jacc.nt.scrs.fsh.png",width=7,height=5)
```

```{r}
patchwork::wrap_plots(jacc.nt.pa.scrs.plot,jacc.mar.pa.scrs.plot,jacc.bold.pa.scrs.plot,nrow=1,guides="collect")
#ggsave("Images/dbrda.jacc.allfsh.scrs.png",width=12,height=6,bg="white")
```

### PERMANOVA  
Adonis checks if groups have different compositions. However, this can be impacted by within-group dispersion. If distance to centroids is unequal between groups, some tests are very sensitive to that and will identify them as significant.   

#### test assumptions  
MANOVA can be confounded by within-group differences
```{r}
mod.rook <- betadisper(jacc.mar.dist,samplelist$Rookery) #are the communities are homogenously dispersed

anova(mod.rook)  #perform an anova on the dispersions; if the ANOVA's p-value is not significant it means that group dispersions are homogenous
```
So any differences identified between rookery/non-rookery samples should not be because the two groups have different dispersions, I think.   

HOWEVER,  
The sites themselves are statistically significant in their dispersions.  
```{r}
mod <- betadisper(jacc.mar.dist,samplelist$Site) #are the communities are homogenously dispersed

anova(mod)  #perform an anova on the dispersions; if the ANOVA's p-value is not significant it means that group dispersions are homogenous

TukeyHSD(mod)  #A Tukey's test can be done to see if and which groups differ in relation to their variances
```

#### test strata  
Strata forces the permutations to occur within the samples inside each block  
Because that variable is included in the block, it shouldn't go into the equation  
```{r}
perm.OR <-how(nperm=999)
setBlocks(perm.OR) <- with(samplelist,Site)

vegan::adonis2(jacc.mar.dist~RookeryBinary+Month+Side+Replicate+FilterNum,samplelist,by='terms',permutations=perm.OR)
vegan::adonis2(jacc.mar.dist~RookeryBinary+Month+Side+Replicate+FilterNum,samplelist,by='terms',permutations=1000)
```

```{r}
perm.mar <-how(nperm=999)
setBlocks(perm.mar) <- with(samplelist,Site)

perm.bold <-how(nperm=999)
setBlocks(perm.bold) <- with(as.data.frame(as.matrix(sample_data(ps.bold.pa))),Site)

perm.nt <-how(nperm=999)
setBlocks(perm.nt) <- with(as.data.frame(as.matrix(sample_data(ps.nt.pa))),Site)
```

```{r}
adonis.jacc.mar <- vegan::adonis2(jacc.mar.dist~RookeryBinary+Month+Side,samplelist,perm=perm.mar,by='terms')

adonis.jacc.bold <- vegan::adonis2(jacc.bold.dist~RookeryBinary+Month+Side,as.data.frame(as.matrix(sample_data(ps.bold.pa))),perm=perm.bold,by='terms')

adonis.jacc.nt <- vegan::adonis2(jacc.nt.dist~RookeryBinary+Month+Side,as.data.frame(as.matrix(sample_data(ps.nt.pa))),perm=perm.nt,by='terms')
```

```{r}
pairwiseAdonis::pairwise.adonis2(jacc.mar.dist~RookeryBinary*Month+Side,samplelist,strata="Site")
```


```{r}
adonis.jacc.mar
adonis.jacc.bold
adonis.jacc.nt
```

## Bray-Curtis dissimilarity  
This takes abundance into account  
```{r}
bray.mar.dist <- distance(ps.mar.rel,method='bray')
bray.bold.dist <- distance(ps.bold.rel,method='bray')
bray.nt.dist <- distance(ps.nt.rel,method='bray')
```

#### test assumptions  
MANOVA can be confounded by within-group differences
```{r}
mod.rook <- betadisper(bray.mar.dist,samplelist$Rookery) #are the communities are homogenously dispersed

anova(mod.rook)  #perform an anova on the dispersions; if the ANOVA's p-value is not significant it means that group dispersions are homogenous
```
```{r}
mod <- betadisper(bray.mar.dist,samplelist$Site)
anova(mod)
```
Looks like dispersions for relative abundance dataset are fine!  

### PCoA  
PCoA is commonly used for (dis)similarities. It is more flexible than a PCA plot, because it will preserve the distances generated by the similarity measure. However, it can only represent the Euclidian distances of a matrix, even if the similarity index you use reports non-euclidean distances or a mix.  

A good PCoA plot will generate only a few axes with large eigenvalues, capturing >50% of the variation  
```{r}
pcoa.bray.mar <- ordinate(ps.mar.rel,bray.mar.dist,method="PCoA")
pcoa.bray.bold <- ordinate(ps.bold.rel,bray.bold.dist,method="PCoA")
pcoa.bray.nt <- ordinate(ps.nt.rel,bray.nt.dist,method="PCoA")
```

```{r}
plot_scree(pcoa.bray.mar)
plot_scree(pcoa.bray.bold)
plot_scree(pcoa.bray.nt)
```

```{r}
plot_ordination(ps.mar.rel,pcoa.bray.mar,type="samples") + geom_point(aes(color=Rookery))
plot_ordination(ps.bold.rel,pcoa.bray.bold,type="samples") + geom_point(aes(color=Rookery))
plot_ordination(ps.nt.rel,pcoa.bray.nt,type="samples") + geom_point(aes(color=Rookery))
```

# GLM 
Want to know if any specific fish species has a response to rookery presence  
Want to know if there is a difference between trophic composition  
poisson distribution because we essentially have count data where the response variable = number of events occurring in a specific timeframe  
https://medium.com/@sahin.samia/a-comprehensive-introduction-to-generalized-linear-models-fd773d460c1d

Trophic index ~ Rookery + Month + Side + Site 
```{r}
lm1 <- glm(Abundance ~ Trophic.Round:Month + Rookery + Month + Site, data=ps.mar.rel.melt)
summary(lm1)
```

Does trophic diversity differ between site types?  
```{r}
lm2 <- lmer(Abundance ~ Trophic.Round:Rookery:Month + (1|Site) , data = ps.mar.pa.melt)
summary(lm2)
```

# Heatmap  
```{r}
mean_cutoff = 0.01
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

melt.mar.rel <- ps.mar.rel %>% tax_glom(.,"Genus",NArm=FALSE) %>% filter_taxa(mean_test, prune=TRUE) %>% psmelt()

melt.mar.rel %>% group_by(Month,Site,Rookery,Family,Genus) %>% summarize(Abund=sum(Abundance)) %>%  ggplot(aes(x=Rookery,y=Genus,fill=Abund)) + geom_tile(,group="Family") + facet_wrap(~Site,scales="free") + theme(axis.text.x = element_text(angle=90))
```
```{r}
melt.mar.rel
```


```{r}
gen.melt.mar.rel <- melt.mar.rel %>% ungroup() %>% mutate(Names=word(NewName,sep="R")) %>% group_by(Genus,Names,Rookery,Site) %>% summarize(Abund=sum(Abundance)) %>% ungroup() %>% group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>% mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund)) %>% group_by(Site) %>% arrange(.,Rookery,.by_group=TRUE) %>% ungroup() %>% select(-c(Rookery,Site))

gen.rel.tab <- gen.melt.mar.rel %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Names,values_from=RelAbund) %>% as.data.frame()
rownames(gen.rel.tab) <- gen.rel.tab$Genus
gen.rel.tab$Genus <- NULL
gen.rel.dt <- melt.mar.rel %>% ungroup() %>% mutate(Names=word(NewName,sep="R")) %>% group_by(.,Names) %>% summarize(Month=unique(Month),Site=unique(Site),Rookery=unique(Rookery))


 
 
 

annotation_col = gen.rel.dt %>% group_by(Rookery,Site) %>% 
  arrange("Rookery",.by_group=TRUE) %>% ungroup() %>% 
  select(c(Names,Rookery,Site)) %>% 
  mutate(Rookery=factor(Rookery,ordered=TRUE),
         Site=factor(Site,ordered=TRUE)) %>% 
  data.frame(.,row.names = "Names")
```

```{r}
pheatmap::pheatmap(gen.rel.tab,cluster_cols=FALSE,annotation_col=annotation_col,gaps_col=c(8,16))
```


# Indicator species   
```{r}
#https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html#data-required-for-indicator-species-analysis  
library(indicspecies) #needs a matrix or data frame with sites as rows and species in columns  
```
```{r}
perm.indic <-how(nperm=999)
setBlocks(perm.indic) <- with(gen.rel.dt,Site)
```

MAR fish species over 1% of db, relative abundance  
```{r}
indic.1pct.melt <- ps.mar.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  filter_taxa(mean_test, prune=TRUE) %>% 
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.1pct.tab <- indic.1pct.melt %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Genus,values_from=RelAbund) %>% as.data.frame() 
rownames(indic.1pct.tab) <- indic.1pct.tab$Names
indic.1pct.tab$Names <- NULL
```

```{r}
indval.1pct <- multipatt(indic.1pct.tab, gen.rel.dt$Rookery, 
                    control = perm.indic) 
summary(indval.1pct)
```

MAR fish species, relative abundance  
```{r}
indic.melt <- ps.mar.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.tab <- indic.melt %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Genus,values_from=RelAbund) %>% as.data.frame() 
rownames(indic.tab) <- indic.tab$Names
indic.tab$Names <- NULL
```

```{r}
indval <- multipatt(indic.tab, gen.rel.dt$Rookery, 
                    control = perm.indic) 
summary(indval)
```

BOLD, relative abundance  
```{r}
indic.bold.melt <- ps.bold.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.bold.tab <- indic.bold.melt %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Genus,values_from=RelAbund) %>% as.data.frame() 
rownames(indic.bold.tab) <- indic.bold.tab$Names
indic.bold.tab$Names <- NULL
```

```{r}
indval.bold <- multipatt(indic.bold.tab, gen.rel.dt$Rookery, 
                    control = perm.indic) 
summary(indval.bold)
```

core nt fish species, relative abundance  
```{r}
indic.nt.melt <- ps.nt.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.nt.tab <- indic.nt.melt %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Genus,values_from=RelAbund) %>% as.data.frame() 
rownames(indic.nt.tab) <- indic.nt.tab$Names
indic.nt.tab$Names <- NULL
```

```{r}
indval.nt <- multipatt(indic.nt.tab, gen.rel.dt$Rookery, 
                    control = perm.indic) 
summary(indval.nt)
```

# Trophic Level  
Violin plot
```{r,warning=FALSE}
ps.nt.rel.melt %>% 
  mutate(Trophic.Rounded = as.numeric(Trophic.Rounded),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Rounded,
           fill=Site)) + 
  facet_wrap(~Site,nrow=3) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Species' trophic index at rookery and non-rookery mangrove cayes",
       caption="taxa assigned using MAR fish species BLAST alias",
       y="Trophic Index")
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```
```{r,warning=FALSE}
ps.mar.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Site)) + 
  facet_wrap(~Site,nrow=3) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Species' trophic index at rookery and non-rookery mangrove cayes",
       caption="taxa assigned using MAR fish species BLAST alias",
       y="Trophic Index")
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```

```{r,warning=FALSE}
ps.mar.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Rookery)) + 
  facet_wrap(Rookery~Site,nrow=2) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Species' trophic index at rookery and non-rookery mangrove cayes",
       caption="taxa assigned using MAR fish species BLAST alias",
       y="Trophic Index")
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```

```{r,warning=FALSE}
ps.nt.rel.melt %>%
  mutate(Trophic.Level = as.numeric(Trophic.Level),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Level,
           fill=Rookery)) + 
  facet_wrap(Rookery~Site,nrow=2) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Species' trophic index at rookery and non-rookery mangrove cayes",
       caption="taxa assigned using nt BLAST database",
       y="Trophic Index")
#ggsave("Images/violin.relabund.trophic.nt.fsh.png",width=10,height=7)
```

(need to figure out how to add index to BOLD assigned taxa)  

```{r}
lmdata.mar <- ps.mar.rel.melt
lmdata.mar$Trophic.Round <- as.factor(lmdata.mar$Trophic.Round)
lmdata.mar$Site <- as.factor(lmdata.mar$Site)
```

```{r}
lm3 <- glm(Abundance ~ Trophic.Round:Month:Site + Trophic.Round:Rookery + Rookery, data=ps.mar.rel.melt)
summary(lm3)
```

```{r}
lme1 <- lmer(Abundance ~ Trophic.Round:Rookery + (1|Site) + (1|Month), data = ps.mar.pa.melt, REML=FALSE)
summary(lme1)
```
plot random effects  
(random effects with a variance of 0 have covariance)
```{r}
dotplot(ranef(lme1))
```

```{r}
lme0 <- lmer(Abundance ~ Trophic.Round  + (1|Site) + (1|Month), data = ps.mar.pa.melt, REML=FALSE)
summary(lme0)
```

Test 
```{r}
anova(lme1,lme0)
```

Test effect on residuals  
```{r}
kruskal.test(residuals(lme1) ~ ps.mar.pa.melt$Sample)
```

```{r}
ggeffects::ggpredict(lme1, terms = c("Trophic.Round","Rookery"),type="fixed") %>% plot()
```


Pie chart?  
```{r}
ps.mar.pa.melt %>%  
  group_by(Trophic.Round) %>% 
  #group_by(Site,Rookery,Trophic.Round) %>% #filter(!is.na(Trophic.Round)) %>%
  summarize(Tlevel=unique(Trophic.Round),Abundance=sum(Abundance)) %>% 
  mutate(Rel=Abundance/sum(Abundance)) %>% 
  mutate(csum = rev(cumsum(rev(Rel))), 
         pos = Rel/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Rel/2, pos)) %>% 
  ggplot(aes(x="",y=Rel,fill=Tlevel)) + 
  #ggplot(aes(x=Rookery,y=Rel,fill=Tlevel)) + 
  geom_bar(stat = "identity",position="stack",color="black") + 
  #facet_wrap(~Site) + 
  #facet_wrap(~Rookery) +
  labs(y="Relative abundance",title="Trophic index of species identified at rookery and non-rookery sites",caption="taxa annotated using core nt database") +
  coord_polar(theta="y",start=0) + theme_void() 
```

```{r}
ps.nt.pa.melt %>%  
  group_by(Trophic.Rounded) %>% 
  #group_by(Site,Rookery,Trophic.Rounded) %>% #filter(!is.na(Trophic.Rounded)) %>%
  summarize(Tlevel=unique(Trophic.Rounded),Abundance=sum(Abundance)) %>% 
  mutate(Rel=Abundance/sum(Abundance)) %>% 
  mutate(csum = rev(cumsum(rev(Rel))), 
         pos = Rel/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Rel/2, pos)) %>% 
  ggplot(aes(x="",y=Rel,fill=Tlevel)) + 
  #ggplot(aes(x=Rookery,y=Rel,fill=Tlevel)) + 
  geom_bar(stat = "identity",position="stack",color="black") + 
  #facet_wrap(~Site) + 
  #facet_wrap(~Rookery) +
  labs(y="Relative abundance",title="Trophic index category of species identified at rookery and non-rookery sites",caption="taxa annotated using core nt database") +
  coord_polar(theta="y",start=0) + theme_void() 
```

# IUCN  
```{r}
#ps.mar.rel.melt %>% 
ps.mar.pa %>% psmelt() %>%
  group_by(Site,Rookery,IUCN.Category) %>%
  summarize(IUCN=unique(IUCN.Category),Abundance=sum(Abundance)) %>% 
  mutate(Rel=Abundance/sum(Abundance)) %>% ggplot(aes(x=Rookery,y=Rel,fill=IUCN)) + geom_bar(stat = "identity",position="stack") + facet_wrap(~Site) + labs(y="Relative abundance",title="IUCN category of species identified at rookery and non-rookery sites",caption="taxa annotated using MAR nt alias database")
#ggsave("Images/iucn.mar.png")
```

```{r}
ps.mar.pa.melt$IUCN.Category[ps.mar.pa.melt$IUCN.Category==""] <- NA 
ps.mar.pa.melt %>% 
  group_by(IUCN.Category) %>% 
  #group_by(Rookery,IUCN.Category) %>% 
  #group_by(Site,Rookery,IUCN.Category) %>% #filter(!is.na(IUCN.Category)) %>%
  summarize(IUCN=unique(IUCN.Category),Abundance=sum(Abundance)) %>% 
  mutate(Rel=Abundance/sum(Abundance)) %>% 
  mutate(csum = rev(cumsum(rev(Rel))), 
         pos = Rel/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Rel/2, pos)) %>% 
  ggplot(aes(x="",y=Rel,fill=IUCN)) + 
  #ggplot(aes(x=Rookery,y=Rel,fill=IUCN)) + 
  geom_bar(stat = "identity",position="stack",color="black") + 
  #facet_wrap(~Site) + 
  #facet_wrap(~Rookery) +
  labs(y="Relative abundance",title="IUCN category of species identified at rookery and non-rookery sites",caption="taxa annotated using MAR species BLAST alias database") +
  coord_polar(theta="y",start=0) + theme_void()  
```

And why are we using the presence/absence database?  
Because using the relative abundance one massively skews the data because of one sample from Hicks Caye that has a huge representation from shark   
```{r}
ps.mar.rel.melt %>%
  mutate(Month=factor(Month,levels=c("August","November","February","May")))  %>% filter(IUCN.Category=="Endangered") %>% ggplot(aes(x=SampleName_Short,y=Abundance,fill=Family)) + geom_bar(stat="identity",position="stack") + facet_grid(Month~Site,scales="free_x") + theme(axis.text.x = element_text(angle=90))
```

```{r}
ps.mar.rel.melt %>%
    mutate(Month=factor(Month,levels=c("August","November","February","May")))  %>% filter(IUCN.Category=="Endangered" & Site=="Hicks Caye" & Month=="August" & Abundance>0) 
```

```{r}
 ps.nt.pa.melt %>%  
# ps.nt.rel.melt %>% 
  group_by(Site,Rookery,IUCN.Category) %>% #filter(!is.na(IUCN.Category)) %>%
  summarize(IUCN=unique(IUCN.Category),Abundance=sum(Abundance)) %>% 
  mutate(Rel=Abundance/sum(Abundance)) %>% ggplot(aes(x=Rookery,y=Rel,fill=IUCN)) + 
  geom_bar(stat = "identity",position="stack") + 
  facet_wrap(~Site) + 
  labs(y="Relative abundance",title="IUCN category of species identified at rookery and non-rookery sites",caption="taxa annotated using core nt database")
#ggsave("Images/iucn.nt.fsh.png")
```
```{r}
 ps.nt.pa.melt %>% 
  group_by(IUCN.Category) %>% 
  #group_by(Rookery,IUCN.Category) %>% 
  #group_by(Site,Rookery,IUCN.Category) %>% #filter(!is.na(IUCN.Category)) %>%
  summarize(IUCN=unique(IUCN.Category),Abundance=sum(Abundance)) %>% 
  mutate(Rel=Abundance/sum(Abundance)) %>% 
  mutate(csum = rev(cumsum(rev(Rel))), 
         pos = Rel/2 + lead(csum, 1),
         pos = if_else(is.na(pos), Rel/2, pos)) %>% 
  ggplot(aes(x="",y=Rel,fill=IUCN)) + 
  #ggplot(aes(x=Rookery,y=Rel,fill=IUCN)) + 
  geom_bar(stat = "identity",position="stack",color="black") + 
  #facet_wrap(~Site) + 
  #facet_wrap(~Rookery) +
  labs(y="Relative abundance",title="IUCN category of species identified at rookery and non-rookery sites",caption="taxa annotated using core nt database") +
  coord_polar(theta="y",start=0) + theme_void() + 
  geom_label_repel(aes(y = pos, label = IUCN.Category),
                   size = 3, nudge_x = 1, show.legend = FALSE,direction="both") 
```

# Session Info   
```{r}
save.image("RData/08_Analysis_Fish.RData")
```

```{r}
sessionInfo()
```
