---
output:
  word_document: default
  html_document: default
---

# Load  
Load libraries  
```{r,message=FALSE}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(vegan)
library(lme4)
```

Load data  
```{r load}
dir_data <- '/Users/lauragivens/Desktop/R/BZrookery_eDNA/Rdata'
dir_man <- "/Volumes/Fuji/Mangroves"
dir_results <- "/Volumes/Fuji/Mangroves/2025_0319_Givens_Canty_Rookery_COI/cutadapt/results"
#dir_results <- "/Users/lauragivens/Desktop/MangroveResults"

ps.og <- readRDS(paste0(dir_results,
                        "/prune.decontam.combined.wordsize.ps.rds"))

ps.wbla <- readRDS(paste0(dir_results,
                        "/prune.decontam.combined.wordsize.ps.keepblanks.rds"))

#taxa_nt_troph <- readRDS(paste0(dir_results,"/taxtable.v2.nt.wtroph.rds"))
#tax.combined.df <- readRDS(paste0(dir_results,"/tax.combined.df.rds"))

taxa.traits.df <- readRDS(paste0(dir_results,"/tax.troph.combined.wordsize.df.rds"))
taxa.taxize.df <- readRDS(paste0(dir_results,"/taxa_traits_df.rds"))
```

Or alternatively,  
```{r,eval=FALSE}
load("RData/08_Analysis.RData")
```

```{r}
ggplot2::theme_set(theme_cowplot() + theme(
  legend.position = "bottom",
  panel.border = element_rect(color='black'),
  legend.box.margin = margin(5,5,5,5),
  legend.spacing = unit(0,"pt"),
  plot.title = element_text(face="bold"),
  plot.caption = element_text(face="italic"),
  text = element_text(size=(15),family="Times"),
  legend.key.height = unit(10,"pt"),
  legend.key.width = unit(10,"pt")
))
```

# Update ps  
```{r}
samplelist <- as.data.frame(as.matrix(sample_data(ps.og)))
```

```{r}
ps <- ps.og  %>% 
  prune_samples(sample_sums(.)>50,.) %>% 
  subset_samples(Rookery!="Practice") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% 
  prune_samples(sample_sums(.)>0,.)
```

Convert rookery to 0/1  
```{r}
samplelist <- as.data.frame(as.matrix(sample_data(ps)))
samplelist <- samplelist %>% 
  mutate(RookeryBinary = case_when(Rookery=="Rookery" ~ 1,
                                   Rookery!="Rookery" ~ 0),
         .after=Rookery )
samplelist <- samplelist %>% mutate(NewName=word(rownames(.),sep="_",end=2),
                                    Trip=factor(Trip,levels=c("Trip 1: Summer 2023","Trip 2: Fall 2023","Trip 3: Winter 2024","Trip 4: Spring 2024")))

sample_data(ps) <- samplelist
```

Update tax list  
```{r}
#tax <- as.data.frame(as.matrix(tax_table(ps)))
tax <- taxa.taxize.df %>%
  mutate(Class=case_when(Class=="Actinopterygii" ~ "Actinopteri",
                         .default=Class))
rownames(tax) <- tax$seqid
tax$seqid <- NULL 

tax_table(ps) <- tax_table(as.matrix(tax))
```

```{r}
tab <- otu_table(ps) %>% as.matrix() %>% as.data.frame() %>% t()
```

# Dataset manipulations  

## Whole dataset  
keep taxa seen at least twice in 2 or more samples  
```{r} 
ps.prune <- filter_taxa(ps, function(x) sum(x >= 2) >= 2, TRUE)
```

Dataset manipulations  
```{r}
ps.prune.pa <- ps.prune %>% 
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.prune.rel <- ps.prune %>% transform_sample_counts(.,function(x)x/sum(x))
ps.rel.melt <- psmelt(ps.prune.rel) %>% mutate(Diettroph=as.double(Diettroph))

tab.prune <- otu_table(ps.prune) %>%
  as.matrix() %>% as.data.frame() %>% t()
```

```{r,eval=FALSE}
ps.pa <- ps %>% 
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.rel <- ps %>% 
  transform_sample_counts(.,function(x) x/sum(x))
```

## Chordate dataframes   
```{r}
ps.ch <-  ps %>% 
  subset_taxa(.,Phylum=="Chordata") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>%
  prune_samples(sample_sums(.)>0,.)
#get_taxa_unique(ps.ch,"Class")

ps.ch.pa <-  ps.ch %>% 
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.ch.rel <-  ps.ch %>% 
  transform_sample_counts(.,function(x) x/sum(x))

tab.ch <- otu_table(ps.ch) %>%
  as.matrix() %>% as.data.frame() %>% t()

samplelist.ch <- as.data.frame(as.matrix(sample_data(ps.ch)))
```

```{r}
ps.othr <- ps.ch %>% 
  subset_taxa(Class!="Actinopteri" & Class != "Ascidiacea" & Class!="Elasmobranchii") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>%
  prune_samples(sample_sums(.)>0,.)

ps.othr.pa <- ps.othr %>% 
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.othr.rel <- ps.othr %>% 
  transform_sample_counts(.,function(x) x/sum(x))

tab.othr <- otu_table(ps.othr) %>% 
  as.matrix() %>% as.data.frame() %>% t()

samplelist.othr <- as.data.frame(as.matrix(sample_data(ps.othr)))
```

## Fish dataframes  
```{r}
ps.fsh <- ps.ch %>% 
  subset_taxa(Class=="Actinopteri" | Class=="Elasmobranchii") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>%
  prune_samples(sample_sums(.)>0,.)

ps.fsh.pa <- ps.fsh %>% 
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.fsh.rel <- ps.fsh %>% 
  transform_sample_counts(.,function(x) x/sum(x))

ps.fsh.rel.melt <- ps.fsh.rel %>% psmelt() %>% mutate(Diettroph=as.double(Diettroph))

tab.fsh <- otu_table(ps.fsh.rel) %>%
  as.matrix() %>% as.data.frame() %>% t()

samplelist.fsh <- as.data.frame(as.matrix(sample_data(ps.fsh)))
```

## Bird dataframes  
```{r}
ps.brd <- ps.ch %>% 
  subset_taxa(Class=="Aves") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>%
  prune_samples(sample_sums(.)>0,.)

ps.brd.pa <- ps.brd %>% 
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.brd.rel <- ps.brd %>% 
  transform_sample_counts(.,function(x) x/sum(x))

tab.brd <- otu_table(ps.brd) %>%
  as.matrix() %>% as.data.frame() %>% t()

samplelist.brd <- as.data.frame(as.matrix(sample_data(ps.brd)))
```

Compare  
```{r}
ps
ps.prune

ps.ch
ps.fsh
ps.othr
ps.brd
```

# Summary  
```{r}
ps
```
```{r}
ps.bla <- ps.wbla %>% subset_samples(Rookery=="Blank") %>% prune_taxa(taxa_sums(.)>0,.)

ps.bla
```
```{r}
range(sample_sums(ps.bla))
```

# Main questions  
1. does rookery presence have an impact on fish community composition?  
2. does rookery presence have an impact on whole community composition?  
3. What species are impacted?  (Commercial/non-commercial; trophic index as well)
4. how does distance from shore and/or distance from populations impact community composition?  
5. how does size of protected area correlate?  

# 3. Analysis (Community)   

## Richness  
No impact of side when looked at by date and site
```{r rich-samplename,warning=FALSE}
#plot_richness(ps,x="SampleName",measures = c("Observed","Shannon"),color="Rookery") + geom_boxplot()

sppr <- specnumber(tab)

rich.df <- full_join((samplelist %>% rownames_to_column("names")),
                     (data.frame(names= names(sppr),spp=sppr))
                     )

side.comps <- list(c("15-Aug_DCNRL","15-Aug_DCNRW"),
                    c("15-Aug_DCRL","15-Aug_DCRW" ),
                    c("15-Aug_HINRL","15-Aug_HINRW"),
                    c("15-Aug_HIRL","15-Aug_HIRW" ),
                    c("17-Aug_TANRL","17-Aug_TANRW"),
                    c("17-Aug_TARL","17-Aug_TARW" ),
                    c("22-Feb_DCNRL","22-Feb_DCNRW"),
                    c("22-Feb_DCRL","22-Feb_DCRW"),
                    c("22-Feb_TANRL","22-Feb_TANRW"),
                    c("22-Feb_TARL","22-Feb_TARW" ),
                   c("23-Feb_HINRL","23-Feb_HINRW" ),
                   c("23-Feb_HIRL","23-Feb_HIRW"),
                   c("23-May_DCNRL","23-May_DCNRW"),
                   c("23-May_DCRL","23-May_DCRW"),
                   c("23-May_HINRL","23-May_HINRW"),
                   c("23-May_HIRL","23-May_HIRW" ),
                   c("23-May_TANRL","23-May_TANRW"),
                   c("23-May_TARL","23-May_TARW" ),
                   c("23-Nov_DCNRL","23-Nov_DCNRW"),
                   c("23-Nov_DCRL","23-Nov_DCRW"),
                   c("23-Nov_HINRL","23-Nov_HINRW"),
                   c("23-Nov_HIRL","23-Nov_HIRW" ),
                   c("23-Nov_TANRL","23-Nov_TANRW"),
                   c("23-Nov_TARL","23-Nov_TARW" )
                    )


ggpubr::ggboxplot(rich.df,x="NewName",y="spp",
          color="Rookery",add="jitter") +
  #facet_wrap(~Trip,scales="free") +
  ggpubr::stat_compare_means(comparisons=side.comps,
                    # symnum.args = list(
                    #   cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
                    #   symbols = c("****", "***", "**", "*", "ns")),
                     #hide.ns=TRUE
                     ) + theme(axis.text.x = element_text(angle=90))
```

Significant impact of side at Turneffe
```{r}
ggpubr::ggboxplot(rich.df,x="SampleName_Short",y="spp",
          color="Rookery",add="jitter") +
  #facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(comparisons=list(c("DCNRL","DCNRW"),
                                              c("DCRL","DCRW"),
                                              c("HINRL","HINRW"),
                                              c("HIRL","HIRW"),
                                              c("TANRL","TANRW"),
                                              c("TARL","TARW")
                                              ),
                    # symnum.args = list(
                    #   cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
                    #   symbols = c("****", "***", "**", "*", "ns")),
                     #hide.ns=TRUE
                     ) + 
  theme(axis.text.x = element_text(angle=90))
#ggsave("Figures/richness.boxplot.side.png",width=7,height=5)
```

```{r}
ggpubr::ggboxplot(rich.df,x="Rookery",y="spp",
          color="Site",add="jitter") +
  facet_wrap(~Trip,scales="free") +
  ggpubr::stat_compare_means(comparisons=list(c("Rookery","Non-rookery")),
                    # symnum.args = list(
                    #   cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
                    #   symbols = c("****", "***", "**", "*", "ns")),
                     #hide.ns=TRUE
                     ) + theme(axis.text.x = element_text(angle=90))

ggpubr::ggboxplot(rich.df,,x="Trip",y="spp",
          color="Rookery",add="jitter") +
  facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(
    aes(group=Rookery),
    #comparisons=list(c("Rookery","Non-rookery")),
    method="wilcox.test",
    label="p.signif",
    symnum.args = list(
      cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
      symbols = c("p < 0.0001", "p < 0.001", "p < 0.01", "p < 0.05", "ns")),
    vjust=0,
    hide.ns=TRUE
    ) + theme(axis.text.x = element_text(angle=45,hjust=1))

#ggsave("Figures/richness.boxplot.rookery.png",width=7,height = 10)
```

## bray dissimilarity
```{r}
bray.dist <- distance(ps.prune.rel,method='bray')
```

### distance-based Redundancy analysis   
We want to know how much variation in community composition can be explained by the environmental variables we measured.  
To do that, we are going to use a distance-based redundancy analysis (db-RDA) to identify the significant variables. (Different from a regular RDA because it can use non-Euclidean dissimilarity indices - if you use a Euclidean index, you will get the same result as rda() ).   

Make base rda  
```{r}
rda.0 <- dbrda(bray.dist~1, data=samplelist)
```

Add environmental variables  
```{r}
rda.bray <- dbrda(bray.dist ~ RookeryBinary + Site + Month + Side, samplelist)

perm <-how(nperm=999)
setBlocks(perm) <- with(samplelist,Site)
```

Calculate significance  
```{r}
step <- vegan::anova.cca(rda.bray,permutations=perm,by='terms')
```

```{r}
step
```

#### Visualize  
```{r}
plot_ordination(ps.prune.rel,rda.bray,
                color="Rookery") + 
     #          color="Rookery") + 
  #facet_wrap(~Month) + 
  #facet_wrap(~Site) +
  facet_grid(Site~Month) +
  stat_ellipse()
#ggsave("Figures/rda.site.month.rookery.png",width=7,height=5)
```

```{r}
plot_ordination(ps.prune.rel,rda.bray,
                color="Site",shape="Site") + 
  facet_wrap(~Month) +
  stat_ellipse()
#ggsave("Figures/rda.site.month.png",width=7,height=5)
```

### PERMANOVA  
```{r}
adonis.bray <- vegan::adonis2(bray.dist~RookeryBinary+Month+Side,
                              samplelist,
                              perm=perm,by='terms')
adonis.bray
```

### Trophic index  
```{r}
# if scale="area" (default), all violins have the same area (before trimming the tails). If "count", areas are scaled proportionally to the number of observations. If "width", all violins have the same maximum width.

ps.rel.melt %>% filter(Abundance > 0) %>%
  ggplot(.,aes(x=Rookery,y=Diettroph,
               #color=Side
               )) + 
  facet_grid(Trip~Site) +
    geom_violin(na.rm=TRUE,
                adjust=0.5,
                #draw_quantiles = c(0.25,0.75),
                #scale="width",
                scale="count" # Scale maximum width proportional to sample size: 
                ) + geom_jitter(height = 0, width = 0.05,alpha=0.5) + 
  geom_boxplot(width=0.1) + theme_bw()

#ggsave("Figures/tlevel.trip.site.side.png",width=10,height=7)
#ggsave("Figures/tlevel.trip.site.png",width=10,height=7)
```

# 2. Analysis (Fish)  

## Richness  
```{r fshrich-samplename,warning=FALSE}
#plot_richness(ps,x="SampleName",measures = c("Observed","Shannon"),color="Rookery") + geom_boxplot()

fsh.sppr <- specnumber(tab.fsh)

rich.df.fsh <- full_join((samplelist.fsh %>% rownames_to_column("names")),(data.frame(names= names(fsh.sppr),spp=fsh.sppr))) %>% mutate(Trip=factor(Trip,levels=c("Trip 1: Summer 2023","Trip 2: Fall 2023","Trip 3: Winter 2024","Trip 4: Spring 2024")))
```

```{r}
ggpubr::ggboxplot(rich.df.fsh,x="Trip",y="spp",
          #color="Rookery",
          add="jitter") +
  facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(
    aes(group=Rookery),
    #comparisons=list(c("Rookery","Non-rookery")),
    method="wilcox.test",
    label="p.signif",
    symnum.args = list(
      cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
      symbols = c("p < 0.0001", "p < 0.001", "p < 0.01", "p < 0.05", "ns")),
    vjust=0,
    hide.ns=TRUE
    ) + 
  theme(axis.text.x = element_text(angle=45,hjust=1))
#ggsave("Figures/fsh.richness.boxplot.site.png",width=7,height=5)
```

```{r}
ggpubr::ggboxplot(rich.df.fsh,x="SampleName_Short",y="spp",
          color="Rookery",add="jitter") +
  #facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(comparisons=list(c("DCNRL","DCNRW"),
                                              c("DCRL","DCRW"),
                                              c("HINRL","HINRW"),
                                              c("HIRL","HIRW"),
                                              c("TANRL","TANRW"),
                                              c("TARL","TARW")
                                              ),
                    # symnum.args = list(
                    #   cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
                    #   symbols = c("****", "***", "**", "*", "ns")),
                     #hide.ns=TRUE
                     ) + 
  theme(axis.text.x = element_text(angle=90))
#ggsave("Figures/fsh.richness.boxplot.side.png",width=7,height=5)
```

## bray dissimilarity    
```{r}
bray.fsh.dist <- distance(ps.fsh.rel,method='bray')
```

### distance-based Redundancy analysis   
We want to know how much variation in community composition can be explained by the environmental variables we measured.  
To do that, we are going to use a distance-based redundancy analysis (db-RDA) to identify the significant variables. (Different from a regular RDA because it can use non-Euclidean dissimilarity indices - if you use a Euclidean index, you will get the same result as rda() ).   

Make base rda  
```{r}
#using phyloseq
#rda.0 <- ordinate(ps.fsh.rel,distance=bray.dist)

#using vegan
rda.0.fsh <- dbrda(tab.fsh~1,data=samplelist.fsh)
```

Add environmental variables  
```{r}
#using phyloseq  
#rda.bray.fsh <- ordinate(ps.fsh.rel,method="RDA",distance=bray.fsh.dist,formula=Site+Month+Side+RookeryBinary)

#using vegan 
rda.bray.fsh <- dbrda(bray.fsh.dist ~ RookeryBinary + Site + Month + Side,
                      samplelist.fsh)
```

Calculate significance  
```{r}
perm.fsh <-how(nperm=999)
setBlocks(perm.fsh) <- with(samplelist.fsh,Site)

#step.fsh <- ordistep(rda.0.fsh,scope=formula(rda.bray.fsh),perm.max=1000)

 #anova.cca performs an ANOVA like permutation test for RDA to assess the significance of constraints  
step.fsh <- vegan::anova.cca(rda.bray.fsh,permutations=perm.fsh,by='terms')
```
```{r}
step.fsh
```

#### Visualize  
```{r}
plot_ordination(ps.fsh.rel,rda.bray.fsh,
                color="Site") +
  facet_grid(~Trip) + stat_ellipse()
#ggsave("Figures/fsh.rda.bray.site.png",width=7,height=5)

plot_ordination(ps.fsh.rel,rda.bray.fsh,
                color="Rookery") +
  facet_grid(Trip~Site) + 
  stat_ellipse()
#ggsave("Figures/fsh.rda.bray.rookery.site.png",width=7,height=5)
```

### Statistical Significance   
```{r}
adonis.bray.fsh <- vegan::adonis2(bray.fsh.dist~RookeryBinary+Month+Side,
                                  samplelist.fsh,
                                  perm=1000,by='terms',permutations=perm.fsh)
adonis.bray.fsh
```

Kruskal test - difference in abundance between rookery/non-rookery at each caye  
```{r}
ps.prune.rel.fsh.melt <- psmelt(ps.fsh.rel)
```

```{r}
kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.fsh.melt %>% filter(Site=="Hicks Caye"))
             )

kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.fsh.melt %>% filter(Site=="Turneffe Atoll"))
             )

kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.fsh.melt %>% filter(Site=="Drowned Caye"))
             )
```

# 3a. Analysis (non-fish chordates)    

## Richness  
```{r othrrich-samplename,warning=FALSE}
othr.sppr <- specnumber(tab.othr)

rich.df.othr <- full_join((samplelist.othr %>% rownames_to_column("names")),(data.frame(names= names(othr.sppr),spp=othr.sppr))) %>% mutate(Trip=factor(Trip,levels=c("Trip 1: Summer 2023","Trip 2: Fall 2023","Trip 3: Winter 2024","Trip 4: Spring 2024")))
```

```{r}
ggpubr::ggboxplot(rich.df.othr,x="Trip",y="spp",
          color="Rookery",add="jitter") +
  facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(
    aes(group=Rookery),
    #comparisons=list(c("Rookery","Non-rookery")),
    method="wilcox.test",
    label="p.signif",
    symnum.args = list(
      cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
      symbols = c("p < 0.0001", "p < 0.001", "p < 0.01", "p < 0.05", "ns")),
    vjust=0,
    hide.ns=TRUE
    ) + 
  theme(axis.text.x = element_text(angle=45,hjust=1))
#ggsave("Figures/othr.richness.boxplot.site.png",width=7,height=5)
```

```{r}
ggpubr::ggboxplot(rich.df.othr,x="SampleName_Short",y="spp",
          color="Rookery",add="jitter") +
  #facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(comparisons=list(c("DCNRL","DCNRW"),
                                              c("DCRL","DCRW"),
                                              c("HINRL","HINRW"),
                                              c("HIRL","HIRW"),
                                              c("TANRL","TANRW"),
                                              c("TARL","TARW")
                                              ),
                    # symnum.args = list(
                    #   cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
                    #   symbols = c("****", "***", "**", "*", "ns")),
                     #hide.ns=TRUE
                     ) + 
  theme(axis.text.x = element_text(angle=90))
#ggsave("Figures/othr.richness.boxplot.side.png",width=7,height=5)
```

## bray dissimilarity  
```{r}
bray.othr.dist <- distance(ps.othr.rel,method='bray')

rda.bray.othr <- dbrda(bray.othr.dist ~ RookeryBinary+Site+Month+Side,
                       samplelist.othr)

perm.othr <-how(nperm=999)
setBlocks(perm.othr) <- with(samplelist.othr,Site)

adonis.bray.othr <- vegan::adonis2(bray.othr.dist~RookeryBinary+Month+Side,
                                   samplelist.othr,
                                   permutations=perm.othr,by='terms')
adonis.bray.othr
```

### db-RDA  
Make base rda  
```{r}
#using phyloseq
#rda.0 <- ordinate(ps.othr.rel,distance=bray.dist)

#using vegan
rda.0.othr <- dbrda(tab.othr~1,data=samplelist.othr)
```

Add environmental variables  
```{r}
#using phyloseq  
#rda.bray.othr <- ordinate(ps.othr.rel,method="RDA",distance=bray.othr.dist,formula=Site+Month+Side+RookeryBinary)

#using vegan 
rda.bray.othr <- dbrda(bray.othr.dist ~ RookeryBinary + Site + Month + Side,
                      samplelist.othr)
```

Calculate significance  
```{r}
perm.othr <-how(nperm=999)
setBlocks(perm.othr) <- with(samplelist.othr,Site)

#step.othr <- ordistep(rda.0.othr,scope=formula(rda.bray.othr),perm.max=1000)

 #anova.cca performs an ANOVA like permutation test for RDA to assess the significance of constraints  
step.othr <- vegan::anova.cca(rda.bray.othr,permutations=perm.othr,by='terms')
```
```{r}
step.othr
```

### Visualize  
```{r}
plot_ordination(ps.othr.rel,rda.bray.othr,
                color="Site") +
  facet_grid(~Trip) + stat_ellipse()
#ggsave("Figures/othr.rda.bray.site.png",width=7,height=5)

plot_ordination(ps.othr.rel,rda.bray.othr,
                color="Rookery") +
  facet_grid(Trip~Site) + 
  stat_ellipse()
#ggsave("Figures/othr.rda.bray.rookery.site.png",width=7,height=5)
```

### Statistical Significance  
PERMANOVA  
```{r}
adonis.bray.othr <- vegan::adonis2(bray.othr.dist~RookeryBinary+Month+Side,
                                  samplelist.othr,
                                  perm=perm.othr,by='terms')
adonis.bray.othr
```

Kruskal test - difference in abundance between rookery/non-rookery at each caye  
```{r}
ps.prune.rel.othr.melt <- psmelt(ps.othr.rel)
```

```{r}
kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.othr.melt %>% filter(Site=="Hicks Caye"))
             )

kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.othr.melt %>% filter(Site=="Turneffe Atoll"))
             )

kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.othr.melt %>% filter(Site=="Drowned Caye"))
             )
```


# 3b. Birds specifically  

## Richness  
```{r brdrich-samplename,warning=FALSE}
brd.sppr <- specnumber(tab.brd)

rich.df.brd <- full_join((samplelist.brd %>% rownames_to_column("names")),(data.frame(names= names(brd.sppr),spp=brd.sppr))) %>% mutate(Trip=factor(Trip,levels=c("Trip 1: Summer 2023","Trip 2: Fall 2023","Trip 3: Winter 2024","Trip 4: Spring 2024")))
```

```{r}
ggpubr::ggboxplot(rich.df.brd,x="Trip",y="spp",
          color="Rookery",add="jitter") +
  facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(
    aes(group=Rookery),
    #comparisons=list(c("Rookery","Non-rookery")),
    method="wilcox.test",
    label="p.signif",
    symnum.args = list(
      cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
      symbols = c("p < 0.0001", "p < 0.001", "p < 0.01", "p < 0.05", "ns")),
    vjust=0,
    #hide.ns=TRUE
    ) + 
  theme(axis.text.x = element_text(angle=45,hjust=1))
#ggsave("Figures/brd.richness.boxplot.site.png",width=7,height=5)
```

```{r}
ggpubr::ggboxplot(rich.df.brd,x="SampleName_Short",y="spp",
          color="Rookery",add="jitter") +
  #facet_wrap(~Site,scales="free") +
  ggpubr::stat_compare_means(comparisons=list(c("DCNRL","DCNRW"),
                                              c("DCRL","DCRW"),
                                              c("HINRL","HINRW"),
                                              c("HIRL","HIRW"),
                                              c("TANRL","TANRW"),
                                              c("TARL","TARW")
                                              ),
                    # symnum.args = list(
                    #   cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf), 
                    #   symbols = c("****", "***", "**", "*", "ns")),
                     #hide.ns=TRUE
                     ) + 
  theme(axis.text.x = element_text(angle=90))
#ggsave("Figures/brd.richness.boxplot.side.png",width=7,height=5)
```

## bray dissimilarity  
```{r}
bray.brd.dist <- distance(ps.brd.rel,method='bray')
```

### db-RDA  
Make base rda  
```{r}
#using phyloseq
#rda.0 <- ordinate(ps.brd.rel,distance=bray.dist)

#using vegan
rda.0.brd <- dbrda(tab.brd~1,data=samplelist.brd)
```

Add environmental variables  
```{r}
#using phyloseq  
#rda.bray.brd <- ordinate(ps.brd.rel,method="RDA",distance=bray.brd.dist,formula=Site+Month+Side+RookeryBinary)

#using vegan 
rda.bray.brd <- dbrda(bray.brd.dist ~ RookeryBinary + Site + Month + Side,
                      samplelist.brd)
```

Calculate significance  
```{r}
perm.brd <-how(nperm=999)
setBlocks(perm.brd) <- with(samplelist.brd,Site)

#step.brd <- ordistep(rda.0.brd,scope=formula(rda.bray.brd),perm.max=1000)

 #anova.cca performs an ANOVA like permutation test for RDA to assess the significance of constraints  
step.brd <- vegan::anova.cca(rda.bray.brd,permutations=perm.brd,by='terms')
```
```{r}
step.brd
```

### Visualize  
```{r}
plot_ordination(ps.brd.rel,rda.bray.brd,
                color="Site") +
  facet_grid(~Trip) #+ stat_ellipse()
#ggsave("Figures/brd.rda.bray.site.png",width=7,height=5)

plot_ordination(ps.brd.rel,rda.bray.brd,
                color="Rookery") +
  facet_grid(Trip~Site) #+ stat_ellipse()
#ggsave("Figures/brd.rda.bray.rookery.site.png",width=7,height=5)
```

### Statistical Significance   
```{r}
adonis.bray.brd <- vegan::adonis2(bray.brd.dist~RookeryBinary+Month+Side,
                                  samplelist.brd,
                                  perm=perm.brd,by='terms')
adonis.bray.brd
```

Kruskal test - difference in abundance between rookery/non-rookery at each caye  
```{r}
ps.prune.rel.brd.melt <- psmelt(ps.brd.rel)
```

```{r}
kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.brd.melt %>% filter(Site=="Hicks Caye"))
             )

kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.brd.melt %>% filter(Site=="Turneffe Atoll"))
             )

kruskal.test(Abundance ~ Rookery,
             data=(ps.prune.rel.brd.melt %>% filter(Site=="Drowned Caye"))
             )
```

# Linear model   

## Whole community  
```{r}
rel.model1 <- lmer(Abundance ~ Diettroph*Rookery + (1|Site) + (1|Month), data = ps.rel.melt)
rel.model2 <- lmer(Abundance ~ Diettroph + (1|Site) + (1|Month), data = ps.rel.melt)
rel.model3 <- lmer(Abundance ~ Diettroph + (1|Month) , data = ps.rel.melt)

anova(rel.model1,rel.model2,rel.model3)

summary(rel.model1)
# help("pvalues",package="lme4")
car::Anova(rel.model1) #calculates type-II or tyle-III analysis of variance tables for model objects produced by..lmer
lmerTest::ranova(rel.model1) #tests for significance of random effects
```

## Fish  
Want to know if any specific fish species has a response to rookery presence  
Want to know if there is a difference between trophic composition  
poisson distribution because we essentially have count data where the response variable = number of events occurring in a specific timeframe  
https://medium.com/@sahin.samia/a-comprehensive-introduction-to-generalized-linear-models-fd773d460c1d

Does trophic diversity differ between site types?  
```{r}
rel.fsh.model1 <- lmer(Abundance ~ Rookery*Diettroph + (1|Site) + (1|Month), data = ps.fsh.rel.melt)
rel.fsh.model2 <- lmer(Abundance ~ Diettroph  + (1|Site) + (1|Month), data = ps.fsh.rel.melt)
rel.fsh.model3 <- lmer(Abundance ~ Diettroph + (1|Month) , data = ps.fsh.rel.melt)

anova(rel.fsh.model1,rel.fsh.model2)
anova(rel.fsh.model1,rel.fsh.model2,rel.fsh.model3)

summary(rel.fsh.model1)
# help("pvalues",package="lme4")
car::Anova(rel.fsh.model1)
lmerTest::ranova(rel.fsh.model1)
```

# Indicator species   
```{r}
#https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html#data-required-for-indicator-species-analysis  
library(indicspecies) #needs a matrix or data frame with sites as rows and species in columns  
```

```{r}
indic.melt <- ps.prune.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.tab <- indic.melt %>% 
  select(c(Names,RelAbund,Genus)) %>%
  distinct() %>% 
  pivot_wider(names_from = Genus,values_from=RelAbund) %>% 
  as.data.frame() 
rownames(indic.tab) <- indic.tab$Names
indic.tab$Names <- NULL

indic.patt <- indic.melt %>% 
  select(c(Names,Rookery)) %>% 
  distinct() %>% 
  mutate(Rookery=case_when(Rookery=="Rookery" ~ 1, .default=0))

indic.df <- indic.melt %>% 
  select(c(Names,Rookery,Site)) %>% 
  distinct()

perm_indic <-how(nperm=999)
setBlocks(perm_indic) <- with(indic.df,Site)

indval <- multipatt(indic.tab, indic.patt$Rookery,
                    #allow.negative=TRUE, 
                    control = perm_indic) 

summary(indval,indvalcomp=TRUE) #using indvalcomp=TRUE gives us the probability of finding the species in sites belonging to the site group 
```
```{r}
indval$sign %>% as.data.frame() %>% arrange(p.value) %>% filter(p.value <0.05) %>% rownames()

coverage(indic.tab,indval)
```

```{r}
indic.ch.melt <- ps.ch.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.ch.tab <- indic.ch.melt %>% 
  select(c(Names,RelAbund,Genus)) %>%
  distinct() %>% 
  pivot_wider(names_from = Genus,values_from=RelAbund) %>% 
  as.data.frame() 
rownames(indic.ch.tab) <- indic.ch.tab$Names
indic.ch.tab$Names <- NULL

indic.ch.patt <- indic.ch.melt %>% 
  select(c(Names,Rookery)) %>% 
  distinct() %>% 
  mutate(Rookery=case_when(Rookery=="Rookery" ~ 1, .default=0))

indic.ch.df <- indic.ch.melt %>% 
  select(c(Names,Rookery,Site)) %>% 
  distinct()

perm_indic.ch <-how(nperm=999)
setBlocks(perm_indic.ch) <- with(indic.ch.df,Site)

indval.ch <- multipatt(indic.ch.tab, indic.ch.patt$Rookery, 
                    control = perm_indic.ch,
                    allow.negative=TRUE) 

summary(indval.ch,indvalcomp=TRUE) #using indvalcomp=TRUE gives us the probability of finding the species in sites belonging to the site group 
```

```{r}
indic.fsh.melt <- ps.fsh.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.fsh.tab <- indic.fsh.melt %>% 
  select(c(Names,RelAbund,Genus)) %>%
  distinct() %>% 
  pivot_wider(names_from = Genus,values_from=RelAbund) %>% 
  as.data.frame() 
rownames(indic.fsh.tab) <- indic.fsh.tab$Names
indic.fsh.tab$Names <- NULL

indic.fsh.patt <- indic.fsh.melt %>% 
  select(c(Names,Rookery)) %>% 
  distinct() %>% 
  mutate(Rookery=case_when(Rookery=="Rookery" ~ 1, .default=0))

indic.fsh.df <- indic.fsh.melt %>% 
  select(c(Names,Rookery,Site)) %>% 
  distinct()

perm_indic.fsh <-how(nperm=999)
setBlocks(perm_indic.fsh) <- with(indic.fsh.df,Site)

indval.fsh <- multipatt(indic.fsh.tab, indic.fsh.patt$Rookery, 
                    control = perm_indic.fsh) 

summary(indval.fsh,indvalcomp=TRUE) #using indvalcomp=TRUE gives us the probability of finding the species in sites belonging to the site group 
```

## Phi coefficient of association  
(additional metric index to analyze association between a species and group of sites. This is a coefficient that emasures the correlation between binary vectors - presence/absence)
```{r}
indic.fsh.pa.melt <- ps.fsh.pa %>% tax_glom(.,"Species",NArm=TRUE) %>% #collapse by Species, removing anything not assigned to Species
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Species,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% 
  mutate(PA=case_when(is.nan(Abund) ~ 0,Abund>0 ~ 1, .default=0))

indic.fsh.pa.tab <- indic.fsh.pa.melt %>% 
  select(c(Names,PA,Species)) %>%
  distinct() %>% 
  pivot_wider(names_from = Species,values_from=PA) %>% 
  as.data.frame() 
rownames(indic.fsh.pa.tab) <- indic.fsh.pa.tab$Names
indic.fsh.pa.tab$Names <- NULL

indic.fsh.pa.patt <- indic.fsh.pa.melt %>% 
  select(c(Names,Rookery)) %>% 
  distinct() %>% 
  mutate(Rookery=case_when(Rookery=="Rookery" ~ 1, .default=0))

indic.fsh.pa.df <- indic.fsh.pa.melt %>% 
  select(c(Names,Rookery,Site)) %>% 
  distinct()

perm_indic.pa.fsh <-how(nperm=999)
setBlocks(perm_indic.pa.fsh) <- with(indic.fsh.pa.df,Site)

phi <- multipatt(indic.fsh.pa.tab, indic.fsh.pa.patt$Rookery,
                 func = "r.g", #corrects for the fact that some groups have more sites  
                 control = perm_indic.pa.fsh) 

summary(phi)
round((phi$str),3)
```

# IUCN  
```{r}
iucn.df <- ps.prune.rel.fsh.melt %>%
   mutate(IUCN.Category=case_when(IUCN.Category=="" ~ "Status not recorded in FishBase",
                                 is.na(IUCN.Category) ~ "Status not recorded in FishBase",
                                 .default=IUCN.Category)) %>% 
  arrange(Rookery,IUCN.Category) %>% 
  mutate(IUCN.Category=factor(IUCN.Category,levels=c("Endangered","Vulnerable",
                                                     "Near Threatened","Least Concern",
                                                     "Data Deficient",
                                                     "Status not recorded in FishBase"),
                              ordered=TRUE)
         ) %>% 
  group_by(Rookery, Site, IUCN.Category) %>%
  filter(Abundance > 0) %>%
  reframe(n=length(unique(Species)))

iucn.df %>% 
  mutate(IUCN.Category=factor(IUCN.Category,levels=c("Status not recorded in FishBase",
                                                     "Data Deficient","Least Concern",
                                                     "Near Threatened","Vulnerable","Endangered"),
                              ordered=TRUE)
         ) %>% arrange(Rookery,IUCN.Category)
```

```{r}
iucn.df %>%
  ggplot(aes(x=Rookery,y=n,fill=IUCN.Category)) +
#ps.prune.rel %>% 
  #plot_bar(x="SampleName_Short",fill="IUCN.Category") + 
  geom_bar(stat="identity",color="black") + 
  ggpubr::theme_pubr() +
  facet_wrap(~Site) + 
  labs(y="Number of species") + 
  theme(axis.title.x=element_blank())

#ggsave("Figures/n.spp.IUCN.site.png",width=7,height=5)
```

# Commercially Relevant  
```{r impspp_plot}
comm.sp <- ps.prune.rel.fsh.melt %>% 
  filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  #select(contains("Commer") | contains("Small.scale")) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% mutate(New=case_when(is.na(New) ~ FALSE,.default=New))

comm.sp.pa <- ps.prune.pa %>%
  subset_taxa(Phylum=="Chordata") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>%
  psmelt() %>%
  filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  #select(contains("Commer") | contains("Small.scale")) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% mutate(New=case_when(is.na(New) ~ FALSE,.default=New))
```
```{r}
comm.sp %>%
  ggplot(aes(x=Month,fill=New,y=Rel)) +
  geom_bar(stat="identity",position="stack") + 
  facet_grid(Site ~ Rookery) +
  guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) +
  theme_bw() +
  labs(y="Relative Abundance")

comm.sp.pa %>%
  ggplot(aes(x=Month,fill=New,y=Rel)) + 
  geom_bar(stat="identity",position="stack") + 
  facet_grid(Site ~ Rookery) +
  guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) +
  theme_bw() +
  labs(y="Relative Abundance")

#ggsave("Figures/commercial.spp.png")
```

# Habitat  
```{r}
colnames(ps.prune.rel.fsh.melt[100:125])

ps.prune.rel.fsh.melt %>% mutate(Type=case_when(Benthic==-1 ~ "Benthic",
                                            Demersal==-1 ~ "Demersal",
                                            Pelagic==-1 ~ "Pelagic"
                                            )) %>% ggplot(aes(x=SampleName_Short,y=Abundance,fill=Type)) + geom_bar(stat="identity")
```


# Session Info   
```{r}
save.image("RData/08_Analysis_bray.RData")
```

```{r}
sessionInfo()
```

