
# Load  
Load libraries  
```{r,message=FALSE}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(vegan)
library(lme4)
```

Load data  
```{r load}
dir_data <- '/Users/lauragivens/Desktop/R/BZrookery_eDNA/Rdata'
dir_results <- "/Volumes/Fuji/Mangroves/2025_0319_Givens_Canty_Rookery_COI/cutadapt/results"

ps.nt <- readRDS(paste0(dir_results,"/prune.decontam.ps.nt.troph.rds"))
ps.mar <- readRDS(paste0(dir_results,"/prune.decontam.ps.troph.rds"))
ps.bold <- readRDS(paste0(dir_results,"/prune.decontam.ps.bold.rds"))

ps.nt <- ps.nt %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")
ps.mar <- ps.mar %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")
ps.bold <- ps.bold %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")

taxa_nt <- readRDS(paste0(dir_results,"/prune.decontam.ps.nt.troph.rds"))
taxa_mar <- readRDS(paste0(dir_results,"/taxtable.wtroph.rds"))
taxa_bold <- readRDS(paste0(dir_results,"/taxtable.bold.rds"))

meta <- readRDS(paste0(dir_results,"/metadata.edit.rds"))
samplelist <- meta %>% filter(Rookery!="Practice" & Rookery !="Blank")
```
Or alternatively,  
```{r}
load("RData/08_Analysis.RData")
```

Convert rookery to 0/1  
```{r}
samplelist<- samplelist %>% mutate(RookeryBinary = case_when(Rookery=="Rookery" ~ 1,
                                Rookery!="Rookery" ~ 0),.after=Rookery )
samplelist <- samplelist %>% mutate(NewName=word(rownames(.),sep="_",end=2))

sample_data(ps.nt) <- samplelist
sample_data(ps.mar) <- samplelist
sample_data(ps.bold) <- samplelist
```

```{r}
ggplot2::theme_set(theme_cowplot() + theme(
  legend.position = "bottom",
  panel.border = element_rect(color='black'),
  legend.box.margin = margin(5,5,5,5),
  legend.spacing = unit(0,"pt"),
  plot.title = element_text(face="bold"),
  plot.caption = element_text(face="italic"),
  text = element_text(size=(15),family="Times"),
  legend.key.height = unit(10,"pt"),
  legend.key.width = unit(10,"pt")
))
```

# Conversions  
Presence/absence  
```{r}
ps.nt.pa <- ps.nt %>% transform_sample_counts(., function(abund) 1*(abund > 0))
ps.mar.pa <- ps.mar %>% transform_sample_counts(., function(abund) 1*(abund > 0))
ps.bold.pa <- ps.bold %>% transform_sample_counts(., function(abund) 1*(abund > 0))
```

Relative abundance  
```{r}
ps.nt.rel <- ps.nt %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.mar.rel <- ps.mar %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.bold.rel <- ps.bold %>% transform_sample_counts(.,function(x) x/sum(x) ) 
```

Vegan-ready  
```{r}
tab.nt <- otu_table(ps.nt) %>% as.matrix() %>% as.data.frame() %>% t()
tab.mar <- otu_table(ps.mar) %>% as.matrix() %>% as.data.frame() %>% t()
tab.bold <- otu_table(ps.bold) %>% as.matrix() %>% as.data.frame() %>% t()
```

```{r}
tab.nt.pa <- otu_table(ps.nt.pa) %>% as.matrix() %>% as.data.frame() %>% t()
tab.mar.pa <- otu_table(ps.mar.pa) %>% as.matrix() %>% as.data.frame() %>% t()
tab.bold.pa <- otu_table(ps.bold.pa) %>% as.matrix() %>% as.data.frame() %>% t()
```

```{r}
tab.nt.rel <- otu_table(ps.nt.rel) %>% as.matrix() %>% as.data.frame() %>% t()
tab.mar.rel <- otu_table(ps.mar.rel) %>% as.matrix() %>% as.data.frame() %>% t()
tab.bold.rel <- otu_table(ps.bold.rel) %>% as.matrix() %>% as.data.frame() %>% t()
```

# Main questions  
1. does rookery presence have an impact on fish community composition?  
2. does rookery presence have an impact on whole community composition?  
3. What species are impacted?  (Commercial/non-commercial; trophic index as well)

# 1. Analysis (Fish)  
Select only fish BOLD samples (since the mar dataset is already only chordates)
May need to further subsample this actually
```{r}
ps.bold.pa.fsh <- ps.bold.pa %>% subset_taxa(Phylum=="Chordata" & Class!="Ascidiacea" & Class != "Reptilia" & Class != "Aves") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
tab.bold.pa.fsh <- otu_table(ps.bold.pa.fsh) %>% as.matrix() %>% as.data.frame() %>% t()
```

## Jaccard distance  
```{r}
jacc.mar.dist <- distance(ps.mar.pa,method='jaccard')
```
```{r}
jacc.bold.fsh.dist <- distance(ps.bold.pa.fsh,method='jaccard')
```

### distance-based Redundancy analysis   
We want to know how much variation in community composition can be explained by the environmental variables we measured.  
To do that, we are going to use a distance-based redundancy analysis (db-RDA) to identify the significant variables. (Different from a regular RDA because it can use non-Euclidean dissimilarity indices - if you use a Euclidean index, you will get the same result as rda() ).   

Make base rda  
```{r}
#using phyloseq
#rda.0.mar <- ordinate(ps.mar.pa,distance=jacc.mar.dist)
#rda.0.bold <- ordinate(ps.bold.pa.fsh,distance=jacc.bold.dist)

#using vegan
rda.0.mar <- dbrda(tab.mar.pa~1,data=samplelist)
rda.0.bold.fsh <- dbrda(tab.bold.pa.fsh~1,data=samplelist)
```

Add environmental variables  
```{r}
#using phyloseq  
#rda.jacc.mar <- ordinate(ps.mar.pa,method="RDA",distance=jacc.mar.dist,formula=Site+Month+Side+RookeryBinary)
#rda.jacc.bold.fsh <- ordinate(ps.bold.pa.fsh,method="RDA",distance=jacc.bold.fsh.dist,formula=Site+Month+Side+RookeryBinary)

#using vegan 
rda.jacc.mar <- dbrda(jacc.mar.dist ~ Site+Month+Side+RookeryBinary,samplelist)
rda.jacc.bold.fsh <- dbrda(jacc.bold.fsh.dist ~ Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.fsh))))
```

Calculate significance  
```{r}
#step.mar <- ordistep(rda.0.mar,scope=formula(rda.jacc.mar),perm.max=1000)
#step.bold.fsh <- ordistep(rda.0.bold.fsh,scope=formula(rda.jacc.bold.fsh),perm.max=1000)

step.mar <- vegan::anova.cca(rda.jacc.mar,permutations=1000,by='terms') #anova.cca performs an ANOVA like permutation test for RDA to assess the significance of constraints  
step.bold.fsh <- vegan::anova.cca(rda.jacc.bold.fsh,permutations=1000,by='terms')
```

```{r}
step.mar
```

```{r}
step.bold.fsh
```

#### Visualize  
```{r}
plot_ordination(ps.mar.pa,rda.jacc.mar,color="Rookery",shape="Side") + facet_wrap(~Site)
```

```{r}
plot_ordination(ps.bold.pa.fsh,rda.jacc.bold.fsh,color="Rookery",shape="Side") + facet_wrap(~Site)
```

```{r}
arrow_map_rda = aes(xend = dbRDA1, 
                yend = dbRDA2,
                x = 0, y = 0, 
                shape = NULL, color = NULL,label=labels
                )
label_map_rda = aes(x = 1.5 * dbRDA1, 
                y = 1.5 * dbRDA2, 
                shape = NULL, color = NULL,label=labels
                )
arrowhead = arrow(length = unit(0.03, "npc"))

# get scores from rda  
sigscrs_rda_mar <- scores(rda.jacc.mar,display="bp") %>% as.data.frame() %>% filter(., abs(dbRDA1)>0.05 & abs(dbRDA2)>0.05)
# add labels  
arrowdf_rda_mar <- data.frame(labels = rownames(sigscrs_rda_mar), sigscrs_rda_mar)

plot_ordination(ps.mar.pa,rda.jacc.mar) + 
  geom_point(aes(color=Rookery)) + 
  geom_segment(arrow_map_rda, data=arrowdf_rda_mar,arrow=arrowhead) + #add vector line  
  #geom_text(label_map_rda, size = 4, data = arrowdf_rda_mar, color = "black")  #add label 
    ggrepel::geom_text_repel(label_map_rda,data = arrowdf_rda_mar, 
                           cex = 4, direction = "both", 
                           segment.size = 0.25,color="black") #+ facet_wrap(~Site)
```

### PERMANOVA  
```{r}
adonis.jacc.mar <- vegan::adonis2(jacc.mar.dist~Site+Month+Side+RookeryBinary,samplelist,perm=1000,by='terms')
adonis.jacc.mar


adonis.jacc.bold.fsh <- vegan::adonis2(jacc.bold.fsh.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.fsh))),perm=1000,by='terms')
adonis.jacc.bold.fsh
```

### test strata  
```{r}
perm.OR <-how(nperm=999)
setBlocks(perm.OR) <- with(samplelist,Site)
# vegan::adonis2(obs.richness~Location*Season,permutations=perm.OR,meta.subset.m)

 vegan::adonis2(jacc.mar.dist~Site+Side+RookeryBinary,samplelist,by='terms',permutations=perm.OR)
```

## GLM  
Want to know if any specific fish species has a response to rookery presence  
Want to know if there is a difference between trophic composition  
poisson distribution because we essentially have count data where the response variable = number of events occurring in a specific timeframe  
https://medium.com/@sahin.samia/a-comprehensive-introduction-to-generalized-linear-models-fd773d460c1d

Trophic index ~ Rookery + Month + Side + Site 
```{r}
glm()
```

Does trophic diversity differ between site types?  
```{r}
ps.mar.rel.melt <- psmelt(ps.mar.rel) 

lmer(Abundance ~ Trophic.Round + Rookery + Month + Side + (1|Site) , data = ps.mar.rel.melt)
#singular fit warning 
```

# 2. Analysis (non-fish chordates)    
```{r}
get_taxa_unique(ps.bold.pa %>% subset_taxa(Phylum=="Chordata"),"Class")
get_taxa_unique(ps.nt.pa %>% subset_taxa(Phylum=="Chordata"),"Class")
```

Presence/absence, nt database  
```{r}
ps.nt.pa.othr <- ps.nt.pa  %>% subset_taxa(Phylum=="Chordata" & Class!="Actinopteri" & Class != "Ascidiacea") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
tab.nt.pa.othr <- otu_table(ps.nt.pa.othr) %>% as.matrix() %>% as.data.frame() %>% t()

jacc.nt.othr.dist <- distance(ps.nt.pa.othr,method='jaccard')

rda.jacc.nt.othr <- dbrda(jacc.nt.othr.dist ~ Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.nt.pa.othr))))


perm.nt <-how(nperm=999)
setBlocks(perm.nt) <- with(as.data.frame(as.matrix(sample_data(ps.nt.pa.othr))),Site)

adonis.jacc.nt.othr <- vegan::adonis2(jacc.nt.othr.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.nt.pa.othr))),permutations=perm.nt,by='terms')
adonis.jacc.nt.othr

#without blocks
#vegan::adonis2(jacc.nt.othr.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.nt.pa.othr))),permutations=1000,by='terms')

plot_ordination(ps.nt.pa.othr,rda.jacc.nt.othr,color="Rookery",shape="Month") + facet_wrap(~Site)
```

```{r}
plot_ordination(ps.nt.pa.othr,rda.jacc.nt.othr,color="Rookery",shape="Month") + facet_wrap(~Site)
```

Presence/absence, bold database  
```{r}
ps.bold.pa.othr <- ps.bold.pa  %>% subset_taxa(Phylum=="Chordata" & Class!="Actinopterygii" & Class != "Ascidiacea") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
tab.bold.pa.othr <- otu_table(ps.bold.pa.othr) %>% as.matrix() %>% as.data.frame() %>% t()

jacc.bold.othr.dist <- distance(ps.bold.pa.othr,method='jaccard')

rda.jacc.bold.othr <- dbrda(jacc.bold.othr.dist ~ Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.othr))))


perm.othr <-how(nperm=999)
setBlocks(perm.othr) <- with(as.data.frame(as.matrix(sample_data(ps.bold.pa.othr))),Site)

adonis.jacc.bold.othr <- vegan::adonis2(jacc.bold.othr.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.othr))),permutations=perm.othr,by='terms')
adonis.jacc.bold.othr
#without blocks
#vegan::adonis2(jacc.bold.othr.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.othr))),permutations=1000,by='terms')

plot_ordination(ps.bold.pa.othr,rda.jacc.bold.othr,color="Rookery",shape="Month") + facet_wrap(~Site)
```

```{r}
ps.nt.pa.othr.melt <- psmelt(ps.nt.pa.othr)
```

```{r}
kruskal.test(ps.nt.pa.othr.melt$Abundance ~ ps.nt.pa.othr.melt$Rookery)
```

### Birds specifically  
Relative abundance, nt database  
```{r}
ps.nt.rel.brd <- ps.nt.rel  %>% subset_taxa(Class=="Aves") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
tab.nt.rel.brd <- otu_table(ps.nt.rel.brd) %>% as.matrix() %>% as.data.frame() %>% t()

bray.nt.brd.dist <- distance(ps.nt.rel.brd,method='bray')

rda.bray.nt.brd <- dbrda(bray.nt.brd.dist ~ Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.nt.rel.brd))))

adonis.bray.nt.brd <- vegan::adonis2(bray.nt.brd.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.nt.rel.brd))),perm=1000,by='terms')
adonis.bray.nt.brd

perm.nt.brd <-how(nperm=999)
setBlocks(perm.nt.brd) <- with(as.data.frame(as.matrix(sample_data(ps.nt.rel.brd))),Month)
 vegan::adonis2(bray.nt.brd.dist~Site+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.nt.rel.brd))),by='terms',permutations=perm.nt.brd)

plot_ordination(ps.nt.rel.brd,rda.bray.nt.brd,color="Rookery",shape="Month") + facet_wrap(~Site) + labs(title="Avian occurrences",subtitle="db-RDA on bray distance calculated with relative abundance dataset",caption="Taxonomy annotated with NCBI nt database")
```

pa, bold 
```{r}
ps.bold.pa.brd <- ps.bold.pa  %>% subset_taxa(Class=="Aves") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
ps.bold.pa.brd.melt <- psmelt(ps.bold.pa.brd)
```
```{r}
kruskal.test(ps.bold.pa.brd.melt$Abundance ~ ps.bold.pa.brd.melt$Rookery)
```

Relative abundance, bold database  
```{r}
ps.bold.rel.brd <- ps.bold.rel  %>% subset_taxa(Class=="Aves") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
tab.bold.rel.brd <- otu_table(ps.bold.rel.brd) %>% as.matrix() %>% as.data.frame() %>% t()

bray.bold.brd.dist <- distance(ps.bold.rel.brd,method='bray')

rda.bray.bold.brd <- dbrda(bray.bold.brd.dist ~ Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.rel.brd))))

adonis.bray.bold.brd <- vegan::adonis2(bray.bold.brd.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.rel.brd))),perm=1000,by='terms')
adonis.bray.bold.brd

perm.brd <-how(nperm=999)
setBlocks(perm.brd) <- with(as.data.frame(as.matrix(sample_data(ps.bold.rel.brd))),Month)
 vegan::adonis2(bray.bold.brd.dist~Site+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.rel.brd))),by='terms',permutations=perm.brd)

plot_ordination(ps.bold.rel.brd,rda.bray.bold.brd,color="Rookery",shape="Side") + facet_wrap(~Site)
```

Presence/absence, bold database  
```{r}
ps.bold.pa.brd <- ps.bold.pa  %>% subset_taxa(Class=="Aves") %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
tab.bold.pa.brd <- otu_table(ps.bold.pa.brd) %>% as.matrix() %>% as.data.frame() %>% t()

bray.bold.pa.brd.dist <- distance(ps.bold.pa.brd,method='bray')

rda.bray.bold.pa.brd <- dbrda(bray.bold.pa.brd.dist ~ Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.brd))))

adonis.bray.pa.bold.brd <- vegan::adonis2(bray.bold.pa.brd.dist~Site+Month+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.brd))),perm=1000,by='terms')
adonis.bray.bold.brd

 vegan::adonis2(bray.bold.pa.brd.dist~Site+Side+RookeryBinary,as.data.frame(as.matrix(sample_data(ps.bold.pa.brd))),by='terms',permutations=perm.brd)

plot_ordination(ps.bold.pa.brd,rda.bray.bold.brd,color="Rookery",shape="Side") + facet_wrap(~Site)
```

#### GLM  
```{r}
ps.nt.rel.brd.melt <- psmelt(ps.nt.rel.brd)
head(ps.nt.rel.brd.melt)
glm(Abundance~Species*RookeryBinary+Site+Month+Side,data=ps.nt.rel.brd.melt,family=gaussian)

#lmer(Abundance~Species*RookeryBinary+(1|Site)+Month+Side,data=ps.nt.rel.brd.melt)
```

```{r}
kruskal.test(ps.nt.rel.brd.melt$Abundance ~ ps.nt.rel.brd.melt$Rookery)
```

# 3. Analysis (Community)   
## Jaccard distance  
```{r}
jacc.nt.dist <- distance(ps.nt.pa,method='jaccard')
jacc.bold.dist <- distance(ps.bold.pa,method='jaccard')
```

### distance-based Redundancy analysis   
We want to know how much variation in community composition can be explained by the environmental variables we measured.  
To do that, we are going to use a distance-based redundancy analysis (db-RDA) to identify the significant variables. (Different from a regular RDA because it can use non-Euclidean dissimilarity indices - if you use a Euclidean index, you will get the same result as rda() ).   

Make base rda  
```{r}
rda.0.bold <- dbrda(tab.bold.pa~1,data=samplelist)
```

Add environmental variables  
```{r}
rda.jacc.nt <- dbrda(jacc.nt.dist ~ Site+Month+Side+RookeryBinary,samplelist)
rda.jacc.bold <- dbrda(jacc.bold.dist ~ Site+Month+Side+RookeryBinary,samplelist)
```

Calculate significance  
```{r}
step.nt <- vegan::anova.cca(rda.jacc.nt,permutations=1000,by='terms')
step.bold <- vegan::anova.cca(rda.jacc.bold,permutations=1000,by='terms')
```

```{r}
step.nt
```

```{r}
step.bold
```

#### Visualize  
```{r}
plot_ordination(ps.nt.pa,rda.jacc.nt,color="Rookery",shape="Month") + facet_wrap(~Site)# + stat_ellipse()
```

```{r}
plot_ordination(ps.nt.pa,rda.jacc.nt,color="Month",shape="Rookery") + facet_wrap(~Site) 
```

```{r}
plot_ordination(ps.bold.pa,rda.jacc.bold,color="Month",shape="Rookery") + facet_wrap(~Site) 
```

### PERMANOVA  
```{r}
adonis.jacc.nt <- vegan::adonis2(jacc.nt.dist~Site+Month+Side+RookeryBinary,samplelist,perm=1000,by='terms')
adonis.jacc.nt
```
with strata
```{r}
perm.nt <-how(nperm=999)
setBlocks(perm.nt) <- with(samplelist,Site)

vegan::adonis2(jacc.nt.dist~Site+Month+Side+RookeryBinary,samplelist,perm=perm.nt,by='terms')
```

```{r}
adonis.jacc.bold <- vegan::adonis2(jacc.bold.dist~Site+Month+Side+RookeryBinary,samplelist,perm=1000,by='terms')
adonis.jacc.bold
```
with strata
```{r}
perm.bld <-how(nperm=999)
setBlocks(perm.bld) <- with(samplelist,Site)

vegan::adonis2(jacc.bold.dist~Site+Month+Side+RookeryBinary,samplelist,perm=perm.bld,by='terms')
```
## GLM  



# 4. Heatmap  

## 4b. fish  
```{r}
mean_cutoff = 0.01
mean_test = function(x) {
  mean(x) >= mean_cutoff
}

melt.mar.rel <- ps.mar.rel %>% tax_glom(.,"Genus",NArm=FALSE) %>% filter_taxa(mean_test, prune=TRUE) %>% psmelt()

melt.mar.rel %>% group_by(Month,Site,Rookery,Family,Genus) %>% summarize(Abund=sum(Abundance)) %>%  ggplot(aes(x=Rookery,y=Genus,fill=Abund)) + geom_tile(,group="Family") + facet_wrap(~Site,scales="free") + theme(axis.text.x = element_text(angle=90))
```
```{r}
melt.mar.rel
```


```{r}
gen.melt.mar.rel <- melt.mar.rel %>% ungroup() %>% mutate(Names=word(NewName,sep="R")) %>% group_by(Genus,Names,Rookery,Site) %>% summarize(Abund=sum(Abundance)) %>% ungroup() %>% group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>% mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund)) %>% group_by(Site) %>% arrange(.,Rookery,.by_group=TRUE) %>% ungroup() %>% select(-c(Rookery,Site))

gen.rel.tab <- gen.melt.mar.rel %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Names,values_from=RelAbund) %>% as.data.frame()
rownames(gen.rel.tab) <- gen.rel.tab$Genus
gen.rel.tab$Genus <- NULL
gen.rel.dt <- melt.mar.rel %>% ungroup() %>% mutate(Names=word(NewName,sep="R")) %>% group_by(.,Names) %>% summarize(Month=unique(Month),Site=unique(Site),Rookery=unique(Rookery))


 
 
 

annotation_col = gen.rel.dt %>% group_by(Rookery,Site) %>% 
  arrange("Rookery",.by_group=TRUE) %>% ungroup() %>% 
  select(c(Names,Rookery,Site)) %>% 
  mutate(Rookery=factor(Rookery,ordered=TRUE),
         Site=factor(Site,ordered=TRUE)) %>% 
  data.frame(.,row.names = "Names")
```

```{r}
pheatmap::pheatmap(gen.rel.tab,cluster_cols=FALSE,annotation_col=annotation_col,gaps_col=c(8,16))
```

## 4b. other verts  
```{r}
melt.bold.rel <- ps.bold.rel %>% 
  subset_taxa(.,Phylum=="Chordata") %>%
  tax_glom(.,"Genus",NArm=TRUE) %>% 
  #filter_taxa(mean_test, prune=TRUE) %>%
  psmelt()
```
```{r}
melt.bold.rel
```


```{r}
gen.melt.bold.rel <- melt.bold.rel %>% ungroup() %>% mutate(Names=word(NewName,sep="R")) %>% group_by(Genus,Names,Rookery,Site) %>% summarize(Abund=sum(Abundance)) %>% ungroup() %>% group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>% mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund)) %>% group_by(Site) %>% arrange(.,Rookery,.by_group=TRUE) %>% ungroup() %>% select(-c(Rookery,Site))

gen.rel.tab <- gen.melt.bold.rel %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Names,values_from=RelAbund) %>% as.data.frame() 
rownames(gen.rel.tab) <- gen.rel.tab$Genus
gen.rel.tab$Genus <- NULL
gen.rel.tab <- gen.rel.tab %>% mutate(tmp=rowSums(.)) %>% arrange(desc(tmp)) %>% filter(tmp>0.01) %>% select(-tmp)

gen.rel.dt <- melt.bold.rel %>% ungroup() %>% mutate(Names=word(NewName,sep="R")) %>% group_by(.,Names) %>% summarize(Month=unique(Month),Site=unique(Site),Rookery=unique(Rookery))


annotation_col = gen.rel.dt %>% group_by(Rookery,Site) %>% 
  arrange("Rookery",.by_group=TRUE) %>% ungroup() %>% 
  select(c(Names,Rookery,Site)) %>% 
  mutate(Rookery=factor(Rookery,ordered=TRUE),
         Site=factor(Site,ordered=TRUE)) %>% 
  data.frame(.,row.names = "Names")

annotation_row = melt.bold.rel %>% select(c(Class,Genus)) %>% distinct() %>% group_by(Class,Genus) %>% arrange("Class",.by_group = TRUE) %>% ungroup() %>% mutate(Class=factor(Class,ordered=TRUE)) %>% data.frame(.,row.names="Genus")
```

```{r}
pheatmap::pheatmap(gen.rel.tab,cluster_cols=FALSE,cluster_rows=FALSE,annotation_row=annotation_row,annotation_col=annotation_col,gaps_col=c(8,16),cellheight = 5)
```

```{r}
pheatmap::pheatmap((gen.rel.tab %>% filter(rownames(.) %in% rownames(annotation_row %>% filter(Class!="Actinopterygii" & Class != "Ascidiacea")))),cluster_cols=FALSE,cluster_rows=FALSE,annotation_row=annotation_row,annotation_col=annotation_col,gaps_col=c(8,16))
```

# Indicator species   
```{r}
#https://cran.r-project.org/web/packages/indicspecies/vignettes/IndicatorSpeciesAnalysis.html#data-required-for-indicator-species-analysis  
library(indicspecies) #needs a matrix or data frame with sites as rows and species in columns  
```

MAR fish species, relative abundance  
```{r}
indic.melt <- ps.mar.rel %>% tax_glom(.,"Genus",NArm=TRUE) %>% #collapse by genus, removing anything not assigned to genus
  psmelt() %>% 
  mutate(Names=word(NewName,sep="R")) %>% #make a sample name that doesn't include replicate or leeward/windwward
  group_by(Genus,Names,Rookery,Site) %>% 
  summarize(Abund=sum(Abundance)) %>% ungroup() %>% 
  group_by(Names) %>% mutate(RelAbund=Abund/sum(Abund)) %>%
  mutate(RelAbund=case_when(is.nan(RelAbund) ~ 0,!is.nan(RelAbund) ~ RelAbund))

indic.tab <- indic.melt %>% select(c(Names,RelAbund,Genus)) %>% distinct() %>% pivot_wider(names_from = Genus,values_from=RelAbund) %>% as.data.frame() 
rownames(indic.tab) <- indic.tab$Names
indic.tab$Names <- NULL
```

```{r}
perm_indic <-how(nperm=999)
setBlocks(perm_indic) <- with(gen.rel.dt,Site)

indval <- multipatt(indic.tab, gen.rel.dt$Rookery, 
                    control = perm_indic) 
summary(indval)
```

(whole dataset)  
```{r}
perm_indic <-how(nperm=999)
setBlocks(perm_indic) <- with(as.data.frame(sample_data(ps.mar.rel)),Site)

indval <- multipatt(t(otu_table(ps.mar.rel)), sample_data(ps.mar.rel)$Rookery, 
                    control = perm_indic) 
summary(indval)
```

# Session Info   
```{r}
save.image("RData/08_Analysis.RData")
```

```{r}
sessionInfo()
```

