---
title: "Untitled"
author: "Laura Givens"
date: "2025-09-09"
output: html_document
---

# Load  
Load libraries  
```{r,message=FALSE}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(patchwork)
```

Load data  
```{r load}
dir_data <- '/Users/lauragivens/Desktop/R/BZrookery_eDNA/Rdata'
dir_man <- "/Volumes/Fuji/Mangroves"
dir_results <- "/Volumes/Fuji/Mangroves/2025_0319_Givens_Canty_Rookery_COI/cutadapt/results"
#dir_results <- "/Users/lauragivens/Desktop/MangroveResults"

ps.og <- readRDS(paste0(dir_results,"/prune.decontam.combined.ps.rds"))
#taxa_nt_troph <- readRDS(paste0(dir_results,"/taxtable.v2.nt.wtroph.rds"))
#tax.combined.df <- readRDS(paste0(dir_results,"/tax.combined.df.rds"))
taxa.traits.df <- readRDS(paste0(dir_results,"/taxa.combined.traitsdf.rds"))
```

```{r}
ggplot2::theme_set(theme_cowplot() + theme(
  legend.position = "bottom",
  panel.border = element_rect(color='black'),
  legend.box.margin = margin(5,5,5,5),
  legend.spacing = unit(0,"pt"),
  plot.title = element_text(face="bold"),
  plot.caption = element_text(face="italic"),
  text = element_text(size=(15),family="Times"),
  legend.key.height = unit(10,"pt"),
  legend.key.width = unit(10,"pt")
))
```

# Update ps  
```{r}
tax <- tax_table(as.matrix(taxa.traits.df %>% select(-seqid)))
tax_table(ps.og) <- tax
```

```{r}
ps <- ps.og 
```

keep chordates  
```{r}
ps <- ps %>% 
  prune_samples(sample_sums(.)>50,.) %>% 
  subset_samples(Rookery!="Practice") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% 
  prune_samples(sample_sums(.)>0,.)

ps.ch <-  ps %>% subset_taxa(.,Phylum=="Chordata") %>% 
  subset_taxa(.,Class != "Reptilia" & Class != "Ascidiacea" & Class != "Aves") 

ps.pa <- ps %>% subset_taxa(.,Phylum=="Chordata") %>% 
  subset_taxa(.,Class != "Reptilia" & Class != "Ascidiacea" & Class != "Aves") %>%
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.pa.melt <- psmelt(ps.pa)
```
keep chordates that make up at least 1 pct of taxa  
```{r,eval=FALSE}
mean_cutoff = 0.01
mean_test = function(x) {
  mean(x) >= mean_cutoff
}
 ps.1pct <- ps %>%
   filter_taxa(mean_test, prune=TRUE) 

 ps.1pct.pa <- ps.1pct %>% 
   subset_taxa(.,Phylum=="Chordata") %>% 
  subset_taxa(.,Class != "Reptilia" & Class != "Ascidiacea" & Class != "Aves") %>%
  transform_sample_counts(., function(abund) 1*(abund > 0))
 
ps.1pct.pa.melt <- psmelt(ps.1pct.pa)
``` 
keep taxa seen at least twice in 2 or more samples  
```{r} 
# keep only taxa seen at least twice in at least 2% of samples
#ps.prune <- filter_taxa(ps, function(x) sum(x >= 2) > (0.02*length(x)), TRUE)

ps.prune <- filter_taxa(ps, function(x) sum(x >= 2) >= 2, TRUE)

ps.prune.pa <- ps.prune %>% 
  subset_taxa(.,Phylum=="Chordata") %>% 
  subset_taxa(.,Class != "Reptilia" & Class != "Ascidiacea" & Class != "Aves") %>%
  transform_sample_counts(., function(abund) 1*(abund > 0))

ps.prune.pa.melt <- psmelt(ps.prune.pa)
```
Compare  
```{r}
ps
#ps.1pct
ps.prune

ps.pa
#ps.1pct.pa
ps.prune.pa
```
Save
```{r}
write.csv((ps.prune.pa.melt %>% filter(Abundance > 0)),"/Users/lauragivens/Downloads/bzrookery.fish.csv")
```

# Descriptive Stats  
Sequence distribution  
```{r distribution}
plot_bar(ps)
```

Unique taxa  
```{r num_taxa}
  tibble(Rank = rank_names(ps)) %>% 
    rowwise %>%
    mutate(Unique = length(get_taxa_unique((ps),Rank)))
```
Unique chordate taxa  
```{r num_cbhordate_taxa}
 tibble(Rank = rank_names(ps.ch)) %>% 
    rowwise %>%
    mutate(Unique = length(get_taxa_unique((ps.ch),Rank)))
```

# Top taxa  
```{r}
top100 <- names(sort(taxa_sums(ps.prune.pa %>%
  subset_taxa(.,Phylum=="Chordata") %>% 
  subset_taxa(.,Class != "Reptilia" & Class != "Ascidiacea" & Class != "Aves")), decreasing=TRUE))[1:100]

t100.tab <- ps.prune.pa %>% prune_taxa(top100, .) %>% tax_table() %>% as.matrix() %>% as.data.frame()
t100.tab %>% rownames_to_column("seqid") %>% .[match(top100,.$seqid),] %>% filter(Class=="Actinopterygii") #%>% .$Species %>% unique()

#t100.tab %>% 
#  rownames_to_column("seqid") %>%
#  .[match(top100,.$seqid),] %>% 
ps.prune.pa.melt %>% 
  filter(Abundance > 0 & Class=="Actinopterygii") %>%
  group_by(Species) %>%
  summarize(Species=unique(Species),
            n=sum(Abundance)
            ) %>% arrange(desc(n))
```

Confirm old top taxa is because of the coverage:  
```{r,eval=FALSE}
setwd("/Users/lauragivens/Desktop/MangroveResults")
marblastResults<-read.table('dada2.uniques.BLAST.martaxid.v1.tsv',header=FALSE,stringsAsFactors=FALSE)
maraccessions<-strsplit(marblastResults[,2],'\\|') #pull accession numbers into a list
martaxaId<-accessionToTaxa(maraccessions,"~/accessionTaxa.sql") 
martaxResults <- as.data.frame(getTaxonomy(unique(martaxaId),'~/accessionTaxa.sql')) %>% 
                    rownames_to_column(.,"taxid") %>% mutate(.,taxid=str_trim(taxid))
t1 <- marblastResults[1:2] #qseqid sseqid
t1$taxid <- unlist(martaxaId) %>% as.character()
t2 <- merge(t1,martaxResults,by.x='taxid',all.x=TRUE) %>% rename('qseqid'="V1","sseqid"="V2")
  
merge(marblastResults,t1,by='V2',all.x=TRUE) %>% 
  merge(.,martaxResults,by="taxid",all.x=TRUE) %>% 
  filter(species%in% c("Xyrichtys martinicensis", "Apogon planifrons", "Diodon hystrix")) #previous top taxa
#these are only covered by ~30 bp
```

# Richness  
```{r rich-samplename}
plot_richness(ps,x="SampleName",measures = c("Observed","Shannon"),color="Rookery")
```

```{r}
tab.pa <- otu_table(ps.pa) %>% as.matrix() %>% as.data.frame() %>% t()

richdf <- rowSums(tab.pa) %>% 
  as.data.frame() %>% 
  `colnames<-` ("samplerichness") %>% 
  mutate(Sample=rownames(.)) %>% left_join(select((sample_data(ps.ch) %>% as.matrix() %>% as.data.frame() %>% mutate(Sample=rownames(.))),Sample,Month,Site,Rookery))
```
```{r}
wrap_plots(
ggplot(richdf) + geom_boxplot(aes(x=Month,y=samplerichness)) + labs(x="Month",y="Sample Richness"),
ggplot(richdf) + geom_boxplot(aes(x=Site,y=samplerichness)) + labs(x="Site",y="Sample Richness"),
ggplot(richdf) + geom_boxplot(aes(x=Rookery,y=samplerichness)) + labs(x="Rookery",y="Sample Richness")
)


wrap_plots(
ggplot(richdf) + geom_boxplot(aes(x=Site,y=samplerichness)) + labs(x="Site",y="Number of taxa"),
ggplot(richdf) + geom_boxplot(aes(x=Rookery,y=samplerichness)) + labs(x="Rookery",y="Number of taxa")
) & theme_bw() & theme(axis.title.x=element_blank())

#ggsave("/Users/lauragivens/Downloads/richness.boxplots.png",width=10,height=5,bg="white")
```

## Species of importance   
```{r impspp_plot}
comm.sp <- ps.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  #select(contains("Commer") | contains("Small.scale")) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% mutate(New=case_when(is.na(New) ~ FALSE,.default=New))

comm.sp.prune <- ps.prune.pa.melt %>% filter(Abundance > 0) %>% 
  group_by(Site,Month) %>%
  mutate(Rel=Abundance/sum(Abundance)) %>% ungroup() %>% 
  #select(contains("Commer") | contains("Small.scale")) %>%
  mutate(Commercial.FSH=ifelse(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes"),"TRUE","FALSE"),.after = Reported.Commercial.FishbaseSpeciesPage) %>% rowwise() %>% 
  mutate(New= any(c_across(contains("Commer") | contains("Small.scale"))=="TRUE")
  ) %>% mutate(New=case_when(is.na(New) ~ FALSE,.default=New))
```
```{r}
comm.sp %>%
  ggplot(aes(x=Month,fill=New,y=Rel)) + geom_bar(stat="identity",position="stack") + facet_grid(Site ~ Rookery) + guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) + theme_bw() + labs(y="Relative Abundance")

comm.sp.prune %>%
  ggplot(aes(x=Month,fill=New,y=Rel)) + geom_bar(stat="identity",position="stack") + facet_grid(Site ~ Rookery) + guides(fill=guide_legend(title="Critical species \n (Commercial and \n small-scale fisheries)")) + theme_bw() + labs(y="Relative Abundance")

#ggsave("/Users/lauragivens/Downloads/hannahfig1.png")
```
```{r}
comm.sp %>% filter(New==TRUE) %>% 
  mutate(Small.scale.Fisheries=case_when(Small.scale.Fisheries==TRUE ~ 1,.default=0),
         Sport=case_when(Gamefish==TRUE ~ 1, .default=0),
         Commercial=case_when(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes") ~ 1,
                              Commercial.FSH==TRUE ~ 1,
                              Reported.Commercial.Exploited.other==TRUE ~ 1,
                              Commercially.Imp.Occurring.IN.Belize==TRUE ~ 1,
                              #matches("ommercial")=="YES" ~ 1,
                              .default=0),
         ) %>%
  select(c(Site,Rookery,
           #Species,
           Common.name,#matches("commerc")
           Small.scale.Fisheries,
           Sport,Commercial
           )) %>% 
  distinct() %>% 
  mutate(n="x") %>% 
  pivot_wider(names_from=c("Site","Rookery"), #output column names 
              names_sort=TRUE,
              values_from="n" #output cell values 
              ) %>% 
  arrange(Common.name)

comm.sp.prune %>% filter(New==TRUE) %>% 
  mutate(Small.scale.Fisheries=case_when(Small.scale.Fisheries==TRUE ~ 1,.default=0),
         Sport=case_when(Gamefish==TRUE ~ 1, .default=0),
         Commercial=case_when(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes") ~ 1,
                              Commercial.FSH==TRUE ~ 1,
                              Reported.Commercial.Exploited.other==TRUE ~ 1,
                              Commercially.Imp.Occurring.IN.Belize==TRUE ~ 1,
                              #matches("ommercial")=="YES" ~ 1,
                              .default=0),
         ) %>%
  select(c(Site,Rookery,
           #Species,
           Common.name,#matches("commerc")
           Small.scale.Fisheries,
           Sport,Commercial
           )) %>% 
  distinct() %>% 
  mutate(n="x") %>% 
  pivot_wider(names_from=c("Site","Rookery"), #output column names 
              names_sort=TRUE,
              values_from="n" #output cell values 
              ) %>% 
  arrange(Common.name)
```

```{r impspp_table}
impspp.table <- comm.sp.prune %>% #comm.sp %>% 
  filter(New==TRUE) %>% 
  mutate(Small.scale.Fisheries=case_when(Small.scale.Fisheries==TRUE ~ 1,.default=0),
         Sport=case_when(Gamefish==TRUE ~ 1, .default=0),
         Commercial=case_when(str_detect(Reported.Commercial.FishbaseSpeciesPage,"Yes") ~ 1,
                              Commercial.FSH==TRUE ~ 1,
                              Reported.Commercial.Exploited.other==TRUE ~ 1,
                              Commercially.Imp.Occurring.IN.Belize==TRUE ~ 1,
                              #matches("ommercial")=="YES" ~ 1,
                              .default=0),
         ) %>%
  select(c(Site,Rookery,
           Species,
           Common.name,#matches("commerc")
           Small.scale.Fisheries,
           Sport,Commercial
           )) %>% 
  distinct() %>% 
  mutate(n="x") %>% 
  pivot_wider(names_from=c("Site","Rookery"), #output column names 
              names_sort=TRUE,
              values_from="n" #output cell values 
              ) %>% 
  arrange(Common.name) %>% 
  mutate(Common.name=paste0(Common.name,", ",Species)) %>% select(-Species)

impspp.table
write.csv(impspp.table,"/Users/lauragivens/Downloads/impspp.table.csv")
```

# Ordination  
```{r}
plot_ordination(ps,ordinate((ps %>% prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)),method="NMDS"),color="Side") + facet_grid(Site~Rookery) + theme_bw()
#ggsave("/Users/lauragivens/Downloads/leeward.windward.nmds.png",width=7,height=5)
```

# Important species  
```{r}
ps.pa.melt %>% 
  group_by(Rookery,Site,Species) %>%
  summarize(Abund=sum(Abundance),spec=unique(Species)) %>% 
  filter(Abund>0) %>% ungroup() %>% 
  group_by(Rookery,Site) %>% 
  summarize(n=length(Species)) %>% pivot_wider(names_from="Rookery",values_from="n") %>% flextable::flextable()
```
```{r}
ps.pa.melt %>% 
  group_by(Month,Rookery,Site,Species) %>%
  summarize(Abund=sum(Abundance),spec=unique(Species)) %>% 
  filter(Abund>0) %>% ungroup() %>% 
  group_by(Month,Rookery,Site) %>% 
  summarize(n=length(Species)) %>% pivot_wider(names_from="Month",values_from="n") %>% arrange(desc(Site)) %>% flextable::flextable()
```

```{r}
hc <- get_taxa_unique(ps.ch %>% subset_samples(Site=="Hicks Caye") %>% prune_taxa(taxa_sums(.)>0,.),'Species')
dc <- get_taxa_unique(ps.ch %>% subset_samples(Site=="Drowned Caye") %>% prune_taxa(taxa_sums(.)>0,.),'Species')
ta <- get_taxa_unique(ps.ch %>% subset_samples(Site=="Turneffe Atoll") %>% prune_taxa(taxa_sums(.)>0,.),'Species')

hc[!hc%in%c(dc,ta)]
dc[!dc%in%c(hc,ta)]
ta[!ta%in%c(hc,dc)]
```

# Questions  
What happened to sportfish that showed up in original report?  
E.g., tarpon, bonefish, and permit
(Megalops atlanticus, Albula vulpes, Trachinotus falcatus)
```{r}
ps %>% subset_taxa(Genus=="Megalops") %>% plot_bar()
ps %>% subset_taxa(Genus=="Trachinotus") %>% plot_bar()
ps %>% subset_taxa(Genus=="Albula") %>% plot_bar()
  
ps %>% subset_taxa(Genus=="Albula") %>% get_taxa_unique("Species")
```


```{r}
sessionInfo()
```

