---
title: "PAMF one pager ocean DNA"
author: "Laura Givens"
date: "2025-06-06"
output: html_document
---
# Load  
Load libraries  
```{r,message=FALSE}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(vegan)
library(lme4)
library(patchwork)
library(ggpattern)
library(ggrepel)
```

```{r}
set.seed(999)
```
```{r}
ggplot2::theme_set(theme_cowplot() + theme(
  legend.position = "bottom",
  panel.border = element_rect(color='black'),
  legend.box.margin = margin(5,5,5,5),
  legend.spacing = unit(0,"pt"),
  plot.title = element_text(face="bold"),
  plot.caption = element_text(face="italic"),
  text = element_text(size=(15),family="Times"),
  legend.key.height = unit(10,"pt"),
  legend.key.width = unit(10,"pt")
))
```

```{r}
load("RData/OceanDNAprop.RData")
```

Load data  
```{r load}
dir_data <- '/Users/lauragivens/Desktop/R/BZrookery_eDNA/Rdata'
dir_results <- "/Volumes/Fuji/Mangroves/2025_0319_Givens_Canty_Rookery_COI/cutadapt/results"

#ps.nt.all <- readRDS(paste0(dir_results,"/prune.decontam.ps.nt.troph.rds")) #no qcov
#ps.mar.all <- readRDS(paste0(dir_results,"/prune.decontam.ps.troph.rds")) #no qcov
ps.nt.all <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.nt.troph.rds"))
ps.mar.all <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.troph.rds"))
ps.bold.all <- readRDS(paste0(dir_results,"/prune.decontam.ps.bold.rds"))

#taxa_nt <- readRDS(paste0(dir_results,"/prune.decontam.ps.nt.troph.rds"))
#taxa_mar <- readRDS(paste0(dir_results,"/taxtable.wtroph.rds"))
#taxa_bold <- readRDS(paste0(dir_results,"/taxtable.bold.rds"))

meta <- readRDS(paste0(dir_results,"/metadata.edit.rds"))
samplelist <- meta %>% filter(Rookery!="Practice" & Rookery !="Blank")

ps.nt.c <- ps.nt.all %>% subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") 
ps.mar.c <- ps.mar.all %>% subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") 
```
Fix randomly unassigned Order 
```{r}
newnttax <- as.data.frame(as.matrix(tax_table(ps.nt.c))) %>% mutate(.,
                      Trophic.Base = word(Trophic.Index,sep=" ") %>% as.numeric(),
                      Trophic.Round = round(as.numeric(Trophic.Base)),
                      .after = Trophic.Index
                      ) 
newnttax$Species[str_detect(newnttax$Species,"sp")] <- NA
newnttax$Species[str_detect(newnttax$Species,"-")] <- NA
newnttax$Species[str_detect(newnttax$Species,"environmental")] <- NA

naordfamlist <- newnttax %>% filter(is.na(Order)) %>% .$Family %>% unique() %>% sort()
naordlist <- c("Mugiliformes", #Ambassidae
               "Carangiformes", #Centropomidae
               "Blenniiformes", #Embiotocidae
               "Acanthuriformes", #Moronidae
               "Blenniiformes", #Plesiopidae
               "Carangiformes", #Polynemidae
               "Acanthuriformes", #Pomacanthidae
               "Blenniiformes", #Pomacentridae
               "Blenniiformes", #Pseudochromidae
               "Acanthuriformes", #Sciaenidae
               "Acanthuriformes", #Sillaginidae
               "Carangiformes", #Sphyraenidae
               "Squatiniformes" #Squatinidae
               )

naordf <- data.frame(fam=naordfamlist,ord=naordlist)

newnttax <- newnttax %>% mutate(Order=case_when(
  is.na(Order) & Family==naordf[1,1] ~ naordf[1,2],
  is.na(Order) & Family==naordf[2,1] ~ naordf[2,2],
  is.na(Order) & Family==naordf[3,1] ~ naordf[3,2],
  is.na(Order) & Family==naordf[4,1] ~ naordf[4,2],
  is.na(Order) & Family==naordf[5,1] ~ naordf[5,2],
  is.na(Order) & Family==naordf[6,1] ~ naordf[6,2],
  is.na(Order) & Family==naordf[7,1] ~ naordf[7,2],
  is.na(Order) & Family==naordf[8,1] ~ naordf[8,2],
  is.na(Order) & Family==naordf[9,1] ~ naordf[9,2],
  is.na(Order) & Family==naordf[10,1] ~ naordf[10,2],
  is.na(Order) & Family==naordf[11,1] ~ naordf[11,2],
  is.na(Order) & Family==naordf[12,1] ~ naordf[12,2],
  is.na(Order) & Family==naordf[13,1] ~ naordf[13,2],
  .default =  Order
  ),
  Class=case_when(
  is.na(Class) & !is.na(Order) ~ "Reptilia",
  .default = Class
  )
)

tax_table(ps.nt.c) <- tax_table(as.matrix(newnttax))
```
```{r}
newmartax <-  as.data.frame(as.matrix(tax_table(ps.mar.c))) %>% mutate(.,
                      Trophic.Base = word(Trophic.Index,sep=" ") %>% as.numeric(),
                      Trophic.Round = round(as.numeric(Trophic.Base)),
                      .after = Trophic.Index
                      ) 
naordfamlist <- newmartax %>% filter(is.na(Order) & !is.na(Family)) %>% .$Family %>% unique() %>% sort()
naordlist <- c("Carangiformes", #Centropomidae
               "Acanthuriformes", #Pomacanthidae
               "Blenniiformes", #Pomacentridae
               "Acanthuriformes", #Sciaenidae
               "Carangiformes" #Sphyraenidae
               )

naordfmar <- data.frame(fam=naordfamlist,ord=naordlist)

newmartax <- newmartax %>% mutate(Order=case_when(
  is.na(Order) & Family==naordfmar[1,1] ~ naordfmar[1,2],
  is.na(Order) & Family==naordfmar[2,1] ~ naordfmar[2,2],
  is.na(Order) & Family==naordfmar[3,1] ~ naordfmar[3,2],
  is.na(Order) & Family==naordfmar[4,1] ~ naordfmar[4,2],
  is.na(Order) & Family==naordfmar[5,1] ~ naordfmar[5,2],
  .default =  Order
  ),
  Class=case_when(
    is.na(Class) & Order=="Testudines" ~ "Reptilia",
    .default = Class
  )
)

tax_table(ps.mar.c) <- tax_table(as.matrix(newmartax))
```

```{r}
ps.nt <- ps.nt.c %>% subset_taxa(Class=="Actinopteri" | Class == "Chondrichthyes") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
ps.mar <- ps.mar.c %>% subset_taxa(Class=="Actinopteri" | Class == "Chondrichthyes") 
ps.bold <- ps.bold.all %>% subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata" & Class=="Actinopterygii" | Class == "Elasmobranchii") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% prune_samples(sample_sums(.)>0,.)
```

```{r}
ps.nt.melt <- psmelt(ps.nt %>% tax_glom("Species",NArm=FALSE))
ps.mar.melt <- psmelt(ps.mar %>% tax_glom("Species",NArm=FALSE))
ps.bold.melt <- psmelt(ps.bold %>% tax_glom("Species",NArm=FALSE))
```

Relative abundance  
```{r}
ps.nt.rel <- ps.nt %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.mar.rel <- ps.mar %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.bold.rel <- ps.bold %>% transform_sample_counts(.,function(x) x/sum(x) ) 
```

```{r}
ps.nt.rel.melt <- psmelt(ps.nt.rel)
ps.mar.rel.melt <- psmelt(ps.mar.rel)
ps.bold.rel.melt <- psmelt(ps.bold.rel)
```

# Basic info  
```{r}
length(get_taxa_unique(ps.nt,"Species"))
length(get_taxa_unique(ps.mar,"Species"))
```

```{r}
length(get_taxa_unique(ps.bold,"Species"))
```

# Test plots  
```{r}
ps.nt.melt %>%
  filter(.,Abundance>0) %>%  group_by(.,Order) %>% 
  reframe(.,Order=unique(Order), 
            db="core nt",
            Abund=sum(Abundance)) %>% 
  mutate(RelAbund=Abund/sum(Abund)) %>%
  ggplot(aes(x=db,y=RelAbund,fill=Order)) + 
  geom_bar(stat="identity",position="stack") + theme(axis.title.x=element_blank(),legend.text = element_text(size=8))
```

```{r}
ps.bold.melt %>%
  filter(.,Abundance>0) %>%  group_by(.,Order) %>% 
  reframe(.,Order=unique(Order), 
          Class=unique(Class),
            db="BOLDv5",
            Abund=sum(Abundance)) %>% 
  mutate(RelAbund=Abund/sum(Abund)) %>%
  ggplot(aes(x=Class,y=Abund,fill=Order)) + 
  geom_bar(stat="identity",position="stack") + theme(axis.title.x=element_blank(),legend.text = element_text(size=8)) + facet_wrap(~Class, scales="free_x")
```

```{r}
ps.nt.melt %>%
  filter(.,Abundance>0) %>% 
  group_by(Class) %>%
  mutate(RelAbund=Abundance/sum(Abundance)) %>% 
  #ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  ggplot(aes(x=Class,y=RelAbund,fill=Order)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  #theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
ps.nt.melt %>%
  filter(.,Abundance>0) %>% 
  group_by(Class) %>%
  mutate(RelAbund=Abundance/sum(Abundance)) %>% 
  #ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  ggplot(aes(x=Class,y=RelAbund,fill=Family)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
ps.nt.melt %>%
  filter(.,Abundance>0) %>% 
  group_by(Class) %>%
  mutate(RelAbund=Abundance/sum(Abundance)) %>% 
  #ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  ggplot(aes(x=Class,y=RelAbund,fill=Genus)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
ps.nt.melt %>%
  filter(.,Abundance>0) %>% 
  group_by(Class) %>%
  mutate(RelAbund=Abundance/sum(Abundance)) %>% 
  #ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  ggplot(aes(x=Class,y=RelAbund,fill=Species)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
ps.bold.melt %>%
  filter(.,Abundance>0) %>% 
  group_by(Class) %>%
  mutate(RelAbund=Abundance/sum(Abundance)) %>% 
  #ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  ggplot(aes(x=Class,y=RelAbund,fill=Species)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
ps.nt.melt %>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Species)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 


ps.bold.melt %>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Species)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
length(get_taxa_unique(ps.bold,"Species"))

length(get_taxa_unique(ps.mar,"Species"))
length(get_taxa_unique(ps.nt,"Species"))
```

```{r}
length(get_taxa_unique(ps.bold,"Genus"))

length(get_taxa_unique(ps.mar,"Genus"))
length(get_taxa_unique(ps.nt,"Genus"))

```

Ok, so we know that the MAR dataset is the species we've definitively identified as being in the MAR.  
This also means each one them was ID'd by some source, so the number of taxa should equal the number of non-NA  
```{r}
ntaxa(ps.mar)
is.na(as.data.frame(as.matrix(tax_table(ps.mar)))$`ID.Source`) %>% summary()

get_taxa_unique(ps.mar,"Species") %>% length()
nrow(distinct(as.data.frame(as.matrix(tax_table(ps.mar)))))
```

Compare this to the nt dataset  
```{r}
ntaxa(ps.nt)
is.na(as.data.frame(as.matrix(tax_table(ps.nt)))$`ID.Source`) %>% summary()

get_taxa_unique(ps.nt,"Species") %>% length()
nrow(distinct(as.data.frame(as.matrix(tax_table(ps.nt)))))

get_taxa_unique((ps.nt %>% subset_taxa(!is.na(`ID.Source`))),"Species") %>% length()
nrow(distinct(as.data.frame(as.matrix(tax_table(ps.nt %>% subset_taxa(!is.na(`ID.Source`)))))))
```
So maybe what we can compare is the fish species we know are present and how many are identified at each level  

```{r}
ps.nt.sub <- ps.nt %>% subset_taxa(!is.na(`ID.Source`))
```
```{r}
ps.nt.sub %>% psmelt() %>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Family)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 

ps.nt.sub %>% psmelt() %>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 

ps.nt.sub %>% psmelt() %>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Species)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```
```{r}
ps.mar.melt%>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Family)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 

ps.mar.melt%>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Genus)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 

ps.mar.melt%>%
  filter(.,Abundance>0) %>% 
  ggplot(aes(x=Class,y=Abundance,fill=Species)) + 
  theme(axis.title.x=element_blank(),legend.text = element_text(size=8))  + 
  theme(legend.position = "none") + 
  geom_bar(stat="identity",position="stack") 
```

```{r}
ps.melt <- ps.nt.sub %>% tax_glom(.,"Species",NArm=FALSE) %>% psmelt()
```
```{r}
ps.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```
```{r}
ps.mar.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```

```{r}
ps.bold.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```

```{r}
ps.nt.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```

```{r}
ps.bold.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% mutate(Rank=factor(Rank,levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa")+ scale_fill_brewer(palette="Dark2")

#ps.bold.melt %>% select(-c(1:2,5:17)) %>% 
#  group_by(Class,Order,Family,Genus,Species) %>% 
#  reframe(Abundance=sum(Abundance)) %>% 
#pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
#  ggplot(aes(y=Rank,x=Abundance,fill=value)) + geom_bar(stat="identity",position="stack") + theme(legend.position="none")
```

```{r}
ps.mar.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% mutate(Rank=factor(Rank,levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa")+ scale_fill_brewer(palette="Dark2")
```

```{r}
ps.nt.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% mutate(Rank=factor(Rank,levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa") + scale_fill_brewer(palette="Dark2")
```

# Comparison figure: fish and elasmobranchs  number of assigned MAR vs BOLD 
```{r}
nassigned.mar.df <- ps.mar.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Family,Genus,Species) %>% 
  #group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Family","Genus","Species"),names_to="Rank") %>% 
  #pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family"))) 
  #mutate(Rank=factor(Rank,levels=c("Class","Order","Family","Genus","Species"))) 
```
```{r,warning=FALSE}
nassigned.mar.plot <- nassigned.mar.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of fish taxa assigned to each rank using a masked GenBank database") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,value,NA))),hjust=2.5) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) + 
  theme(legend.position="right")

nassigned.mar.plot
```

```{r}
nassigned.bold.df <- ps.bold.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Family,Genus,Species) %>% 
  #group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Family","Genus","Species"),names_to="Rank") %>% 
  #pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family"))) 
  #mutate(Rank=factor(Rank,levels=c("Class","Order","Family","Genus","Species"))) 
```
```{r,warning=FALSE}
nassigned.bold.plot <- nassigned.bold.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of fish taxa assigned to each rank using BOLD v5") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=value),hjust=1) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) +
  theme(legend.position="none")

nassigned.bold.plot
```
```{r}
ps.bold.melt %>% select(-c(1:17)) %>% distinct() %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class"))) %>% 
  ggplot(aes(y=Rank,fill=isassgn)) + geom_bar(stat="count") + theme(legend.position="none")
```

```{r}
patchwork::wrap_plots((nassigned.mar.plot+theme(axis.title.x = element_blank())),nassigned.bold.plot,nrow=2)
#ggsave("/Users/lauragivens/Downloads/taxcoverage.family.png",width=10,height=5)
```

# Comparison figure: All in phylum  
```{r}
nassigned.nt.df <- ps.nt.c %>% 
  tax_glom("Species",NArm=FALSE) %>% 
  psmelt() %>% select(-c(1:2,5:16)) %>% 
  group_by(Phylum,Class,Order,Family,Genus,Species) %>% 
  #group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Phylum","Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class","Phylum"))) 
```
```{r}
nassigned.phy.nt.plot <- nassigned.nt.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of chordate taxa assigned to each rank using GenBank") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=value),hjust=0) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,150)) + 
  theme(legend.position="none")
nassigned.phy.nt.plot
```

```{r}
nassigned.boldall.df <- ps.bold.all %>%
  subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") %>% 
  tax_glom("Species",NArm=FALSE) %>% 
  psmelt() %>% select(-c(1:2,5:17)) %>% 
  group_by(Phylum,Class,Order,Family,Genus,Species) %>% 
  #group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Phylum","Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class","Phylum"))) 
```
```{r}
nassigned.phy.bold.plot <- nassigned.boldall.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of chordate taxa assigned to each rank using BOLD v5") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=value),hjust=1) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,200)) + 
  theme(legend.position="none")

nassigned.phy.bold.plot
```

```{r}
nassigned.marall.df <- ps.mar.c %>% 
  tax_glom("Species",NArm=FALSE) %>% 
  psmelt() %>% select(-c(1:2,5:16)) %>% 
  group_by(Phylum,Class,Order,Family,Genus,Species) %>% 
  #group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Phylum","Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class","Phylum"))) 
```
```{r}
nassigned.phy.mar.plot <- nassigned.marall.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of chordate taxa assigned to each rank using a masked GenBank database") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=value),hjust=1) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) + 
  theme(legend.position="right")

nassigned.phy.mar.plot
```

```{r}
patchwork::wrap_plots(
  #(nassigned.mar.plot+theme(axis.title.x = element_blank())),
  (nassigned.phy.nt.plot+theme(axis.title.x = element_blank())),
  (nassigned.phy.mar.plot+theme(axis.title.x = element_blank())),
  (nassigned.phy.bold.plot+theme(legend.position="none")),
  nrow=3) + plot_layout(design="
                        AAAAA
                        BBB##
                        CCC##
                        ")
#ggsave("/Users/lauragivens/Downloads/taxcoverage.phy.png",width=12,height=8)
```

# Trophic Level  
Violin plot
```{r,eval=FALSE,warning=FALSE}
ps.nt.rel.melt %>% 
  mutate(Trophic.Rounded = as.numeric(Trophic.Rounded),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Rounded,
           fill=Site)) + 
  facet_wrap(~Site,nrow=3) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Species' trophic index at rookery and non-rookery mangrove cayes",
       caption="taxa assigned using MAR fish species BLAST alias",
       y="Trophic Index")
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```
```{r,eval=FALSE,warning=FALSE}
ps.mar.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Site)) + 
  facet_wrap(~Site,nrow=3) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Species' trophic index at rookery and non-rookery mangrove cayes",
       caption="taxa assigned using MAR fish species BLAST alias",
       y="Trophic Index")
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```

```{r,warning=FALSE}
#ps.nt %>% psmelt() %>% 
#ps.nt.all  %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")  %>% psmelt() %>% 
violin.nt.plot <- ps.nt.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Rookery)) + 
  facet_grid(Rookery~Site) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Relative abundance of species' trophic index at rookery and non-rookery mangrove cayes",
       subtitle="taxa assigned using NCBI GenBank core nt database",
       #caption="taxa assigned using NCBI GenBank core nt database",
       y="Trophic Index")

violin.nt.plot
#ggsave("Images/violin.relabund.trophic.nt.png",width=10,height=7)
```

```{r,warning=FALSE}
#ps.mar %>% psmelt() %>%
violin.mar.plot <- ps.mar.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Rookery)) + 
  facet_grid(Rookery~Site) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(#title="Species' trophic index at rookery and non-rookery mangrove cayes",
       subtitle="using a masked MAR fish species GenBank BLAST alias",
       #caption="taxa assigned using a masked MAR fish species GenBank BLAST alias",
       y="Trophic Index") + 
  theme(legend.position="right")

violin.mar.plot
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```

```{r,warning=FALSE}
patchwork::wrap_plots(
  (violin.nt.plot+theme(axis.title.x = element_blank(),legend.position="none")),
  violin.mar.plot,
  nrow=2)
#ggsave("/Users/lauragivens/Downloads/violin.nt.comp.png",width=12,height=8)
```

# Check accuracy  
So we can probably assume some of the nt assignments are incorrect  
```{r}
#shared species assignments between nt and MAR masked  
unique(newnttax$Species)[unique(newnttax$Species)%in%unique(newmartax$Species)] %>% length()
```
Make a new dataframe for nt taxa that only include fish and elasmo  
```{r}
newnttaxclass <- as.data.frame(as.matrix(tax_table(ps.nt)))
newboldtaxclass <- as.data.frame(as.matrix(tax_table(ps.bold)))
```
How many species does that bring us down to?  
```{r}
#number of species in original df  
length(unique(newnttax$Species))

#number of fish and elasmo species  
length(unique(newnttaxclass$Species))
length(unique(newboldtaxclass$Species))
```

Get a list of species that are NOT in the MAR masked alias  
```{r}
ntsp <- unique(newnttaxclass$Species)[!unique(newnttaxclass$Species)%in%unique(newmartax$Species)] #subsetting the taxa df we just made to keep only those NOT in the MAR alias dataset  
boldsp <- unique(newboldtaxclass$Species)[!unique(newboldtaxclass$Species)%in%unique(newmartax$Species)]  

#for later: get the list of species that were in the mar dataset  
marsp <- unique(newnttaxclass$Species)[unique(newnttaxclass$Species)%in%unique(newmartax$Species)]
marspbd <- unique(newboldtaxclass$Species)[unique(newboldtaxclass$Species)%in%unique(newmartax$Species)]

# check math 
length(unique(newnttaxclass$Species))
sum(length(marsp),length(ntsp)) #should equal the number of unique species above

length(unique(newboldtaxclass$Species))
sum(length(marspbd),length(boldsp)) #should equal the number of unique species above

#how many species are we left with  
length(ntsp)
length(boldsp)
```
That's still too many to check manually.  

How many species from the nt dataset are NOT in the mar nt dataset and do not share a genus with the mar dataset  
```{r}
ntgen <- newnttaxclass[newnttaxclass$Species%in%ntsp,] %>%  #subset taxa df to keep Species unique to the nt dataset  
  filter(!Genus%in%unique(newmartax$Genus)) %>% #remove genera that are in the mar dataset  
  .$Genus %>% unique() %>% sort()

length(ntgen)
```
```{r}
boldgen <- newboldtaxclass[newboldtaxclass$Species%in%boldsp,] %>%  #subset taxa df to keep Species unique to the bold dataset  
  filter(!Genus%in%unique(newmartax$Genus)) %>% #remove genera that are in the mar dataset  
  .$Genus %>% unique() %>% sort()

length(boldgen)
```

```{r}
spp <- newnttaxclass[newnttaxclass$Species%in%ntsp,] %>% #subset taxa df to keep Species unique to the nt dataset  
  filter(Genus%in%ntgen) %>% #keep only genera that were not in the mar dataset
  distinct() %>% 
  filter(is.na(`ID.Source`)) %>% #remove any that have an ID.Source, because that means we knew the species was in the mar but it just didn't show up in the masked dataset  
  .$Species %>% unique()

sppbd <- newboldtaxclass[newboldtaxclass$Species%in%boldsp,] %>% #subset taxa df to keep Species unique to the bold dataset  
  filter(Genus%in%boldgen) %>% #keep only genera that were not in the mar dataset
  distinct() %>% 
  .$Species %>% unique()


#for later: get the list of species that were in the mar dataset  
sppg <- newnttaxclass[newnttaxclass$Species%in%ntsp,] %>% #subset taxa df to keep Species unique to the nt dataset 
  filter(.,!Genus%in%ntgen) %>% #keep genera that are in the mar dataset  
  .$Species %>% unique()

#for later: get the list of species that were in the mar dataset  
sppgbd <- newboldtaxclass[newboldtaxclass$Species%in%boldsp,] %>% #subset taxa df to keep Species unique to the bold dataset 
  filter(.,!Genus%in%boldgen) %>% #keep genera that are in the mar dataset  
  .$Species %>% unique()

#add to marsp list  
marspp <- c(marsp,sppg)
marsppbd <- c(marspbd,sppgbd)

#check addition  
length(sppg)
length(marsp)
sum(length(sppg),length(marsp))
length(marspp)

sum(length(marspp),length(spp))

sum(length(marsppbd),length(sppbd))
```

```{r}
#remotes::install_github("ropensci/rfishbase")
library("rfishbase")
```
```{r}
fshbsecountrydf <- rfishbase::country(
  species_list = spp,
  #fields = NULL,
  server = c("fishbase", "sealifebase"),
  version = "latest"
  #db = NULL,
)

fshbsecountrydfbd <- rfishbase::country(
  species_list = sppbd,
  #fields = NULL,
  server = c("fishbase", "sealifebase"),
  version = "latest"
  #db = NULL,
)
```

How many species were kept?  
```{r}
length(fshbsecountrydf$Species %>% unique())
length(spp)


length(fshbsecountrydfbd$Species %>% unique())
length(sppbd)
#so some were not found in fishbase  
```
```{r}
#get list of species not found in fishbase  
ntfshbse <- spp[!spp%in%unique(fshbsecountrydf$Species)]
bdfshbse <- sppbd[!sppbd%in%unique(fshbsecountrydfbd$Species)]

#check math  
sum(length(ntfshbse),length(unique(fshbsecountrydf$Species)))
length(spp)

sum(length(bdfshbse),length(unique(fshbsecountrydfbd$Species)))
length(sppbd)

```

What countries are in the dataset?  
```{r}
#fshbsecountrydf$country %>% unique() 
```

Species known to be in the MAR  
```{r}
#df of species in the MAR  
#fshbsecountrydf[fshbsecountrydf$country %in% c("Belize","Mexico","Guatemala","Honduras"),] 

#list of species found in MAR  
fshsppmar <- fshbsecountrydf[fshbsecountrydf$country %in% c("Belize","Mexico","Guatemala","Honduras"),] %>% .$Species %>% unique()
fshsppmarbd <- fshbsecountrydfbd[fshbsecountrydfbd$country %in% c("Belize","Mexico","Guatemala","Honduras"),] %>% .$Species %>% unique()

fshsppmar
fshsppmarbd

#df of fish not found in MAR countries  
fshsppmardf <- fshbsecountrydf %>% filter(!Species%in%fshsppmar)
fshsppmarbddf <- fshbsecountrydfbd %>% filter(!Species%in%fshsppmarbd)

length(fshsppmar)
length(fshsppmardf %>% filter(!Species %in% fshsppmar) %>% .$Species %>% unique())
```
Species known to be in Caribbean  
```{r}
#list of caribbean countries
carcountries <- c("Bel","Mex","Guat","Hond","Baha","Jamai","Arub","Virgin","Cayman","Antil","Antig","Costa","Cuba","Dominic","Barbad","Venez","Nicar","Puert","Guya","Hait","Trini","Surin","United States","Colom","Panam","Grena","Lucia","Vincent","US","Braz","Bras","Cura","Kitts")

#list of non-caribbean countries  
#noncarcountries <- fshsppdf$country[!str_detect(fshsppdf$country,paste(carcountries,collapse="|"))] %>% unique() %>% sort()

#list of species found in caribbean countries 
fshspp <- fshbsecountrydf[str_detect(fshbsecountrydf$country,paste(carcountries,collapse="|")),] %>% .$Species %>% unique() 
fshsppbd <- fshbsecountrydfbd[str_detect(fshbsecountrydfbd$country,paste(carcountries,collapse="|")),] %>% .$Species %>% unique() 

fshspp 
fshsppbd
```
```{r}
#df of fish not found in caribbean countries  
fshsppdf <- fshbsecountrydf %>% filter(!Species%in%fshspp)
fshsppbddf <- fshbsecountrydfbd %>% filter(!Species%in%fshsppbd)
```
Species not recorded in the Caribbean  
```{r}
noncarfsh <- fshsppdf$Species %>% unique() %>% sort()
noncarfshbd <- fshsppbddf$Species %>% unique() %>% sort()

fshsppdf$Species %>% unique() %>% sort()
```

Check numbers  
```{r}
noncarfsh <- noncarfsh[!is.na(noncarfsh)]
fshspp <- fshspp[!is.na(fshspp)]
marspp <- marspp[!is.na(marspp)]
ntfshbse <- ntfshbse[!is.na(ntfshbse)]

length(noncarfsh) # number of fish not found in caribbean
length(fshspp) # number of fish found in caribbean but not mar
length(marspp) # number of fish found in mar  
length(ntfshbse) # number of fish not found in fishbase

sum(length(noncarfsh),length(fshspp),length(marspp),length(ntfshbse))
length(get_taxa_unique(ps.nt,"Species"))
```

Check numbers  
```{r}
noncarfshbd <- noncarfshbd[!is.na(noncarfshbd)]
fshsppbd <- fshsppbd[!is.na(fshsppbd)]
marsppbd <- marsppbd[!is.na(marsppbd)]
bdfshbse <- bdfshbse[!is.na(bdfshbse)]

length(noncarfshbd) # number of fish not found in caribbean
length(fshsppbd) # number of fish found in caribbean but not mar
length(marsppbd) # number of fish found in mar  
length(bdfshbse) # number of fish not found in fishbase

sum(length(noncarfshbd),length(fshsppbd),length(marsppbd),length(bdfshbse))
length(get_taxa_unique(ps.bold,"Species"))
```
### nt plot  
```{r}
accuracy.nt <- ps.nt.melt %>% 
  mutate(FalseAssgn = case_when(Species%in%noncarfsh ~ "Geographically Unlikely", 
                             Species%in%fshspp ~ "Known in Caribbean",
                             Species%in%marspp ~ "Known in MAR",
                             Species%in%ntfshbse ~ "Not in Fishbase",
                             #Genus%in%ntgen & is.na(Species) ~ "Not Defined",
                             is.na(Species) ~ NA
                             ),.after=Species)
```
```{r}
accuracy.plot.df <- accuracy.nt %>% select(-c(1:17)) %>% distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         FalseAssgn=case_when(Rank=="Species" ~ FalseAssgn, 
                              #remove false assignments from everything other than species  
                              .default=NA)
         ) %>% distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn,FalseAssgn) %>% 
  reframe(rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn,
          FalseAssgn=FalseAssgn) %>% 
  distinct() %>% group_by(Rank)  %>% arrange(desc(isassgn),.by_group = TRUE) %>%
  mutate(FinalNum=case_when(isassgn=="Not Assigned" & is.na(FalseAssgn) ~ NotAssgn,
                            isassgn=="Assigned" & is.na(FalseAssgn) ~ rows-NotAssgn,
                            #isassgn=="Not Assigned" & is.na(FalseAssgn) ~ (rows-sum(numuniqassgn)+1),
                            #uniqranks==numuniqassgn ~ rows,
                            .default=numuniqassgn),
        FalseAssgn=factor(FalseAssgn,
                     levels=c(NA,"Not in Fishbase","Geographically Unlikely",
                              "Known in Caribbean","Known in MAR"),ordered=TRUE)
         ) %>% ungroup() %>% group_by(Rank)  %>% arrange(desc(FalseAssgn),.by_group = TRUE) %>% arrange(desc(isassgn),.by_group = TRUE) %>% mutate(,
         LabelNum=( cumsum(FinalNum)))

accuracy.plot.df
```
```{r}
plotpatt <- c("stripe","crosshatch","circle",NA)
plotcol <- c("#e29578","#e9c46a","white","#5a189a")
names(plotpatt) <- names(plotcol) <- levels(accuracy.plot.df$FalseAssgn)




basic.bar <- accuracy.plot.df %>% 
  filter(Rank!="Class" & Rank != "Order") %>% 
  ggplot(.,aes(fill=isassgn)) +
  geom_bar(stat="identity",
           position="stack",
           color="black",
           aes(y=Rank,x=FinalNum,fill=isassgn)) +
  scale_fill_brewer(palette="Dark2") 
basic.bar

bar.patt <- basic.bar + 
  ggpattern::geom_col_pattern(aes(y=Rank,x=FinalNum,
                                  pattern=FalseAssgn,pattern_angle=FalseAssgn,
                                  #pattern_color=FalseAssgn,
                                  pattern_fill=FalseAssgn),
                              color="black",
                              pattern_density=0.7,pattern_spacing=0.02,
                              pattern_alpha=0.8,
                              pattern_key_scale_factor=1
                              )
bar.patt

bar.patt.man <- bar.patt + 
  scale_pattern_manual(values = plotpatt) + 
  scale_pattern_fill_manual(values = plotcol)
bar.patt.man

bar.patt.theme <- bar.patt.man + 
  guides(fill=guide_legend(order=1,
                           title="Taxonomy assigned",
                           override.aes = list(pattern="none")),
         pattern_fill=guide_legend(title="Likely accuracy of assignment",
                                   override.aes = list(pattern_alpha=0.5,
                                                       fill="#1B9E77FF")),
         pattern=guide_legend(title="Likely accuracy of assignment"),
         pattern_angle=guide_legend(title="Likely accuracy of assignment")
         ) + 
  labs(x="Number of fish taxa", 
       title="Number of chordate taxa assigned to each rank \n using the NCBI GenBank core nt database") + 
  scale_x_continuous(expand = c(0,NA),limits=c(-10,125)) + 
  theme(legend.position="right",
        legend.spacing = unit(1,"cm"),
        legend.box.margin = margin(c(1,0,1,1),unit="cm"),
        legend.key.spacing.y = unit(0.1, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.key.width = unit(1, "cm")
        ) 
bar.patt.theme
  
bar.patt.label <- bar.patt.theme +  
  geom_label_repel(aes(x=(LabelNum), y=Rank, label=FinalNum),
                   #position=position_stack(vjust = 0.8),
                   #label.padding = unit(0.15,"line"),
                   nudge_x=(-5),
                   min.segment.length = 0.2,
                   xlim=c(-10,150),
                   size=5,label.size=1,
                   show.legend=FALSE
                   ) 
bar.patt.label
```

```{r}
nassigned.mar.plot2 <- nassigned.mar.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack",color="black") + 
  guides(fill=guide_legend(title="Taxonomy Assigned")) + 
  labs(x="Number of fish taxa", title="Number of fish taxa assigned to each rank \n using only fish known to occur in the MAR") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,value,NA))),hjust=2.5,
                   size=5,label.size=1,show.legend=FALSE) + 
  scale_x_continuous(expand = c(0,NA),limits=c(-10,300)) +  
  theme(legend.position="right",
        legend.spacing = unit(1,"cm"),
        legend.box.margin = margin(c(1,0,1,1),unit="cm"),
        legend.key.spacing.y = unit(0.1, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.key.width = unit(1, "cm")
        ) 

```

```{r}
design1="
AAAA
BBB#
"
wrap_plots(bar.patt.label, (nassigned.mar.plot2)) + 
  plot_layout(design=design1) + plot_annotation(caption='A) Likelihood that a species could be found in the MAR was determined through a search of known countries in FishBase on 11 June 2025 \n B) Species assignments were restricted by taxid of known species',tag_levels='A',tag_sep = ".") & theme(text=element_text(size=20))

#ggsave("/Users/lauragivens/Downloads/taxcov.fish.accuracy.png",width=15,height=9,bg="white")
```

### BOLD v5 plot  
```{r}
accuracy.bold <- ps.bold.melt %>% 
  mutate(FalseAssgn = case_when(Species%in%noncarfshbd ~ "Geographically Unlikely", 
                             Species%in%fshsppbd ~ "Known in Caribbean",
                             Species%in%marsppbd ~ "Known in MAR",
                             Species%in%bdfshbse ~ "Not in Fishbase",
                             #Genus%in%ntgen & is.na(Species) ~ "Not Defined",
                             is.na(Species) ~ NA
                             ),.after=Species)

#sum(length(noncarfshbd),length(fshsppbd),length(marsppbd),length(bdfshbse))
```
```{r}
accuracy.bd.plot.df <- accuracy.bold %>% select(-c(1:17)) %>% distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         FalseAssgn=case_when(Rank=="Species" ~ FalseAssgn, 
                              #remove false assignments from everything other than species  
                              .default=NA)
         ) %>% distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn,FalseAssgn) %>% 
  reframe(rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn,
          FalseAssgn=FalseAssgn) %>% 
  distinct() %>% group_by(Rank)  %>% arrange(desc(isassgn),.by_group = TRUE) %>%
  mutate(FinalNum=case_when(isassgn=="Not Assigned" & is.na(FalseAssgn) ~ NotAssgn,
                            isassgn=="Assigned" & is.na(FalseAssgn) ~ rows-NotAssgn,
                            #isassgn=="Not Assigned" & is.na(FalseAssgn) ~ (rows-sum(numuniqassgn)+1),
                            #uniqranks==numuniqassgn ~ rows,
                            .default=numuniqassgn),
        FalseAssgn=factor(FalseAssgn,
                     levels=c(NA,"Not in Fishbase","Geographically Unlikely",
                              "Known in Caribbean","Known in MAR"),ordered=TRUE)
         ) %>% ungroup() %>% group_by(Rank)  %>% arrange(desc(FalseAssgn),.by_group = TRUE) %>% arrange(desc(isassgn),.by_group = TRUE) %>% mutate(,
         LabelNum=( cumsum(FinalNum)))

```
```{r}
basic.bar.bd <- accuracy.bd.plot.df %>% 
  filter(Rank!="Class" & Rank != "Order") %>% 
  ggplot(.,aes(fill=isassgn)) +
  geom_bar(stat="identity",
           position="stack",
           color="black",
           aes(y=Rank,x=FinalNum,fill=isassgn)) +
  scale_fill_brewer(palette="Dark2") 
basic.bar.bd

bar.bd.patt <- basic.bar.bd + 
  ggpattern::geom_col_pattern(aes(y=Rank,x=FinalNum,
                                  pattern=FalseAssgn,
                                  pattern_angle=FalseAssgn,
                                  #pattern_color=FalseAssgn,
                                  pattern_fill=FalseAssgn),
                              color="black",
                              pattern_density=0.7,pattern_spacing=0.02,
                              pattern_alpha=0.8,
                              pattern_key_scale_factor=1
                              )
bar.bd.patt

bar.bd.patt.man <- bar.bd.patt + 
  scale_pattern_manual(values = plotpatt[names(plotpatt)%in%levels(accuracy.bd.plot.df$FalseAssgn)]
                       ) + 
  scale_pattern_fill_manual(values =  plotcol[names(plotcol)%in%levels(accuracy.bd.plot.df$FalseAssgn)]
                            ) 
bar.bd.patt.man

bar.bd.patt.theme <- bar.bd.patt.man + 
  guides(fill=guide_legend(order=1,
                           title="Taxonomy assigned",
                           override.aes = list(pattern="none")),
         pattern_fill=guide_legend(title="Likely accuracy of assignment",
                                   override.aes = list(pattern_alpha=0.5,
                                                       fill="#1B9E77FF")),
         pattern=guide_legend(title="Likely accuracy of assignment"),
         pattern_angle=guide_legend(title="Likely accuracy of assignment")
         ) + 
  labs(x="Number of fish taxa", 
       title="Number of chordate taxa assigned to each rank \n using the NCBI GenBank core nt database") + 
  scale_x_continuous(expand = c(0,NA),limits=c(-5,140)) + 
  theme(legend.position="right",
        legend.spacing = unit(1,"cm"),
        legend.box.margin = margin(c(1,0,1,1),unit="cm"),
        legend.key.spacing.y = unit(0.1, "cm"),
        legend.key.height = unit(1, "cm"),
        legend.key.width = unit(1, "cm")
        ) 
bar.bd.patt.theme
  
bar.bd.patt.label <- bar.bd.patt.theme +  
  geom_label_repel(aes(x=(LabelNum), y=Rank, label=FinalNum),
                   #position=position_stack(vjust = 0.8),
                   #label.padding = unit(0.15,"line"),
                   nudge_x=(-5),
                   min.segment.length = 0.2,
                   xlim=c(-10,800),
                   size=5,label.size=1,
                   show.legend=FALSE
                   ) 
bar.bd.patt.label
```

```{r,warning=FALSE}
design2="
AAAAA
BBBB#
CCC##
"

patchwork::wrap_plots(bar.patt.label, nassigned.mar.plot2, bar.bd.patt.label,nrow=3,guides="collect") + 
  plot_layout(design=design2) + plot_annotation(caption='A) Likelihood that a species could be found in the MAR was determined through a search of known countries in FishBase on 11 June 2025 \n B) Species assignments were restricted by taxid of known species',tag_levels='A',tag_sep = ".") & theme(text=element_text(size=20))

#ggsave("/Users/lauragivens/Downloads/taxcov.fish.accuracy.png",width=15,height=9,bg="white")
```

# Session Info  
```{r}
save.image("RData/OceanDNAprop.v2.RData")
sessionInfo()
```

