---
title: "PAMF one pager ocean DNA"
author: "Laura Givens"
date: "2025-06-06"
output: html_document
---
# Load  
Load libraries  
```{r,message=FALSE}
library(phyloseq)
library(tidyverse)
library(ggplot2)
library(cowplot)
library(vegan)
library(lme4)
library(patchwork)
library(ggpattern)
library(ggrepel)
```

```{r}
set.seed(999)
```
```{r}
ggplot2::theme_set(theme_cowplot() + theme(
  legend.position = "bottom",
  panel.border = element_rect(color='black'),
  legend.box.margin = margin(5,5,5,5),
  legend.spacing = unit(0,"pt"),
  plot.title = element_text(face="bold"),
  plot.caption = element_text(face="italic"),
  text = element_text(size=(15),family="Times"),
  legend.key.height = unit(10,"pt"),
  legend.key.width = unit(10,"pt")
))
```

```{r,eval=FALSE}
load("RData/OceanDNAprop.RData")
```

Load data  
```{r load}
dir_data <- '/Users/lauragivens/Desktop/R/BZrookery_eDNA/Rdata'
dir_results <- "/Volumes/Fuji/Mangroves/2025_0319_Givens_Canty_Rookery_COI/cutadapt/results"

ps.nt.all2 <- readRDS(paste0(dir_results,"/ps.wordsize.nt.rds")) #qcov=80;wordsize=17? 
#ps.mar.all <- readRDS(paste0(dir_results,"/prune.decontam.ps.troph.rds")) #no qcov
ps.nt.all <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.nt.troph.rds"))
ps.mar.all <- readRDS(paste0(dir_results,"/prune.decontam.v2.ps.troph.rds"))
ps.bold.all <- readRDS(paste0(dir_results,"/prune.decontam.ps.bold.rds"))
ps.comb.all  <- readRDS(paste0(dir_results,"/prune.decontam.combined.wordsize.ps.rds"))

#taxa_nt <- readRDS(paste0(dir_results,"/prune.decontam.ps.nt.troph.rds"))
#taxa_mar <- readRDS(paste0(dir_results,"/taxtable.wtroph.rds"))
#taxa_bold <- readRDS(paste0(dir_results,"/taxtable.bold.rds"))

meta <- readRDS(paste0(dir_results,"/metadata.edit.rds"))
samplelist <- meta %>% filter(Rookery!="Practice" & Rookery !="Blank")

ps.nt.c <- ps.nt.all %>% subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") 
ps.mar.c <- ps.mar.all %>% subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") 
ps.comb.c <- ps.comb.all %>% subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") 
```
Fix randomly unassigned Order 
```{r}
newnttax <- as.data.frame(as.matrix(tax_table(ps.nt.c))) %>% mutate(.,
                      Trophic.Base = word(Trophic.Index,sep=" ") %>% as.numeric(),
                      Trophic.Round = round(as.numeric(Trophic.Base)),
                      .after = Trophic.Index
                      ) 
newnttax$Species[str_detect(newnttax$Species,"sp")] <- NA
newnttax$Species[str_detect(newnttax$Species,"-")] <- NA
newnttax$Species[str_detect(newnttax$Species,"environmental")] <- NA

naordfamlist <- newnttax %>% 
  filter(is.na(Order)) %>% #for which assignments is Order blank
  .$Family %>% #get the family assignments for those taxa  
  unique() %>% sort()
naordfamlist 
#"Centropomidae" "Moronidae"     "Pomacanthidae" "Pomacentridae" "Sciaenidae"    "Sphyraenidae" 

# from this list, manually assign the correct order name  
naordlist <- c(#"Mugiliformes", #Ambassidae
               "Carangiformes", #Centropomidae
               #"Blenniiformes", #Embiotocidae
               "Acanthuriformes", #Moronidae
               #"Blenniiformes", #Plesiopidae
               #"Carangiformes", #Polynemidae
               "Acanthuriformes", #Pomacanthidae
               "Blenniiformes", #Pomacentridae
               #"Blenniiformes", #Pseudochromidae
               "Acanthuriformes", #Sciaenidae
               #"Acanthuriformes", #Sillaginidae
               "Carangiformes" #Sphyraenidae
               #"Squatiniformes" #Squatinidae
               )

naordf <- data.frame(fam=naordfamlist,ord=naordlist)
naordf #check to make sure they match up 

newnttax <- newnttax %>% mutate(Order=case_when(
  is.na(Order) & Family==naordf[1,1] ~ naordf[1,2],
  is.na(Order) & Family==naordf[2,1] ~ naordf[2,2],
  is.na(Order) & Family==naordf[3,1] ~ naordf[3,2],
  is.na(Order) & Family==naordf[4,1] ~ naordf[4,2],
  is.na(Order) & Family==naordf[5,1] ~ naordf[5,2],
  is.na(Order) & Family==naordf[6,1] ~ naordf[6,2],
  .default =  Order
  ),
  #Class=case_when(
  #is.na(Class) & !is.na(Order) ~ "Reptilia",
  #.default = Class
  #)
)

tax_table(ps.nt.c) <- tax_table(as.matrix(newnttax))
```
```{r}
newmartax <-  as.data.frame(as.matrix(tax_table(ps.mar.c))) %>% mutate(.,
                      Trophic.Base = word(Trophic.Index,sep=" ") %>% as.numeric(),
                      Trophic.Round = round(as.numeric(Trophic.Base)),
                      .after = Trophic.Index
                      ) 
naordfamlist <- newmartax %>% 
  filter(is.na(Order) & !is.na(Family)) %>% 
  .$Family %>% 
  unique() %>% sort()
naordfamlist
# "Centropomidae" "Pomacanthidae" "Pomacentridae" "Sphyraenidae" 

naordlist <- c("Carangiformes", #Centropomidae
               "Acanthuriformes", #Pomacanthidae
               "Blenniiformes", #Pomacentridae
               #"Acanthuriformes", #Sciaenidae
               "Carangiformes" #Sphyraenidae
               )

naordfmar <- data.frame(fam=naordfamlist,ord=naordlist)

newmartax <- newmartax %>% mutate(Order=case_when(
  is.na(Order) & Family==naordfmar[1,1] ~ naordfmar[1,2],
  is.na(Order) & Family==naordfmar[2,1] ~ naordfmar[2,2],
  is.na(Order) & Family==naordfmar[3,1] ~ naordfmar[3,2],
  is.na(Order) & Family==naordfmar[4,1] ~ naordfmar[4,2],
  #is.na(Order) & Family==naordfmar[5,1] ~ naordfmar[5,2],
  .default =  Order
  ),
  #Class=case_when(
  #  is.na(Class) & Order=="Testudines" ~ "Reptilia",
  #  .default = Class
  #)
)

tax_table(ps.mar.c) <- tax_table(as.matrix(newmartax))
```
```{r}
newcombtax <-  as.data.frame(as.matrix(tax_table(ps.comb.c))) %>% mutate(.,
                      Trophic.Base = word(Trophic.Index,sep=" ") %>% as.numeric(),
                      Trophic.Round = round(as.numeric(Trophic.Base)),
                      .after = Trophic.Index
                      ) 
naordfamlist <- newcombtax %>%
  filter(is.na(Order) & !is.na(Family)) %>%
  .$Family %>%
  unique() %>% sort()
naordfamlist
#"Centropomidae" 

naordlist <- c("Carangiformes" #Centropomidae
               )

naordfmar <- data.frame(fam=naordfamlist,ord=naordlist)

newcombtax <- newcombtax %>% mutate(Order=case_when(
  is.na(Order) & Family==naordfmar[1,1] ~ naordfmar[1,2],
  .default =  Order
  )
)

tax_table(ps.comb.c) <- tax_table(as.matrix(newcombtax))
```

```{r}
ps.nt <- ps.nt.c %>% 
  #subset_taxa(Class=="Actinopteri" | Class == "Chondrichthyes") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% 
  prune_samples(sample_sums(.)>0,.)

ps.mar <- ps.mar.c %>% 
  #subset_taxa(Class=="Actinopteri" | Class == "Chondrichthyes")  %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% 
  prune_samples(sample_sums(.)>0,.)

ps.bold <- ps.bold.all %>% 
  subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") %>% 
  #subset_taxa(Phylum=="Chordata" & Class=="Actinopterygii" | Class == "Elasmobranchii") %>% 
  prune_taxa(taxa_sums(.)>0,.) %>%
  prune_samples(sample_sums(.)>0,.)

ps.comb <- ps.comb.c %>% 
  subset_samples(Rookery!="Practice" & Rookery !="Blank") %>% 
  subset_taxa(Phylum=="Chordata") %>% 
  #subset_taxa(Phylum=="Chordata" & Class=="Actinopterygii" | Class == "Elasmobranchii" | Class=="Actinopterygii" ) %>% 
  prune_taxa(taxa_sums(.)>0,.) %>% 
  prune_samples(sample_sums(.)>0,.)
```

```{r}
ps.nt.melt <- psmelt(ps.nt )
ps.mar.melt <- psmelt(ps.mar )
ps.bold.melt <- psmelt(ps.bold)

ps.comb.melt <- psmelt(ps.comb )
```

Relative abundance  
```{r}
ps.nt.rel <- ps.nt %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.mar.rel <- ps.mar %>% transform_sample_counts(.,function(x) x/sum(x) ) 
ps.bold.rel <- ps.bold %>% transform_sample_counts(.,function(x) x/sum(x) ) 

ps.comb.rel <- ps.comb %>% transform_sample_counts(.,function(x) x/sum(x) ) 
```

```{r}
ps.nt.rel.melt <- psmelt(ps.nt.rel)
ps.mar.rel.melt <- psmelt(ps.mar.rel)
ps.bold.rel.melt <- psmelt(ps.bold.rel)

ps.comb.rel.melt <- psmelt(ps.comb.rel)
```

# Basic info  
This is Chordates (all chordates), so also including Aves, Ascidiacea, and Mammalia   

How many including higher level taxa?  
```{r}
ntaxa(ps.mar)
ntaxa(ps.nt)
ntaxa(ps.bold)
ntaxa(ps.comb)
```

How many species are in each database?  
```{r}
length(get_taxa_unique(ps.nt,"Species")) #core nt db
length(get_taxa_unique(ps.mar,"Species")) #mar masked db
length(get_taxa_unique(ps.bold,"Species")) #bold db

length(get_taxa_unique(ps.comb,"Species")) #combined bold and core nt db
```

Ok, so we know that the MAR dataset is the species we've definitively identified as being in the MAR.  
This also means each one them was ID'd by some source, so the number of taxa should equal the number of non-NA  
```{r}
# comparing the number of taxa in the MAR dataset 
# to the number of taxa without a specified ID source  

ntaxa(ps.mar) #total number of taxa 

is.na( #number of taxa without an ID source  
  as.data.frame(as.matrix(
    tax_table(ps.mar)))$`ID.Source`) %>% 
  summary() #false = ID source available (non-NA)
```

Compare this to the nt dataset  
```{r}
ntaxa(ps.nt)
is.na(
  as.data.frame(as.matrix(
    tax_table(ps.nt)))$`ID.Source`) %>%
  summary()
```

and the combined dataset
```{r}
ntaxa(ps.comb)
is.na(
  as.data.frame(as.matrix(
    tax_table(ps.comb)))$`ID.Source`) %>% 
  summary()
```

```{r}
ps.nt.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()

ps.mar.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()

ps.bold.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()

ps.comb.melt %>% filter(Abundance > 0) %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```

Compare assignments between the different types  
```{r}
ps.mar.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>%
  reframe("Assigned"=sum(!is.na(value)),
           "Not Assigned"=sum(is.na(value))) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% mutate(Rank=factor(Rank, levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + 
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa with MAR masked database") +
  scale_fill_brewer(palette="Dark2") +
scale_x_continuous(limits=c(0,250))


ps.nt.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))) %>% 
  #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% mutate(Rank=factor(Rank, levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + 
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa with NCBI nt database") + 
  scale_fill_brewer(palette="Dark2")+
scale_x_continuous(limits=c(0,250))


ps.bold.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))) %>% 
  #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% mutate(Rank=factor(Rank, levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) +
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa with BOLD v5") +
  scale_fill_brewer(palette="Dark2")+
scale_x_continuous(limits=c(0,250))


ps.comb.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))) %>% 
  #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% mutate(Rank=factor(Rank, levels=c("Class","Order","Family","Genus","Species"))) %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + 
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa with BOLD v5 and NCBI nt") + 
  scale_fill_brewer(palette="Dark2") +
scale_x_continuous(limits=c(0,300))
```

```{r}
n.asv.plot <- tax_table(ps.comb) %>% 
  as.matrix() %>% as.data.frame() %>% 
  select(c(#Domain,Phylum,
           Class,Order,Family,Genus,Species)) %>%
  mutate(Class=case_when(Class=="Actinopterygii" ~ "Actinopteri",.default=Class)) %>%
 # summarize(is.na())
 pivot_longer(cols=c(#Domain,Phylum,
                     Class,Order,Family,Genus,Species)) %>%
 mutate(val=case_when(is.na(value) ~ NA,
                      .default="Assigned"),
         sub=case_when(name=="Species" & !is.na(value) ~ "Assigned",
                       name=="Species" & is.na(value) ~ NA,
                       name=="Genus" & is.na(value) ~ NA,
                       name=="Genus" & !is.na(value) ~ "Assigned",
                       name=="Family" & is.na(value) ~ NA,
                       name=="Family" & !is.na(value) ~ "Assigned",
                       name=="Order" & is.na(value) ~ NA,
                       name=="Order" & !is.na(value) ~ "Assigned",
                      .default=value),
         name=factor(name,levels=c("Species","Genus","Family","Order","Class","Phylum","Domain"))) %>%
  mutate(sub=factor(sub,levels=c("Actinopteri","Amphibia","Aves","Ascidiacea","Elasmobranchii","Reptilia", "Mammalia","Assigned" ,NA ))) %>%
  #ggplot(aes(x=name,fill=val)) +
  ggplot(aes(y=name,fill=sub)) +
  geom_bar(stat="count",color="black") + 
  scale_x_continuous(expand=c(0,NA),
                     limits=c(0,700),
                     breaks=seq(from=0,to=700,by=100)) + 
  ggpubr::theme_pubclean() +
  scale_fill_brewer(name="Assigned taxa",palette="Set3") +
  #theme(legend.position="none") +
  labs(x="Number of chordate taxa",y="") + theme(legend.position="right")

n.asv.plot 
```

# Comparison figure: fish and elasmobranchs  number of assigned MAR vs BOLD 
```{r}
nassigned.mar.df <- ps.mar.melt %>% 
  select(-c(1:2,5:17)) %>% 
  group_by(Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))) %>% 
  #mutate(Rel=Abundance/sum(Abundance)) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% 
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family"))) %>% ungroup() %>%
  arrange(Rank,desc(val)) %>% 
  mutate(val=factor(val,levels=c("Assigned","Not Assigned")))
```
```{r,warning=FALSE}
nassigned.mar.plot <- nassigned.mar.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + 
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", 
       title="Number of fish taxa assigned to each rank using a masked GenBank database") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,
                               value,NA))),
             hjust=2.5) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) + 
  theme(legend.position="right")

nassigned.mar.plot
```

```{r}
nassigned.bold.df <- ps.bold.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% 
  mutate(Rank=factor(Rank,
                     levels=c("Species","Genus","Family"))) 
```
```{r,warning=FALSE}
nassigned.bold.plot <- nassigned.bold.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of fish taxa assigned to each rank using BOLD v5") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,
                               value,NA))),
             hjust=2) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) +
  theme(legend.position="none")

nassigned.bold.plot
```

```{r}
nassigned.comb.df <- ps.comb.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% 
  mutate(Rank=factor(Rank,
                     levels=c("Species","Genus","Family"))) 
```
```{r,warning=FALSE}
nassigned.comb.plot <- nassigned.comb.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of fish taxa assigned to each rank using agreements between BOLD and NCBI") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,
                               value,NA))),
             hjust=2) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) +
  theme(legend.position="none")

nassigned.comb.plot
```

```{r}
patchwork::wrap_plots(
  (nassigned.mar.plot + 
     scale_x_continuous(limits=c(0,275))
   ),
  (nassigned.bold.plot + 
     scale_x_continuous(limits=c(0,275))
   ),
  (nassigned.comb.plot + 
     scale_x_continuous(limits=c(0,275))
   ),
  nrow=3)
#ggsave("/Users/lauragivens/Downloads/taxcoverage.family.png",width=10,height=5)
```

# Comparison figure: All in phylum  
```{r}
nassigned.nt.df <-   ps.nt.melt %>% 
  select(-c(1:2,5:16)) %>% 
  group_by(Phylum,Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Phylum","Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value))
          )  %>%
  pivot_longer(cols=c("Not Assigned","Assigned"),
               names_to="val") %>% 
  group_by(Rank) %>% mutate(pos=case_when(val=="Not Assigned" ~ value+1, .default=sum(value))) %>% ungroup() %>%
  mutate(Rank=factor(Rank,
                     levels=c("Species","Genus","Family","Order","Class","Phylum"))) 
```
```{r}
nassigned.phy.nt.plot <- nassigned.nt.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + 
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", 
       title="Number of chordate taxa assigned to each rank using GenBank") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,
                               value,NA)),
                 x=pos
                 ),
             hjust=1) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,150)) + 
  theme(legend.position="none")
nassigned.phy.nt.plot
```

```{r}
nassigned.boldall.df <- ps.bold.melt %>% select(-c(1:2,5:17)) %>% 
  group_by(Phylum,Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Phylum","Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  group_by(Rank) %>% mutate(pos=case_when(val=="Not Assigned" ~ value+1, .default=sum(value))) %>% ungroup() %>%
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class","Phylum"))) 
```
```{r}
nassigned.phy.bold.plot <- nassigned.boldall.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + 
  geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", 
       title="Number of chordate taxa assigned to each rank using BOLD v5") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,
                               value,NA)),
                 x=pos
                 ),
             hjust=1) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,200)) + 
  theme(legend.position="none")

nassigned.phy.bold.plot
```

```{r}
nassigned.marall.df <- ps.mar.melt %>% select(-c(1:2,5:16)) %>% 
  group_by(Phylum,Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Phylum","Class","Order","Family","Genus","Species"),names_to="Rank") %>% 
  group_by(Rank) %>% reframe("Assigned"=sum(!is.na(value)),
                             "Not Assigned"=sum(is.na(value))) %>% 
  pivot_longer(cols=c("Not Assigned","Assigned"),names_to="val") %>% 
  group_by(Rank) %>% mutate(pos=case_when(val=="Not Assigned" ~ value+1, .default=sum(value))) %>% ungroup() %>%
  mutate(Rank=factor(Rank,levels=c("Species","Genus","Family","Order","Class","Phylum"))) 
```
```{r}
nassigned.phy.mar.plot <- nassigned.marall.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) + geom_bar(stat="identity",position="stack") + 
  guides(fill=guide_legend(title="Assigned")) + 
  labs(x="Number of assigned taxa", title="Number of chordate taxa assigned to each rank using a masked GenBank database") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,
                               value,NA)),
                 x=pos
                 ),
             hjust=1) + 
  #scale_x_continuous(expand = c(0,NA),limits=c(0,100)) + 
  theme(legend.position="right")

nassigned.phy.mar.plot
```

```{r}
patchwork::wrap_plots(
  #(nassigned.mar.plot+theme(axis.title.x = element_blank())),
  (nassigned.phy.nt.plot+theme(axis.title.x = element_blank())),
  (nassigned.phy.mar.plot+theme(axis.title.x = element_blank())),
  (nassigned.phy.bold.plot+theme(legend.position="none")),
  nrow=3) + plot_layout(design="
                        CCCCC
                        AAAA#
                        BBB##
                        ")
#ggsave("/Users/lauragivens/Downloads/taxcoverage.phy.png",width=12,height=8)
```

# Trophic Level  
Violin plot
```{r,warning=FALSE}
#ps.nt %>% psmelt() %>% 
#ps.nt.all  %>% subset_samples(Rookery!="Practice" & Rookery !="Blank")  %>% psmelt() %>% 
violin.nt.plot <- ps.nt.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Rookery)) + 
  facet_grid(Rookery~Site) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(title="Relative abundance of species' trophic index at rookery and non-rookery mangrove cayes",
       subtitle="taxa assigned using NCBI GenBank core nt database",
       #caption="taxa assigned using NCBI GenBank core nt database",
       y="Trophic Index")

violin.nt.plot
#ggsave("Images/violin.relabund.trophic.nt.png",width=10,height=7)
```

```{r,warning=FALSE}
#ps.mar %>% psmelt() %>%
violin.mar.plot <- ps.mar.rel.melt %>% 
  mutate(Trophic.Base = as.numeric(Trophic.Base),
         Month=factor(Month,levels=c("August","November","February","May"))) %>% 
  filter(Abundance>0) %>%
ggplot(.,
       aes(x=Month,
           y=Trophic.Base,
           fill=Rookery)) + 
  facet_grid(Rookery~Site) + 
  geom_violin() + 
  #geom_jitter(shape=16, position=position_jitter(0.2)) + 
  geom_boxplot(width=0.15) +
 scale_color_viridis_d(begin=0.3,end=0.9,option='plasma') + scale_fill_viridis_d(begin=0.3,end=0.9,option='plasma') +
  labs(#title="Species' trophic index at rookery and non-rookery mangrove cayes",
       subtitle="using a masked MAR fish species GenBank BLAST alias",
       #caption="taxa assigned using a masked MAR fish species GenBank BLAST alias",
       y="Trophic Index") + 
  theme(legend.position="right")

violin.mar.plot
#ggsave("Images/violin.relabund.trophic.mar.png",width=10,height=7)
```

```{r,warning=FALSE}
patchwork::wrap_plots(
  (violin.nt.plot+theme(axis.title.x = element_blank(),legend.position="none")),
  violin.mar.plot,
  nrow=2)
#ggsave("/Users/lauragivens/Downloads/violin.nt.comp.png",width=12,height=8)
```

# Check accuracy   
```{r}
library(rfishbase)
```
```{r}
#list of caribbean countries
carcountries <- c(#"Bel","Mex","Guat","Hond",
                  "Baha","Jamai","Arub","Virgin","Cayman","Antil","Antig","Costa","Cuba","Dominic","Barbad","Venez","Nicar","Puert","Guya","Hait","Trini","Surin","United States","Colom","Panam","Grena","Lucia","Vincent","US","Braz","Bras","Cura","Kitts")
```

We want to know:   
1. how many taxa are there  
2. how many are actually known to be in the MAR  
3. how many are potentially in the MAR  
4. how many are wrong?   

So to do that, we can first use the MAR masked database to check what species are in common. Any species shared with the MAR dataset are likely to be correct or at least actually known to be present in the region 

Make a new dataframe that only include fish and elasmo  
```{r fishelasmodf}
newnttaxclass <- as.data.frame(as.matrix(
  tax_table(ps.nt %>%
              subset_taxa(Class!="Aves" & Class != "Ascidiacea" & Class != "Mammalia") 
            )))
newnt2taxclass <- as.data.frame(as.matrix(
  tax_table(ps.nt.all2 %>%
              subset_taxa(Class=="Actinopteri") 
            )))

newbdtaxclass <- as.data.frame(as.matrix(
  tax_table(ps.bold %>% 
              subset_taxa(Class!="Aves" & Class != "Ascidiacea" & Class != "Reptilia" & Class != "Mammalia"))
  ))

newcbtaxclass <- as.data.frame(as.matrix(
  tax_table(ps.comb %>% 
              subset_taxa(Class!="Aves" & Class != "Ascidiacea" & Class != "Reptilia" & Class != "Amphibia" & Class != "Mammalia" & !is.na(Class)) 
            )))
```

### NCBI database   
Using the new dataframe of only fish and elasmo species,  
Get a list of species that are NOT in the MAR masked alias    

ntsp: by subsetting the taxa df we just made to keep only those NOT in the MAR alias dataset   
ntmarsp: Then get the list of species that ARE in the MAR dataset  

After querying fishbase using the ntsp list, we will subdivide the species into fish:  
1. known to occur in the MAR (ntmarsp + ntfshbse_mar)  
2. fish from the Caribbean (ntfshbse_car)  
3. fish that are not known to be in the region/found outside the region (ntfshbse_out)  
4. fish that were not found in the databse (ntfshbse_notfound)

```{r}
ntsp <- unique(newnttaxclass$Species)[!unique(newnttaxclass$Species)%in%unique(newmartax$Species)] #species not in mar db
nt2sp <- unique(newnt2taxclass$Species)[!unique(newnt2taxclass$Species)%in%unique(newmartax$Species)]
ntmarsp <- unique(newnttaxclass$Species)[unique(newnttaxclass$Species)%in%unique(newmartax$Species)] #species in mar db
nt2marsp <- unique(newnt2taxclass$Species)[unique(newnt2taxclass$Species)%in%unique(newmartax$Species)] 
```

check that those numbers add up    
```{r}
#nt
length(unique(newnttaxclass$Species)) #total = 
length(ntsp) # taxa not in database +
length(ntmarsp) # taxa in database
# +- 1 for NA 
```

Query fishbase  
Use list of species not in the MAR masked alias database (ntsp)  
```{r}
fshbsecountrydf <- rfishbase::country(
  species_list = ntsp,#spp,
  #fields = NULL,
  server = c("fishbase", "sealifebase"),
  version = "latest"
  #db = NULL,
)

fshbsecountrydf2 <- rfishbase::country(
  species_list = nt2sp,#spp,
  #fields = NULL,
  server = c("fishbase", "sealifebase"),
  version = "latest"
  #db = NULL,
)
```

Get lists of species from fishbase output  
ntfshbse_notfound: species not found in fishbase  
ntfshbse_mar: species found in the MAR  
ntfshbse_car: species not found in the MAR but found in the Caribbean  
ntfshbse_out: species not recorded in the Caribbean 
```{r}
#list of species not found in fishbase  
ntfshbse_notfound <- ntsp[!ntsp%in%unique(fshbsecountrydf$Species)]
nt2fshbse_notfound <- nt2sp[!nt2sp%in%unique(fshbsecountrydf2$Species)]

#list of species found in MAR  
ntfshbse_mar <- fshbsecountrydf[fshbsecountrydf$country %in% c("Belize","Mexico","Guatemala","Honduras"),] %>% .$Species %>%  unique()
nt2fshbse_mar <- fshbsecountrydf2[fshbsecountrydf2$country %in% c("Belize","Mexico","Guatemala","Honduras"),] %>% .$Species %>%  unique()

#list of species found in caribbean countries 
ntfshbse_car <- fshbsecountrydf[str_detect(fshbsecountrydf$country,
                                     paste(carcountries,collapse="|")),] %>% 
  filter(!Species%in%ntfshbse_mar) %>% #dont include species already identified as being in a MAR country
  .$Species %>% 
  unique() 
nt2fshbse_car <- fshbsecountrydf2[str_detect(fshbsecountrydf2$country,
                                     paste(carcountries,collapse="|")),] %>% 
  filter(!Species%in%nt2fshbse_mar) %>% #dont include species already identified as being in a MAR country
  .$Species %>% 
  unique() 

#Species not recorded in the Caribbean  
ntfshbse_out <- fshbsecountrydf %>% 
  filter(!Species%in%ntfshbse_mar) %>% #don't include species found in MAR countries
  filter(!Species%in%ntfshbse_car) %>% #don't include species found in caribbean countries 
  .$Species %>% 
  unique() %>%
  sort()
nt2fshbse_out <- fshbsecountrydf2 %>% 
  filter(!Species%in%nt2fshbse_mar) %>% #don't include species found in MAR countries
  filter(!Species%in%nt2fshbse_car) %>% #don't include species found in caribbean countries 
  .$Species %>% 
  unique() %>%
  sort()
```

Confirm numbers add up  
```{r}
#total num species
length(unique(newnttaxclass$Species))
sum(length(ntmarsp),length(ntsp)) #sum of species found in MAR db and those not found in MAR db  

#num species queried to fishbase  
length(ntsp) #ntsp was used to query fishbase  
sum(length(ntfshbse_notfound), #species not found in fishbase  
    length(ntfshbse_mar), #species found in MAR
    length(ntfshbse_car), #species found in Caribbean
    length(ntfshbse_out) #species found outside of Caribbean  
    )

#total num species 
length(unique(newnttaxclass$Species))
sum(length(ntmarsp), #species found in MAR database
    length(ntfshbse_notfound), #species not found in fishbase  
    length(ntfshbse_mar), #species found in MAR
    length(ntfshbse_car), #species found in Caribbean
    length(ntfshbse_out) #species found outside of Caribbean  
    )
```
```{r}
ntmarsp <-  ntmarsp[!is.na(ntmarsp)] #species found in MAR database
ntfshbse_notfound <- ntfshbse_notfound[!is.na(ntfshbse_notfound)] #species not found in fishbase  
ntfshbse_mar <- ntfshbse_mar[!is.na(ntfshbse_mar)] #species found in MAR
ntfshbse_car <- ntfshbse_car[!is.na(ntfshbse_car)] #species found in Caribbean
ntfshbse_out <- ntfshbse_out[!is.na(ntfshbse_out)] #species found outside of Caribbean  
```

```{r}
accuracy.nt <- ps.nt.melt %>% 
  filter(Class=="Actinopteri" | Class=="Chondrichthyes") %>%
  mutate(FalseAssgn = case_when(Species%in%ntmarsp ~ "Known in MAR",
                                Species%in%ntfshbse_mar ~ "Known in MAR",
                                Species%in%ntfshbse_car ~ "Known in Caribbean",
                                Species%in%ntfshbse_out ~ "Geographically Unlikely", 
                             Species%in%ntfshbse_notfound ~ "Not in Fishbase",
                             is.na(Species) ~ NA,
                             .default="Not a fish species"
                             ),.after=Species)

accuracy.nt2 <- ps.nt.all2 %>% 
  subset_taxa(Class=="Actinopteri") %>% psmelt() %>%
  mutate(FalseAssgn = case_when(Species%in%nt2marsp ~ "Known in MAR",
                                Species%in%nt2fshbse_mar ~ "Known in MAR",
                                Species%in%nt2fshbse_car ~ "Known in Caribbean",
                                Species%in%nt2fshbse_out ~ "Geographically Unlikely", 
                             Species%in%nt2fshbse_notfound ~ "Not in Fishbase",
                             is.na(Species) ~ NA,
                             .default="Not a fish species"
                             ),.after=Species) 
```
```{r}
accuracy.nt.plot.df <- accuracy.nt %>% 
  select(-c(1:17)) %>% 
  distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
               names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,
                     levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         FalseAssgn=case_when(Rank=="Species" ~ FalseAssgn, 
                              #remove false assignments from everything other than species  
                              .default=NA)
         ) %>% 
  distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn,FalseAssgn) %>% 
  reframe(rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn,
          FalseAssgn=FalseAssgn) %>% 
  distinct() %>% 
  group_by(Rank) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(FinalNum=case_when(isassgn=="Not Assigned" & is.na(FalseAssgn) ~ NotAssgn,
                            #isassgn=="Assigned" & is.na(FalseAssgn) ~ rows-NotAssgn,
                            isassgn=="Assigned" & is.na(FalseAssgn) ~ uniqranks-NotAssgn,
                            #isassgn=="Not Assigned" & is.na(FalseAssgn) ~ (rows-sum(numuniqassgn)+1),
                            #uniqranks==numuniqassgn ~ rows,
                            .default=numuniqassgn),
        FalseAssgn=factor(FalseAssgn,
                     levels=c("Not a fish species",
                              "Not in Fishbase",
                              "Geographically Unlikely",
                              "Known in Caribbean","Known in MAR"),ordered=TRUE)
         ) %>% 
  ungroup() %>% 
  group_by(Rank) %>% 
  arrange(desc(FalseAssgn),
          .by_group = TRUE) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(,
         LabelNum=(cumsum(FinalNum)))


accuracy.nt2.plot.df <- accuracy.nt2 %>% select(-c(1:17)) %>% 
  distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
               names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,
                     levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         FalseAssgn=case_when(Rank=="Species" ~ FalseAssgn, 
                              #remove false assignments from everything other than species  
                              .default=NA)
         ) %>% 
  distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn,FalseAssgn) %>% 
  reframe(rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn,
          FalseAssgn=FalseAssgn) %>% 
  distinct() %>% 
  group_by(Rank) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(FinalNum=case_when(isassgn=="Not Assigned" & is.na(FalseAssgn) ~ NotAssgn,
                            #isassgn=="Assigned" & is.na(FalseAssgn) ~ rows-NotAssgn,
                            isassgn=="Assigned" & is.na(FalseAssgn) ~ uniqranks-NotAssgn,
                            #isassgn=="Not Assigned" & is.na(FalseAssgn) ~ (rows-sum(numuniqassgn)+1),
                            #uniqranks==numuniqassgn ~ rows,
                            .default=numuniqassgn),
        FalseAssgn=factor(FalseAssgn,
                     levels=c("Not a fish species",
                              "Not in Fishbase",
                              "Geographically Unlikely",
                              "Known in Caribbean","Known in MAR"),ordered=TRUE)
         ) %>% 
  ungroup() %>% 
  group_by(Rank) %>% 
  arrange(desc(FalseAssgn),
          .by_group = TRUE) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(,
         LabelNum=(cumsum(FinalNum)))
```

### BOLD database   
Using the new dataframe of only fish and elasmo species,  
Get a list of species that are NOT in the MAR masked alias    

bdsp: by subsetting the taxa df we just made to keep only those NOT in the MAR alias dataset   
bdmarsp: Then get the list of species that ARE in the MAR dataset  

After querying fishbase using the bdsp list, we will subdivide the species into fish:  
1. known to occur in the MAR (bdmarsp + bdfshbse_mar)  
2. fish from the Caribbean (bdfshbse_car)  
3. fish that are not known to be in the region/found outside the region (bdfshbse_out)  
4. fish that were not found in the databse (bdfshbse_notfound)

```{r}
bdsp <- unique(newbdtaxclass$Species)[!unique(newbdtaxclass$Species)%in%unique(newmartax$Species)] #species not in mar db
bdmarsp <- unique(newbdtaxclass$Species)[unique(newbdtaxclass$Species)%in%unique(newmartax$Species)] #species in mar db
```

check that those numbers add up    
```{r}
#bd
length(unique(newbdtaxclass$Species)) #total = 
length(bdsp) # taxa not in database +
length(bdmarsp) # taxa in database
# +- 1 for NA 
```

Query fishbase  
Use list of species not in the MAR masked alias database (bdsp)  
```{r}
bdfshbsecountrydf <- rfishbase::country(
  species_list = bdsp,#spp,
  #fields = NULL,
  server = c("fishbase", "sealifebase"),
  version = "latest"
  #db = NULL,
)
```

Get lists of species from fishbase output  
bdfshbse_notfound: species not found in fishbase  
bdfshbse_mar: species found in the MAR  
bdfshbse_car: species not found in the MAR but found in the Caribbean  
bdfshbse_out: species not recorded in the Caribbean 
```{r}
#list of species not found in fishbase  
bdfshbse_notfound <- bdsp[!bdsp%in%unique(bdfshbsecountrydf$Species)]

#list of species found in MAR  
bdfshbse_mar <- bdfshbsecountrydf[bdfshbsecountrydf$country %in% c("Belize","Mexico","Guatemala","Honduras"),] %>%
  .$Species %>% 
  unique()

#list of species found in caribbean countries 
bdfshbse_car <- bdfshbsecountrydf[str_detect(bdfshbsecountrydf$country,
                                     paste(carcountries,collapse="|")),] %>% 
  filter(!Species%in%bdfshbse_mar) %>% #dont include species already identified as being in a MAR country
  .$Species %>% 
  unique() 

#Species not recorded in the Caribbean  
bdfshbse_out <- bdfshbsecountrydf %>% 
  filter(!Species%in%bdfshbse_mar) %>% #don't include species found in MAR coubdries
  filter(!Species%in%bdfshbse_car) %>% #don't include species found in caribbean countries 
  .$Species %>% 
  unique() %>%
  sort()
```

Confirm numbers add up  
```{r}
#total num species
length(unique(newbdtaxclass$Species))
sum(length(bdmarsp),length(bdsp)) #sum of species found in MAR db and those not found in MAR db  

#num species queried to fishbase  
length(bdsp) #bdsp was used to query fishbase  
sum(length(bdfshbse_notfound), #species not found in fishbase  
    length(bdfshbse_mar), #species found in MAR
    length(bdfshbse_car), #species found in Caribbean
    length(bdfshbse_out) #species found outside of Caribbean  
    )

#total num species 
length(unique(newbdtaxclass$Species))
sum(length(bdmarsp), #species found in MAR database
    length(bdfshbse_notfound), #species not found in fishbase  
    length(bdfshbse_mar), #species found in MAR
    length(bdfshbse_car), #species found in Caribbean
    length(bdfshbse_out) #species found outside of Caribbean  
    )
```
```{r}
bdmarsp <-  bdmarsp[!is.na(bdmarsp)] #species found in MAR database
bdfshbse_notfound <- bdfshbse_notfound[!is.na(bdfshbse_notfound)] #species not found in fishbase  
bdfshbse_mar <- bdfshbse_mar[!is.na(bdfshbse_mar)] #species found in MAR
bdfshbse_car <- bdfshbse_car[!is.na(bdfshbse_car)] #species found in Caribbean
bdfshbse_out <- bdfshbse_out[!is.na(bdfshbse_out)] #species found outside of Caribbean  
```

```{r}
accuracy.bd <- ps.bold.melt %>%
  filter(Class=="Actinopteri" | Class =="Actinopterygii") %>% 
  mutate(FalseAssgn = case_when(Species%in%bdmarsp ~ "Known in MAR",
                                Species%in%bdfshbse_mar ~ "Known in MAR",
                                Species%in%bdfshbse_car ~ "Known in Caribbean",
                                Species%in%bdfshbse_out ~ "Geographically Unlikely", 
                             Species%in%bdfshbse_notfound ~ "Not in Fishbase",
                             is.na(Species) ~ NA,
                             .default="Not a fish species"
                             ),.after=Species)
```
```{r}
accuracy.bd.plot.df <- accuracy.bd %>% 
  select(-c(1:17)) %>% 
  distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
               names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,
                     levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         FalseAssgn=case_when(Rank=="Species" ~ FalseAssgn, 
                              #remove false assignments from everything other than species  
                              .default=NA)
         ) %>% 
  distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn,FalseAssgn) %>% 
  reframe(rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn,
          FalseAssgn=FalseAssgn) %>% 
  distinct() %>% 
  group_by(Rank) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(FinalNum=case_when(isassgn=="Not Assigned" & is.na(FalseAssgn) ~ NotAssgn,
                            #isassgn=="Assigned" & is.na(FalseAssgn) ~ rows-NotAssgn,
                            isassgn=="Assigned" & is.na(FalseAssgn) ~ uniqranks-NotAssgn,
                            #isassgn=="Not Assigned" & is.na(FalseAssgn) ~ (rows-sum(numuniqassgn)+1),
                            #uniqranks==numuniqassgn ~ rows,
                            .default=numuniqassgn),
        FalseAssgn=factor(FalseAssgn,
                     levels=c(NA,"Not a fish species",
                              "Not in Fishbase",
                              "Geographically Unlikely",
                              "Known in Caribbean","Known in MAR"),ordered=TRUE)
         ) %>% 
  ungroup() %>% 
  group_by(Rank) %>% 
  arrange(desc(FalseAssgn),
          .by_group = TRUE) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(,
         LabelNum=(cumsum(FinalNum)))
```

### Combined database   
Using the new dataframe of only fish and elasmo species,  
Get a list of species that are NOT in the MAR masked alias    

cbsp: by subsetting the taxa df we just made to keep only those NOT in the MAR alias dataset   
cbmarsp: Then get the list of species that ARE in the MAR dataset  

After querying fishbase using the cbsp list, we will sucbivide the species into fish:  
1. known to occur in the MAR (cbmarsp + cbfshbse_mar)  
2. fish from the Caribbean (cbfshbse_car)  
3. fish that are not known to be in the region/found outside the region (cbfshbse_out)  
4. fish that were not found in the databse (cbfshbse_notfound)

```{r}
cbsp <- unique(newcbtaxclass$Species)[!unique(newcbtaxclass$Species)%in%unique(newmartax$Species)] #species not in mar db
cbmarsp <- unique(newcbtaxclass$Species)[unique(newcbtaxclass$Species)%in%unique(newmartax$Species)] #species in mar db
```

check that those numbers add up    
```{r}
#cb
length(unique(newcbtaxclass$Species)) #total = 
length(cbsp) # taxa not in database +
length(cbmarsp) # taxa in database
# +- 1 for NA 
```

Query fishbase  
Use list of species not in the MAR masked alias database (cbsp)  
```{r}
cbfshbsecountrydf <- rfishbase::country(
  species_list = cbsp,#spp,
  #fields = NULL,
  server = c("fishbase", "sealifebase"),
  version = "latest"
  #db = NULL,
)
```

Get lists of species from fishbase output  
cbfshbse_notfound: species not found in fishbase  
cbfshbse_mar: species found in the MAR  
cbfshbse_car: species not found in the MAR but found in the Caribbean  
cbfshbse_out: species not recorded in the Caribbean 
```{r}
#list of species not found in fishbase  
cbfshbse_notfound <- cbsp[!cbsp%in%unique(cbfshbsecountrydf$Species)]

#list of species found in MAR  
cbfshbse_mar <- cbfshbsecountrydf[cbfshbsecountrydf$country %in% c("Belize","Mexico","Guatemala","Honduras"),] %>%
  .$Species %>% 
  unique()

#list of species found in caribbean countries 
cbfshbse_car <- cbfshbsecountrydf[str_detect(cbfshbsecountrydf$country,
                                     paste(carcountries,collapse="|")),] %>% 
  filter(!Species%in%cbfshbse_mar) %>% #dont include species already identified as being in a MAR country
  .$Species %>% 
  unique() 

#Species not recorded in the Caribbean  
cbfshbse_out <- cbfshbsecountrydf %>% 
  filter(!Species%in%cbfshbse_mar) %>% #don't include species found in MAR coucbries
  filter(!Species%in%cbfshbse_car) %>% #don't include species found in caribbean countries 
  .$Species %>% 
  unique() %>%
  sort()
```

Confirm numbers add up  
```{r}
#total num species
length(unique(newcbtaxclass$Species))
sum(length(cbmarsp),length(cbsp)) #sum of species found in MAR db and those not found in MAR db  

#num species queried to fishbase  
length(cbsp) #cbsp was used to query fishbase  
sum(length(cbfshbse_notfound), #species not found in fishbase  
    length(cbfshbse_mar), #species found in MAR
    length(cbfshbse_car), #species found in Caribbean
    length(cbfshbse_out) #species found outside of Caribbean  
    )

#total num species 
length(unique(newcbtaxclass$Species))
sum(length(cbmarsp), #species found in MAR database
    length(cbfshbse_notfound), #species not found in fishbase  
    length(cbfshbse_mar), #species found in MAR
    length(cbfshbse_car), #species found in Caribbean
    length(cbfshbse_out) #species found outside of Caribbean  
    )
```
```{r}
cbmarsp <-  cbmarsp[!is.na(cbmarsp)] #species found in MAR database
cbfshbse_notfound <- cbfshbse_notfound[!is.na(cbfshbse_notfound)] #species not found in fishbase  
cbfshbse_mar <- cbfshbse_mar[!is.na(cbfshbse_mar)] #species found in MAR
cbfshbse_car <- cbfshbse_car[!is.na(cbfshbse_car)] #species found in Caribbean
cbfshbse_out <- cbfshbse_out[!is.na(cbfshbse_out)] #species found outside of Caribbean  
```

```{r}
accuracy.cb <- ps.comb.melt %>% 
  filter(Class=="Actinopteri" | Class =="Actinopterygii") %>%
  mutate(FalseAssgn = case_when(Species%in%cbmarsp ~ "Known in MAR",
                                Species%in%cbfshbse_mar ~ "Known in MAR",
                                Species%in%cbfshbse_car ~ "Known in Caribbean",
                                Species%in%cbfshbse_out ~ "Geographically Unlikely", 
                             Species%in%cbfshbse_notfound ~ "Not in Fishbase",
                             is.na(Species) ~ NA,
                             .default="Not a fish species"
                             ),.after=Species)
```
```{r}
accuracy.cb.plot.df <- accuracy.cb %>% 
  select(-c(1:17)) %>% 
  distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  #mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
               names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,
                     levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         FalseAssgn=case_when(Rank=="Species" ~ FalseAssgn, 
                              #remove false assignments from everything other than species  
                              .default=NA)
         ) %>% 
  distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn,FalseAssgn) %>% 
  reframe(#rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn,
          FalseAssgn=FalseAssgn) %>% 
  distinct() %>% 
  group_by(Rank) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(FinalNum=case_when(isassgn=="Not Assigned" & is.na(FalseAssgn) ~ NotAssgn,
                            #isassgn=="Assigned" & is.na(FalseAssgn) ~ rows-NotAssgn,
                            isassgn=="Assigned" & is.na(FalseAssgn) ~ uniqranks-NotAssgn,
                            #isassgn=="Not Assigned" & is.na(FalseAssgn) ~ (rows-sum(numuniqassgn)+1),
                            #uniqranks==numuniqassgn ~ rows,
                            .default=numuniqassgn),
        FalseAssgn=factor(FalseAssgn,
                     levels=c(NA,"Not a fish species",
                              "Not in Fishbase",
                              "Geographically Unlikely",
                              "Known in Caribbean","Known in MAR"),ordered=TRUE)
         ) %>% 
  ungroup() %>% 
  group_by(Rank) %>% 
  arrange(desc(FalseAssgn),
          .by_group = TRUE) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  mutate(,
         LabelNum=(cumsum(FinalNum)))
```

## Plot  
```{r}
plotpatt <- c("stripe", #"Not a fish species"
              "crosshatch", #"Not in Fishbase" 
              "stripe", #"Geographically Unlikely"
              "circle",#"Known in Caribbean"
              "circle"#"Known in MAR" 
              )

plotcol <- c(#"#D95F02FF", #NA
             "darkblue", #"Not a fish species"
             #"#e29578", 
             "darkgreen", #"Not in Fishbase" 
             "white",#"Geographically Unlikely"
             "#5a189a",#"Known in Caribbean"
             "#1B9E77FF"#"Known in MAR" 
             )

plotden <- c(#10, #NA
             0.1, #"Not a fish species"
             0.5, #"Not in Fishbase" 
             0.4,#"Geographically Unlikely"
             0.8,#"Known in Caribbean"
             10#"Known in MAR" 
             )

names(plotpatt) <- names(plotcol) <- names(plotden) <- levels(accuracy.nt.plot.df$FalseAssgn)

legend.guide <- guide_legend(override.aes=list(pattern_fill=NA,pattern_color=NA,pattern_alpha=0))
#legend.guide <- guide_legend(override.aes=list(pattern_fill=NA,pattern_color=NA,pattern_alpha=NA,fill=NA))
legend.fill.guide <- guide_legend(override.aes=list(fill="#1B9E77FF"),
                                  nrow=1,
                                  theme = theme(#legend.margin=margin(2,2,2,2,"line"),
        legend.key.spacing.y=unit(0.5,"line"),
        legend.key.height=unit(1.5,"line"),
        legend.key.width=unit(1.5,"line")
        ) )
```

```{r}
accuracy.nt.plot.base <- accuracy.nt.plot.df %>% 
  filter(Rank!="Class" & Rank != "Order") %>% 
  ggplot(.,aes(fill=isassgn)
         ) +
  geom_bar(stat="identity",
           position="stack",
           color="black",
           aes(y=Rank,x=FinalNum,fill=isassgn)) +
  scale_fill_brewer(name="Taxonomy Assigned",palette="Dark2",guide="none") 

accuracy.bd.plot.base <- accuracy.bd.plot.df %>% 
  filter(Rank!="Class" & Rank != "Order") %>% 
  ggplot(.,aes(fill=isassgn)
         ) +
  geom_bar(stat="identity",
           position="stack",
           color="black",
           aes(y=Rank,x=FinalNum,fill=isassgn)) +
  scale_fill_brewer(name="Taxonomy Assigned",palette="Dark2",guide="none") 

accuracy.cb.plot.base <- accuracy.cb.plot.df %>% 
  filter(Rank!="Class" & Rank != "Order") %>% 
  ggplot(.,aes(fill=isassgn)
         ) +
  geom_bar(stat="identity",
           position="stack",
           color="black",
           aes(y=Rank,x=FinalNum,fill=isassgn)) +
  scale_fill_brewer(name="Taxonomy Assigned",palette="Dark2",guide="none") 
```

```{r}
accuracy.nt.plot <- accuracy.nt.plot.base  +
  ggpattern::geom_col_pattern(aes(y=Rank,x=FinalNum,
                                  pattern=FalseAssgn,
                                  pattern_color=FalseAssgn,
                                  pattern_fill=FalseAssgn
                                  ),
                              color="black",
                              pattern_density=0.3,
                              pattern_spacing=0.03,
                              pattern_alpha=0.8,
                              pattern_key_scale_factor=1,
                              na.rm=TRUE
                              ) + 
  scale_pattern_manual(name="Likely accuracy of assignment",values = plotpatt) + 
  scale_pattern_color_manual(name="Likely accuracy of assignment",values = plotcol) +  
  scale_pattern_fill_manual(name="Likely accuracy of assignment",values = plotcol) +  
  geom_label_repel(aes(x=(LabelNum), y=Rank, label=ifelse(FinalNum>5,FinalNum,NA)),
                   nudge_x=(-4),
                   min.segment.length = 0.2,
                  size=5,
                  label.size=1,
                   show.legend=FALSE
                   ) + 
  guides(fill=legend.guide,
         pattern_density=legend.fill.guide,
         pattern=legend.fill.guide,
         pattern_color=legend.fill.guide,
         pattern_fill=legend.fill.guide) 
  #theme(legend.position="right",
  #      legend.margin=margin(2,2,2,2,"line"),
  #      legend.key.spacing.y=unit(1,"line"),
  #      legend.key.height=unit(3,"line"),
  #      legend.key.width=unit(3,"line")
  #      ) 

accuracy.bd.plot <- accuracy.bd.plot.base +
  ggpattern::geom_col_pattern(aes(y=Rank,x=FinalNum,
                                  pattern=FalseAssgn,
                                  pattern_color=FalseAssgn,
                                  pattern_fill=FalseAssgn
                                  ),
                              color="black",
                              pattern_density=0.3,
                              pattern_spacing=0.03,
                              pattern_alpha=0.8,
                              pattern_key_scale_factor=1,
                              na.rm=TRUE
                              ) + 
  scale_pattern_density_manual(name="Likely accuracy of assignment",values=plotden) +
  scale_pattern_manual(name="Likely accuracy of assignment",values = plotpatt) + 
  scale_pattern_color_manual(name="Likely accuracy of assignment",values = plotcol) +  
  scale_pattern_fill_manual(name="Likely accuracy of assignment",values = plotcol) +  
  geom_label_repel(aes(x=(LabelNum), y=Rank, label=ifelse(FinalNum>5,FinalNum,NA)),
                  nudge_x=(-4),
                   min.segment.length = 0.2,
                   size=5,
                  label.size=1,
                   show.legend=FALSE
                   ) + 
  guides(fill=legend.guide,
         pattern=legend.fill.guide,
         pattern_color=legend.fill.guide,
         pattern_fill=legend.fill.guide)

accuracy.cb.plot <- accuracy.cb.plot.base +
  ggpattern::geom_col_pattern(aes(y=Rank,x=FinalNum,
                                  pattern=FalseAssgn,
                                  pattern_color=FalseAssgn,
                                  pattern_fill=FalseAssgn
                                  ),
                              color="black",
                              pattern_density=0.3,
                              pattern_spacing=0.03,
                              pattern_alpha=0.8,
                              pattern_key_scale_factor=1,
                              na.rm=TRUE
                              ) + 
  scale_pattern_manual(name="Taxonomic \n assignment",values = plotpatt) + 
  scale_pattern_color_manual(name="Taxonomic \n assignment",values = plotcol) +  
  scale_pattern_fill_manual(name="Taxonomic \n assignment",values = plotcol) +  
  geom_label_repel(aes(x=(LabelNum), y=Rank, label=ifelse(FinalNum>5,FinalNum,NA)),
                   nudge_x=(-4),
                   min.segment.length = 0.2,
                   size=5,
                  label.size=1,
                   show.legend=FALSE
                   )+ 
  guides(#fill=legend.guide,
         pattern=legend.fill.guide,
         pattern_color=legend.fill.guide,
         pattern_fill=legend.fill.guide) +
  labs(x="Number of fish taxa")
```
```{r}
accuracy.cb.plot 
#ggsave("combinedDatabase.accuracy.png",accuracy.cb.plot,width=10,height=5)
```

```{r,warning=FALSE}
wrap_plots((accuracy.nt.plot + theme(legend.position="none",axis.title.x=element_blank()) & scale_x_continuous(expand=c(0,NA),limits=c(0,200))),
           (accuracy.bd.plot + theme(legend.position="none",axis.title.x=element_blank()) & scale_x_continuous(expand=c(0,NA),limits=c(0,200))),
           (accuracy.cb.plot & scale_x_continuous(expand=c(0,NA),limits=c(0,200))),
           guides = "collect",
           nrow=3) + 
  plot_annotation(tag_levels = "A", 
                  caption="A) Annotated with the NCBI nt database; B) Annotated with the BOLD v5 database;
                  C) Annotated with both NCBI and BOLD and retaining only taxonomic assignments without disagreement")

#ggsave("databaseComparison.png",width=10,height=7)
```
```{r}
comp.mar.plot.df <- ps.mar.melt %>% 
  select(-c(1:17)) %>% 
  distinct() %>% #get a unique dataframe; should have one row for each species, one for each genus, etc  
  mutate(rows=nrow(.)) %>%
  pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
               names_to="Rank") %>%
  mutate(isassgn = case_when(is.na(value) ~ "Not Assigned",
                             !is.na(value) ~ "Assigned"),
         Rank=factor(Rank,
                     levels=c("Species","Genus","Family","Order","Class"))
         ) %>% 
  group_by(Rank) %>% 
  mutate(NotAssgn=sum(isassgn=="Not Assigned"),
         uniqgroups=length(unique(value)), #number of unique members of that rank
         ) %>% 
  distinct() %>% #collapse down to unique rows
  group_by(Rank,isassgn) %>% 
  reframe(rows=rows, #size of df
          NotAssgn=NotAssgn, #how many individuals in each group are not assigned
          numuniqassgn=length(unique(value)), #number of assigned (should stay the same except for species and any ranks with NA values)
          uniqranks=uniqgroups,# number of unique members of that rank
          isassgn=isassgn) %>% 
  distinct() %>% 
  group_by(Rank) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE) %>%
  ungroup() %>% 
  group_by(Rank) %>% 
  arrange(desc(isassgn),
          .by_group = TRUE)  %>%
  mutate(Place=case_when(isassgn=="Not Assigned" ~ NotAssgn,
                            isassgn=="Assigned"  ~ NotAssgn+numuniqassgn,
                            .default=numuniqassgn),
         Label=case_when(isassgn=="Not Assigned" ~ NotAssgn,
                            .default=numuniqassgn)
         ) %>%  
  filter(Rank!="Class" & Rank != "Order") 


comp.mar.plot <- comp.mar.plot.df %>% 
   ggplot(.,aes(fill=isassgn)
         ) +
  geom_bar(stat="identity",
           position="stack",
           color="black",
           aes(y=Rank,x=Label,fill=isassgn)) +
  scale_fill_brewer(name="Taxonomy Assigned",palette="Dark2",guide="none") +  
  geom_label_repel(aes(x=(Place), y=Rank, label=Label),
                   nudge_x=(-4),
                   #min.segment.length = 0.2,
                  size=5,
                  label.size=1,
                   show.legend=FALSE
                   ) 
comp.mar.plot
```
```{r,warning=FALSE}
wrap_plots((accuracy.nt.plot + 
              scale_x_continuous(expand=c(0,NA),limits=c(0,175)) + 
              theme(axis.title.x=element_blank(),legend.position = "none")
            ),
           (comp.mar.plot + 
              scale_x_continuous(expand=c(0,NA),limits=c(0,175)) + 
              theme(axis.title.x=element_blank(),legend.position = "none")
            ),
           plot_spacer(),
           (accuracy.cb.plot & scale_x_continuous(expand=c(0,NA),limits=c(0,175))),
           guides = "collect",
           nrow=4
           ) + 
  plot_annotation(tag_levels = "A", 
                  caption="A) Annotated with the NCBI nt database; B) Annotated with a masked NCBI alias restricted to only fish species known to be present in the MAR; 
                  C) Annotated with both NCBI and BOLD and retaining only taxonomic assignments without disagreement
                  
                  Patterning indicates likelihood that a species could be found in the MAR as determined through a search of known countries in FishBase on 10 Sept 2025.")

#ggsave("databaseComparison.nt.mar.comb.png",width=10,height=7)
```

```{r}
wrap_plots(
  n.asv.plot, 
  (accuracy.cb.plot + ggpubr::theme_pubclean() & scale_x_continuous(expand=c(0,NA),limits=c(0,175))),
  nrow=2
)+ 
  plot_annotation(tag_levels = "A", 
                  caption="A) Number of assigned taxa at each level. For all ranks except Class, color indicates whether the organism is assigned to that level; 
                  B) Annotated with both NCBI and BOLD and retaining only taxonomic assignments without disagreement
                  Patterning indicates likelihood that a species could be found in the MAR 
                  as determined through a search of known countries in FishBase on 10 Sept 2025.")

ggsave("databaseComparison.asv.comb.png",width=12,height=7)
```

```{r}
wrap_plots((accuracy.cb.plot+ scale_x_continuous(expand=c(0,NA),limits=c(0,175))),
           (comp.mar.plot + scale_x_continuous(expand=c(0,NA),limits=c(0,175)) + theme(axis.title.x = element_blank())),
           nrow=2
           ) + 
  plot_annotation(tag_levels = "A", 
                  caption="A) Annotated with both NCBI and BOLD and retaining only taxonomic assignments without disagreement; B) Annotated with a masked NCBI alias restricted to only fish species known to be present in the MAR
                   Patterning indicates likelihood that a species could be found in the MAR as determined through a search of known countries in FishBase on 10 Sept 2025.")

#ggsave("databaseComparison.mar.png",width=10,height=7)
```

```{r}
ps.comb.melt  %>% 
  filter(Class=="Actinopteri" | Class=="Actinopterygii") %>%
  select(c(OTU,Phylum,Class,Order,Family,Genus,Species)) %>% 
  distinct() %>% 
  group_by(Genus) %>% 
  reframe(n_asv=n(),
          n_na=sum(is.na(Species)),
          n_uniq=length(unique(Species[!is.na(Species)]))) %>% 
  filter(n_na>0 & !is.na(Genus)) %>%
  pivot_longer(cols=c(n_na,n_uniq),names_to="Number") %>%
  mutate(Number=case_when(Number=="n_na" ~ "Not assigned",.default="Unique species")) %>%
  ggplot(aes(x=Genus,y=value,fill=Number)) + 
  geom_bar(stat="identity") + 
  theme_bw() + 
  scale_fill_brewer(name="ASVs assigned to species level",palette="Paired") +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5)) +
  scale_y_continuous(name="Number of ASVs",breaks=seq(from=0,to=14,by=2),expand=c(0,0))
  #guides(fill=guide_colorbar(barwidth=16,title="Number of ASVs \n not assigned to species"))+
  #labs(y="Number of ASVs",x="Genus")
  #filter(!is.na(Genus) & is.na(Species)) %>% distinct()
```
```{r}
ps.comb.melt  %>% 
  filter(Class=="Actinopteri" | Class=="Actinopterygii") %>%
  select(c(OTU,Phylum,Class,Order,Family,Genus,Species)) %>% 
  distinct() %>% 
  group_by(Genus) %>% 
  reframe(n_asv=n(),
          n_na=sum(is.na(Species)),
          n_uniq=length(unique(Species[!is.na(Species)]))) %>% 
  filter(n_na>0 & !is.na(Genus)) %>%
  pivot_longer(cols=c(n_na,n_uniq),names_to="Number") %>%
  mutate(Number=case_when(Number=="n_na" ~ "Not assigned",.default="Unique species")) %>% 
  treemap::treemap(.,index=c("Genus","Number"),
                   vSize="value",
                   #sortID = "color",sortID = "size",
                   type="categorical",
                   vColor="Number",
                   fontsize.labels = c(12,0),
                   bg.labels = "transparent",
                   align.labels=c("left","top"),
                   overlap.labels = 0.8
                   )
```

```{r}
nassigned.mar.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) +
  geom_bar(stat="identity",position="stack",color="black") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,value,NA))),hjust=2.5,
                   size=5,label.size=1,show.legend=FALSE) 

nassigned.bold.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) +
  geom_bar(stat="identity",position="stack",color="black") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,value,NA))),hjust=2.5,
                   size=5,label.size=1,show.legend=FALSE) 

nassigned.comb.df %>%
  ggplot(aes(y=Rank,x=value,fill=val)) +
  geom_bar(stat="identity",position="stack",color="black") + 
  scale_fill_brewer(palette="Dark2") + 
  geom_label(aes(label=(ifelse(value>0 & value<275,value,NA))),hjust=2.5,
                   size=5,label.size=1,show.legend=FALSE) 
```

# Summary   

## A comparison of database assignment numbers    
How many chordate ASVs identified with the nt data base?  
And how many were assigned to species level?  
```{r}
ntaxa(ps.nt)
ntaxa(ps.nt %>% subset_taxa(is.na(Species)) %>% prune_taxa(taxa_sums(.)>0,.) )
ntaxa(ps.nt.c %>% subset_taxa(!is.na(Species)) %>% prune_taxa(taxa_sums(.)>0,.) )

ntaxa(ps.nt %>% subset_taxa(is.na(Genus)) %>% prune_taxa(taxa_sums(.)>0,.) )
```
386 total ASVs
334/386 = 86.5% of ASVs could not be assigned to species level  
26/386 = 6.7% of ASVs could not be assigned to genus level  

```{r}
tax_table(ps.nt.all2 %>% subset_taxa(Phylum=="Chordata")) %>% 
  as.matrix() %>% as.data.frame() %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()

ps.nt.all2 %>% 
  subset_taxa(Phylum=="Chordata") %>% 
  subset_taxa(!is.na(Species)) %>%
  prune_taxa(taxa_sums(.)>0,.) %>% 
  get_taxa_unique("Species") %>% 
  length()

accuracy.nt2.plot.df
```


Is that impacted with the masked database?  
```{r}
ntaxa(ps.mar)
ntaxa(ps.mar %>% subset_taxa(is.na(Species)) %>% prune_taxa(taxa_sums(.)>0,.) )

#ntaxa(ps.mar %>% subset_taxa(is.na(Genus)) %>% prune_taxa(taxa_sums(.)>0,.) )
```
287 ASVs 
246/287=85.7% of ASVs could not be assigned to species level  
100% of ASVs could be assigned to genus level  

BOLD 
```{r}
tax_table(ps.bold %>% subset_taxa(Phylum=="Chordata")) %>% 
  as.matrix() %>% as.data.frame() %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```

Finally, the combined dataset (all chordates)  
```{r}
tax_table(ps.comb) %>% 
  as.matrix() %>% as.data.frame() %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```

```{r}
ntaxa(ps.comb)
ntaxa(ps.comb %>% subset_taxa(is.na(Species)) %>% prune_taxa(taxa_sums(.)>0,.) )

ntaxa(ps.comb %>% subset_taxa(is.na(Genus)) %>% prune_taxa(taxa_sums(.)>0,.) )
```
679 ASVs
213/679 = 31.4% of ASVs not assigned to Species level  
54/679 = 8% of ASVs not assigned to Genus level  

How does this change when only looking at fish taxa?  
```{r}
tax_table(ps.comb %>% subset_taxa(Class=="Actinopteri" | Class =="Actinopterygii")) %>% 
  as.matrix() %>% as.data.frame() %>%
  select(c(Class,Order,Family,Genus,Species)) %>% 
  reframe(across(everything(),
           ~ is.na(.))) %>% summary()
```
 451 fish asvs  
 16/451 genera unassigned (3.55%)
 97/451 species unassigned (21.51%)  

## A comparison of accuracy  

At the taxon level,  
```{r}
#ps.comb.melt %>%
psmelt(ps.comb) %>%
  select(-c(1:2,5:17)) %>% 
  group_by(Class,Order,Family,Genus,Species) %>% 
  reframe(Abundance=sum(Abundance))  %>% 
pivot_longer(cols=c("Class","Order","Family","Genus","Species"),
             names_to="Rank") %>% 
  group_by(Rank) %>% 
  reframe("Assigned"=sum(!is.na(value)),
          "Not Assigned"=sum(is.na(value)))

accuracy.cb.plot.df %>% select(-c(uniqranks,FinalNum,LabelNum))
```
47 species not identified
111 species known in the mar 
6 species known in caribbean
2 not in fishbase

12 genera not assigned
89 assigned

52 families assigned 

~26.7% of species cound not be identified  (47/(47+111+6+12))
4.54% of species that were identified were geographically unlikely (8/(47+111+6+12))

11.9% of genera could not be identified  (12/(12+89))

When looking at the masked database:  
```{r}
comp.mar.plot.df %>% select(-c(rows,Place))
```
we see less species being annotated and a lower diversity in general  

```{r}
accuracy.nt.plot.df
accuracy.nt2.plot.df
accuracy.bd.plot.df 
accuracy.cb.plot.df 
```



# Session Info  
```{r}
save.image("RData/OceanDNAprop.v2.RData")
sessionInfo()
```

